### 2017-05-16 elsfrontmatter.tex
INITIAL COMMIT - this has a full paper build

\begin{frontmatter}

\title{A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models}

%% Group authors per affiliation:
\author[iiasa,uw]{Matthew J. Gidden\corref{corref}}
\ead{gidden@iiasa.ac.at}
\cortext[corref]{Corresponding author}
\address[iiasa]{International Institute for Applied Systems Analysis,
  Schlossplatz 1, A-2361 Laxenburg, Austria}

\begin{abstract}
\TODO do this \cite{Dirac1953888}
\end{abstract}

\begin{keyword}
\TODO do this
\end{keyword}

\end{frontmatter}

\linenumbers

### 2017-05-16 paper.tex
INITIAL COMMIT - this has a full paper build

%%% ELS includes
\documentclass[review]{elsarticle}
\usepackage{lineno,hyperref}
\modulolinenumbers[5]
\journal{Environmental Modelling \& Software}
\bibliographystyle{elsarticle-num}
%%%

%%% General usepackages, definitions, etc.
\input{preamble}

%%% General setup
\graphicspath{{./figs/}}
%\include{acros}
\makeglossaries

\begin{document}

%%% ELS input
\input{elsfrontmatter} 
%%%

\newpage
stuff
%% \input{intro}
%% \input{methods}
%% \input{results}
%% \input{conclusion}
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}

fill in

\newpage
\section*{\refname}
\bibliography{refs}
\end{document}

### 2017-05-17 meth.tex
starting to work on methods

\section{Methodology}

IAM emission results are provided along a number of different dimensions,
including, temporal (normally half decade or decade), spatial (i.e., model
regions), gas species, and sectoral. Each individual temporal trajectory (i.e.,
unique combinations of regions, species, and sectors) must be harmonized to the
initial time period. Given a model trajectory, $m_{r, g, s}(t)$, historical
trajectory, $h_{r, g, s}(t)$, and initial time period $t_i$, a harmonized
trajectory can be calcluated.

In previous studies, two \textit{families} of methods have been used: those that
operate on the ratio of base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and
those that operate on the offset of base year values (i.e., $h(t_i) -
m(t_i)$). The \code{aneris} code base provides a number of variations on the
classic methods, including ratio-convergence methods shown in Equation
\ref{eqs:ratio}, offset-convergence methods shown in Equation \ref{eqs:offset},
and interpolation methods shown in Equation \ref{eqs:interpolate}; the
convergence factor for associated methods is shown first in Equation
\ref{eqs:factor}. Each is a function of time, model trajectory, historical
trajectory, harmonization time period, and, potentially, a convergence time
period ($t_f$), where the harmonized trajectory is equivalent to the
unharmonized trajectory.

\begin{equation}\label{eqs:factor}
  \beta(t, t_i, t_f) =
  \begin{cases}
    1 - \frac{t - t_i}{t_f - t_i},& \text{if } t \leq t_f\\
    0,                        & \text{otherwise}
  \end{cases}
\end{equation}

\begin{equation}\label{eqs:ratio}
  m^{rat}(t, m, h, t_i, t_f) = [\beta(t, t_i, t_f) (\frac{h(t_i)}{m(t_i)} - 1) + 1] m(t)
\end{equation}

\begin{equation}\label{eqs:offset}
  m^{off}(t, m, h, t_i, t_f) = \beta(t, t_i, t_f) (h(t_i) - m(t_i)) + m(t)
\end{equation}

\begin{equation}\label{eqs:interpolate}
  m^{int}(t, m, h, t_i, t_f) =
  \begin{cases}
    \frac{m(t_f) - h(t_i)}{t_f - t_i}(t - t_i) + h(t_i), & \text{if } t \leq t_f\\
    m(t), & \text{otherwise}
  \end{cases}
\end{equation}

\section{Implementation}

foo

### 2017-05-18 meth.tex
adding cov figure

@@
-\section{Methodology}
+\section{Methodology \& Implementation}
 
@@
 \ref{eqs:factor}. Each is a function of time, model trajectory, historical
-trajectory, harmonization time period, and, potentially, a convergence time
-period ($t_f$), where the harmonized trajectory is equivalent to the
-unharmonized trajectory.
+trajectory, harmonization time period, and a convergence time period ($t_f$),
+where the harmonized trajectory converges to the unharmonized trajectory.
 
@@
 
-\section{Implementation}
+While historically expert opinion has guided the determination of harmonization
+methods to choose for a given trajectory, a new, \textit{decision tree} approach
+has been implemented in this work. In order to provide reasonable
+\textit{default} methods, the historical trajectory, unharmonized model
+trajectory, and relative difference between history and model values in the
+harmonization year are analyzed. A figure of the decision tree used in this
+analysis is shown in Figure \ref{}.
 
-foo
+% decision tree figure
+
+A number of characteristics impact the decision of which default method to
+select based on the effect of the characteristic on the potential harmonized
+trajectory. For example, it is possible for models to report zero values in the
+harmonization year in situations in which technologies are introduced in future
+time periods in regions or for sectors which produce an emissions species that
+is absent in the initial modeling period. In such cases, an offset method is
+required as a ratio method would mask future emissions and erroneously harmonize
+the trajectory.
+
+Variability of the historical trajectory is also an important characteristic
+when considering the choice of harmonization method. Such trajectories are
+normally indicative of emissions from land use sectors. Land use sectors are
+modeled with varying fidelity by IAMs, but in general proxies are used for
+emissions trajectories. Take for example the emissions of black carbon (BC) from
+grassland burning in Africa \TODO{check that this is a good sector}: for most
+models, the use of grass and paturelands is modeled, and the emissions of
+species from the burning of these lands is determined by applying a emissions
+factor to a proxy model of burning of these lands; in general, the more
+grassland usage occurs, the higher the emissions from burning of
+grasslands. Therefore, consistency in harmonization method is important due to
+the approaches taken in modeling these sectors by IAMs. In order to reliably
+detect land-use-like emissions, a measure of the covariance of the historical
+trajectory is calculated as shown in Equation \ref{eqs:cov} and tested against a
+threshold. To determine this threshold, an analysis of the recent CEDS
+\TODO{cite ceds paper} historical data has been performed. Figure \ref{fig:cov}
+shows the distribution of land-use covariances and non-land-use covariances as
+determined for \TODO{list regions for 11 regions}. The minimum covariance
+observed for land-use sectors is \TODO{find}, thus theshold used in this study
+is \TODO{do this}. Notably, some non-land-use sectors , however,
+
+\begin{equation}\label{eqs:cov}
+    \Omega =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
+\end{equation}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{cov.pdf}
+    \caption[]{
+      \label{fig:cov}
+      The distribution of $\Omega$ values for LUC and non-LUC historical
+      trajectories are shown. The CEDS historical data is used for non-LUC data
+      and \TODO{cite LUC data} is used for LUC data. All historical data has
+      been aggregated to the 11-RCP \TODO{correct nomenclature?, cite} regions,
+      and all gas species included in the historical datasets are included. The
+      solid black line indicates the threshold value chosen in this analysis.
+    }
+  \end{center}
+\end{figure}

### 2017-05-18 meth.tex
using all models for cov figure

@@
 \section{Methodology \& Implementation}
+
+\subsection{Harmonization Methods}
 
@@
 
+\subsection{Default Method Decision Tree}
+
 While historically expert opinion has guided the determination of harmonization
@@
 harmonization year are analyzed. A figure of the decision tree used in this
-analysis is shown in Figure \ref{}.
+analysis is shown in Figure \ref{fig:decision_tree}.
 
-% decision tree figure
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{decision_tree.png}
+    \caption[]{
+      \label{fig:decision_tree}
+      The default method decision tree used in the \code{aneris} software
+      library. Decision-making threshold are described below, and the methods
+      applied are described in Section \ref{sec:impl-meths}.}
+  \end{center}
+\end{figure}
 
@@
 required as a ratio method would mask future emissions and erroneously harmonize
-the trajectory.
+the trajectory. In most cases, however, models do report values in the
+harmonization year. When model and historical values are relatively close, a
+convergence method is provided in order to be as representative as possible to
+the underlying unharmonized model results. The final model-based characteristic
+of import is whether the model reports negative emissions. Such a case is
+possible for gas species which can be extracted from the enviornment, as is the
+case for CO2 in future scenarios with high mitigation strategies.
 
@@
 shows the distribution of land-use covariances and non-land-use covariances as
-determined for \TODO{list regions for 11 regions}. The minimum covariance
-observed for land-use sectors is \TODO{find}, thus theshold used in this study
-is \TODO{do this}. Notably, some non-land-use sectors , however,
+determined for historical data aggregated to 11 world regions \TODO{list regions
+  for 11 regions}. A threshold value of 10 has been chosen based on the
+11-region observations, but, importantly, any regional aggregation is possible
+as this is model depedent.
 
@@
 \end{figure}
+
+Finally, consideration is taken with respect to the relative difference between
+the historic and model values in the harmonization year.
+
+\subsection{Implemented Methods}\label{sec:impl-meths}

### 2017-05-18 meth.tex
updating cov with better threshold

@@
 shows the distribution of land-use covariances and non-land-use covariances as
-determined for historical data aggregated to 11 world regions \TODO{list regions
-  for 11 regions}. A threshold value of 10 has been chosen based on the
-11-region observations, but, importantly, any regional aggregation is possible
-as this is model depedent.
+determined for historical data aggregated to the model regions of 5 different
+flagship IAMs \TODO{cite all 5 IAMs}. A threshold value of 20 has been chosen
+based on these observations. Importantly, tails of the LUC and non-LUC overlap,
+thus there are both false positives (~10\% of non-LUC trajectories) and false
+negatives (~7\% of LUC trajectories). However, as any regional definition is
+model depedent and thus any regional aggregation is possible an automated
+detection mechanism is necessary.
 
@@
       The distribution of $\Omega$ values for LUC and non-LUC historical
-      trajectories are shown. The CEDS historical data is used for non-LUC data
-      and \TODO{cite LUC data} is used for LUC data. All historical data has
-      been aggregated to the 11-RCP \TODO{correct nomenclature?, cite} regions,
-      and all gas species included in the historical datasets are included. The
-      solid black line indicates the threshold value chosen in this analysis.
+      trajectories are shown. CEDS historical data is used for non-LUC data and
+      \TODO{cite LUC data} is used for LUC data. All historical data has been
+      aggregated from countries to IAM model regional definitions, and all gas
+      species included in the historical datasets are included in the
+      analysis. The solid black line indicates the threshold value used by
+      default in \code{aneris}.  
     }

### 2017-05-18 meth.tex
dh fig and analysis added

@@
 possible for gas species which can be extracted from the enviornment, as is the
-case for CO2 in future scenarios with high mitigation strategies.
+case for CO2 in future scenarios with high mitigation strategies. If a model
+provides a trajectory that transitions from positive to negative emissions, then
+a convergence method is used in order to guarantee capture of this
+transition. Otherwise, it is possible, given the relative values between history
+and model in the harmonization year, for a negative trajectory to be
+inappropriatley harmonized to positive, but decreasing, trajectory.
 
@@
       The distribution of $\Omega$ values for LUC and non-LUC historical
-      trajectories are shown. CEDS historical data is used for non-LUC data and
+      trajectories is shown. CEDS historical data is used for non-LUC data and
       \TODO{cite LUC data} is used for LUC data. All historical data has been
@@
 Finally, consideration is taken with respect to the relative difference between
-the historic and model values in the harmonization year.
+the historic and model values in the harmonization year. In order to investigate
+the possible values that these relative differences can take, the flagship model
+values used in the SSP \TODO{CITE} and (ongoing) CMIP6 intercomparison exercises
+are used. A distribution of these differences for all models in the study is
+presented in Figure \ref{fig:dh}. Given the available data, a threshold value of
+50\% was chosen to be used as a default in \code{aneris}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{dh.pdf}
+    \caption[]{
+      \label{fig:dh}
+      The distribution of relative differences between model and historical
+      values in the harmonization year is shown. The solid black line indicates
+      the threshold value used by default in \code{aneris}.  
+    }
+  \end{center}
+\end{figure}
 

### 2017-05-18 meth.tex
add method table

@@
 \end{equation}
+
+The \code{aneris} code base provides a family of methods to choose from for each
+of the harmonization cases. A summary of all available methods is provided below
+in Table \ref{tab:meths}.
+
+\begin{table}[]
+\centering
+\caption{All Harmonization Methods Provided in \code{aneris}}
+\label{my-label}
+\begin{tabular}{|l|l|l|}
+\hline
+Method Name                             & Harmonization Type & Convergence Year\\
+\hline
+\code{constant\_ratio}                  & ratio              & $t_f = \infty$\\
+\code{reduce\_ratio\_<year>}            & ratio              & $t_f = $\code{<year>}\\
+\code{constant\_offset}                 & offset             & $t_f = \infty$\\
+\code{reduce\_offset\_<year>}           & offset             & $t_f = $\code{<year>}\\
+\code{linear\_interpolate\_<year>}      & interpolation      & $t_f = $\code{<year>}\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 \end{figure}
-
-\subsection{Implemented Methods}\label{sec:impl-meths}

### 2017-05-18 meth.tex
adding methods section with table

@@
 \caption{All Harmonization Methods Provided in \code{aneris}}
-\label{my-label}
+\label{tab:meths}
 \begin{tabular}{|l|l|l|}
@@
 \end{figure}
+
+\subsection{\code{aneris} Workflow}
+
+The full \code{aneris} workflow is comprised of a number of components shown
+graphically in Figure \ref{fig:workflow}. Unharmonized model data and
+run-control configuration are read in via an Excel spreadsheet. Data is assumed
+to be in the \textit{IAM Consortium} format, i.e., using \code{Model},
+\code{Scenario}, \code{Region}, \code{Variable}, and \code{Unit} columns in
+addition to column representing each modeled time period. Users are able to
+control the harmonization process via a number of controls. First, users can
+provide \textit{override} methods for any combination of region and variable. If
+override methods are provided, they are used instead of the default methods as
+determined by the default method decision tree. Further, users can set the
+above-mentioned thresholds as well as the LUC method used in the decision
+tree. Further detail of input parameters can be found at
+\hyperlink{http://aneris.readthedocs.io/en/latest/}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[height=0.85\textheight]{aneris-workflow.pdf}
+    \caption[]{
+      \label{fig:workflow}
+      The full harmonization process as executed by \code{aneris}. Operations
+      that can be configured with user-based input configurations are shown in
+      purple. The core harmonization process is shown in orange.  }
+  \end{center}
+\end{figure}
+
+Input data then undergoes a cleaning operation, which adds any trajectories in
+the historical dataset but not provided by the model and detects any issues that
+would cause the harmonization process to fail. The methods used to harmonize the
+data are then determined and the harmonization process is executed. Upon
+completion of the harmonization process, aggregation of common analysis regions
+is performed (e.g., the 5 RCP regions \TODO{CITE}). Finally, any exogenous
+trajectories the user provides are added. Exogenous trajectories are normally
+provided for unmodeled gases with well-accepted scenario trajectories, e.g.,
+Chlorofloro Carbons provided by WMO \TODO{CITE}. Upon completion, the harmonized
+trajectories and metadata regarding the harmonization process are returned. A
+description of all returned metadata is provided in Table \ref{tab:metadata}.
+
+\begin{table}[]
+\centering
+\caption{Metadata provided by the \code{aneris} harmonization routine. This metadata is provided for every combination of region, sector, and emissions species.}
+\label{my-tab:metadata}
+\begin{tabular}{|p{2cm}|p{8cm}|}
+\hline
+Column       & Description                                                                  \\
+\hline
+\hline
+method       & The harmonization method used.                                               \\
+\hline
+default      & The default harmonization method as determined by the default decision tree. \\
+\hline
+override     & The method provided as an override (if any).                                 \\
+\hline
+offset       & The offset value between history and model in the harmonization year.        \\
+\hline
+ratio        & The ratio value between history and model in the harmonization year.         \\
+\hline
+cov          & The covariance value of the historical trajectory.                           \\
+\hline
+unharmonized & The unharmonized value in the harmonization year.                            \\
+\hline
+history      & This historical value in the harmonization year.                             \\
+\hline
+harmonized   & The resulting harmonized value in the harmonization year.\\
+\hline
+\end{tabular}
+\end{table}

### 2017-05-18 results.tex
adding all result files

\section{Results \& Conclusions}

\subsection{A Harmonization Test Case}

In order to test the \code{aneris} code base and harmonization procedure,
results from the IAM MESSAGE-GLOBIOM \TODO{CITE} was processed using the
previously described combined historical dataset. The SSP2-RCP4.5 scenario was
chosen a test cases as it incorporates moderate to heavy mitigation strategies,
and therefore produces negative CO2 emissions for some regions and sectors. In
general, SSP2 scenarios describe a ``middle of the road'' world within the SSP
framework context. The 4.5 designation implies a scenario that results in
approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$ warming effect by the end of
the century, due to both warming from Greenhouse Gases (GHGs) and flourinated
gases as well as cooling from aerosols.

\subsection{The Use of Harmonization Overrides}

\subsection{Disccusion \& Future Work}

### 2017-05-18 analysis/make_results.py

@@ -1,0 +1,11 @@
+import sys
+import pandas as pd
+sys.path.append('../../../iiasa/ceds_harmonization_analysis/')
+import utils
+
+x = utils.read_data('45-no-overrides_harmonized', ext='xlsx')
+y = utils.read_data(
+    'MESSAGE-GLOBIOM_SSP2-45-SPA2-V25_unharmonized_harmonized', ext='xlsx')
+z = pd.read_csv('read_data_just_ceds.csv')
+
+pd.concat([x, y, z]).to_csv('results.csv', index=False)

### 2017-05-19 analysis/_utils.py
about to rm files

@@ -1,0 +1,7 @@
+import sys
+
+pth = '../../../iiasa/ceds_harmonization_analysis/'
+sys.path.append(pth)
+
+
+from utils import *

### 2017-05-19 analysis/make_results.py

@@ -1,11 +1,16 @@
-import sys
 import pandas as pd
-sys.path.append('../../../iiasa/ceds_harmonization_analysis/')
-import utils
 
-x = utils.read_data('45-no-overrides_harmonized', ext='xlsx')
-y = utils.read_data(
-    'MESSAGE-GLOBIOM_SSP2-45-SPA2-V25_unharmonized_harmonized', ext='xlsx')
-z = pd.read_csv('read_data_just_ceds.csv')
+import _utils as utils
 
-pd.concat([x, y, z]).to_csv('results.csv', index=False)
+# read in latest snapshot, get only ceds and message results
+x = pd.read_csv(utils.pth + 'v25.csv')
+lower = x.REGION == x.REGION.str.lower()  # rm isos
+x = x[~lower]
+x = x[(x.SCENARIO == 'CEDS') | (x.MODEL.str.contains('MESSAGE'))]
+
+# you have to dump to csv first because read_data only reads files for now
+x.to_csv('results_v25.csv', index=False)
+y = utils.read_data('results_v25')
+z = utils.read_data('45-no-overrides_harmonized', ext='xlsx')
+z.columns = [str(_) for _ in z.columns]
+pd.concat([y, z]).to_csv('results.csv', index=False)

### 2017-05-19 analysis/make_results.py
rm make results

@@ -1,16 +1,0 @@
-import pandas as pd
-
-import _utils as utils
-
-# read in latest snapshot, get only ceds and message results
-x = pd.read_csv(utils.pth + 'v25.csv')
-lower = x.REGION == x.REGION.str.lower()  # rm isos
-x = x[~lower]
-x = x[(x.SCENARIO == 'CEDS') | (x.MODEL.str.contains('MESSAGE'))]
-
-# you have to dump to csv first because read_data only reads files for now
-x.to_csv('results_v25.csv', index=False)
-y = utils.read_data('results_v25')
-z = utils.read_data('45-no-overrides_harmonized', ext='xlsx')
-z.columns = [str(_) for _ in z.columns]
-pd.concat([y, z]).to_csv('results.csv', index=False)

### 2017-05-19 meth.tex
updating paper and meth
saving work, meth and results mostly done

@@
 trajectory, $h_{r, g, s}(t)$, and initial time period $t_i$, a harmonized
-trajectory can be calcluated.
+trajectory can be calculated.
 
@@
 
-\subsection{Default Method Decision Tree}
+\subsection{Default Method Decision Tree}\label{sec:tree}
 
@@
       The default method decision tree used in the \code{aneris} software
-      library. Decision-making threshold are described below, and the methods
-      applied are described in Section \ref{sec:impl-meths}.}
+      library. Decision-making threshold are described below.}
   \end{center}
@@
 of import is whether the model reports negative emissions. Such a case is
-possible for gas species which can be extracted from the enviornment, as is the
+possible for gas species which can be extracted from the environment, as is the
 case for CO2 in future scenarios with high mitigation strategies. If a model
@@
 and model in the harmonization year, for a negative trajectory to be
-inappropriatley harmonized to positive, but decreasing, trajectory.
+inappropriately harmonized to positive, but decreasing, trajectory.
 
@@
 the approaches taken in modeling these sectors by IAMs. In order to reliably
-detect land-use-like emissions, a measure of the covariance of the historical
+detect land-use-like emissions, a measure of the co-variance of the historical
 trajectory is calculated as shown in Equation \ref{eqs:cov} and tested against a
@@
 \TODO{cite ceds paper} historical data has been performed. Figure \ref{fig:cov}
-shows the distribution of land-use covariances and non-land-use covariances as
+shows the distribution of land-use co-variances and non-land-use co-variances as
 determined for historical data aggregated to the model regions of 5 different
@@
 negatives (~7\% of LUC trajectories). However, as any regional definition is
-model depedent and thus any regional aggregation is possible an automated
+model dependent and thus any regional aggregation is possible an automated
 detection mechanism is necessary.
@@
 the possible values that these relative differences can take, the flagship model
-values used in the SSP \TODO{CITE} and (ongoing) CMIP6 intercomparison exercises
+values used in the SSP \TODO{CITE} and (ongoing) CMIP6 inter-comparison exercises
 are used. A distribution of these differences for all models in the study is
@@
 \caption{Metadata provided by the \code{aneris} harmonization routine. This metadata is provided for every combination of region, sector, and emissions species.}
-\label{my-tab:metadata}
+\label{tab:metadata}
 \begin{tabular}{|p{2cm}|p{8cm}|}

### 2017-05-19 results.tex
result figures generated
saving work, meth and results mostly done

@@
 
-In order to test the \code{aneris} code base and harmonization procedure,
-results from the IAM MESSAGE-GLOBIOM \TODO{CITE} was processed using the
-previously described combined historical dataset. The SSP2-RCP4.5 scenario was
-chosen a test cases as it incorporates moderate to heavy mitigation strategies,
-and therefore produces negative CO2 emissions for some regions and sectors. In
-general, SSP2 scenarios describe a ``middle of the road'' world within the SSP
-framework context. The 4.5 designation implies a scenario that results in
+In order to test the \code{aneris} harmonization procedure, results from the IAM
+MESSAGE-GLOBIOM \TODO{CITE} was processed. Two scenarios from the Shared Socioeconomic
+Pathways scenario library \TODO{CITE} are investigated. The reference SSP2, or
+``middle of the road'' is chosen first because MESSAGE-GLOBIOM is the marker
+scenario for this SSP. Notably, emissions from multiple sectors tend to increase
+in SSP2, thus testing the harmonization mechanism with a increasing emission
+pathway. Additionally, the SSP-45 scenario is chosen, where the ``45''
+designation refers to a scenario with end-of-century emissions resulting in
 approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$ warming effect by the end of
-the century, due to both warming from Greenhouse Gases (GHGs) and flourinated
-gases as well as cooling from aerosols.
+the century, due to both warming from Greenhouse Gases (GHGs) and fluoridated
+gases as well as cooling from aerosols. In this scenario, mitigation
+technologies and policies are enacted causing a general reduction in pollutants
+and GHGs, including (eventual) negative CO2 emissions in some regions and
+sectors due to carbon capture and sequestration and afforestation. This scenario
+explores the harmonization mechanism's response to generally decreasing emission
+trajectories.
 
-\subsection{The Use of Harmonization Overrides}
+MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
+mapped directly to the 5-region definition used in the RCPs. Historical data is
+taken from CEDS and LUC \TODO{get correct name}, which comprise 10 separate
+pollutant and GHG species and 13 sectors. Therefore, a total of 1430 distinct
+trajectories were harmonized.
 
-\subsection{Disccusion \& Future Work}
+The effect of harmonization overrides is analyzed by harmonizing each scenario
+first without overrides and then with overrides. Of the 1430 trajectories,
+approximately 2\% required the use of harmonization overrides after an initial
+investigation; thus, 98\% of all trajectories were satisfactorily determined
+using the default methods as chosen by the harmonization algorithm. Figure
+\ref{fig:nox} provides one such example of a emissions species and sector in
+which all regions were satisfactorily harmonized with the default methods. As
+can be seen, results with and without overrides are identical.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_NOx_Energy_Sector.pdf}
+    \caption[]{
+      \label{fig:nox}
+      NOx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
+      trajectories for SSP2 and SSP2-45 are presented. SSP-45 is denoted with
+      markers. Panel \textbf{a} shows 5-Region trajectories for scenarios
+      \textit{without} any overrides. Panel \textbf{b} shows 5-Region
+      trajectories for scenarios \textit{with} overrides. Panels \textbf{c} and
+      \textbf{d} show global trajectories without and with overrides,
+      respectively.  }
+  \end{center}
+\end{figure}
+
+The harmonization of emissions pathways is performed in order to accurately
+represent an updated historical emissions trajectory while also maintaining
+consistency with the original, unharmonized pathway. When the default methods as
+provided by the harmonization procedure distort or otherwise sufficiently
+misrepresent the underlying unharmonized results, an override method is required
+to be provided for the trajectory of the region, sector, and species in
+question. Based on the 2\% of total trajectories required to be overridden, two
+classifications were observed: regional trajectories whose \textit{magnitude}
+was overly distorted and regional trajectories whose \textit{shape} was overly
+distorted.
+
+Figure \ref{fig:co} presents a situation in which the magnitude of a trajectory
+is distorted. A large discrepancy (~300\% relative difference) is observed in
+the harmonization year for carbon monoxide (CO) emissions in the industrial
+sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region, which
+comprises most of the Asian subcontinent. The default method chosen
+(\code{constant_ratio} maintains model trends for the region, overall model
+results are distorted. By applying a \code{constant_offset} override, the
+regional trend and shape is maintained. With the new harmonization method for
+the SAS region, the global trajectory for industrial CO also corresponds more
+closely with the unharmonized trajectory.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_CO_Industrial_Sector.pdf}
+    \caption[]{
+      \label{fig:co}
+      CO Industrial Sector harmonized and unharmonized emissions are presented
+      for SSP2 and SSP2-45 scenarios. Scenarios as denoted identically to Figure
+      \ref{fig:nox}. Panels \textbf{a} and \textbf{b} show harmonized and
+      overridden-harmonized (respectively) regional trajectories for the 3
+      MESSAGE-GLOBIOM regions that comprise the R5ASIA region: Centrally Planned
+      Asia (CPA), Other Pacific Asia (PAS), and South Asia (SAS). Notably, the
+      SAS regional trajectory displays a distorted trajectory due to the
+      harmonization-year difference between history and model results in both
+      scenarios. The distortion is large enough to affect global results, as
+      shown in Panels \textbf{c} and \textbf{d}.  
+}
+  \end{center}
+\end{figure}
+
+In certain circumstances, the application of the default harmonization methods
+can affect not only the magnitude but also the shape of regional
+trajectories. Figure \ref{fig:nh3} shows emissions trajectories for ammonia
+(NH3) resulting from the agriculture sector in Asia. Again, the SAS region shows
+a large discrepancy in the harmonization year (> 150\% in this case). The
+resulting trajectory harmonized with a \code{constant_ratio} method by default
+provides a large increase after 2080 in the SSP2 reference scenario. Notably,
+the SSP2-45 scenario is not affected to the same degree. While this distortion
+affects the magnitude of the SAS trajectory, it largely affects the post-2080
+shape of the global trajectory (see Figure \ref{fig:nh3}, panel \textbf{c}). By
+using a \code{constant_offset} method as an override, this distortion is
+addressed and more accurately reflects unharmonized results both in the SAS
+region and global results for agricultural ammonia emissions.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_NH3_Agriculture.pdf}
+    \caption[]{
+      \label{fig:nh3}
+      NH3 Agriculture harmonized and unharmonized emissions are presented for
+      SSP2 and SSP2-45 scenarios. Scenarios and panel layouts are identical to
+      Figure \ref{fig:co}. In this case, the SAS trajectory again shows not only
+      a magnitude distortion, but also a shape distortion at the tail of the
+      trajectory. Override methods have been applied to correct the distortion.
+    }
+  \end{center}
+\end{figure}
+
+\subsection{Discussion \& Future Work}

### 2017-05-23 intro.tex
intro added

\section{Introduction}

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economoic systems, land use, the climate,
and air pollution, among other sectors \TODO{cite some IAMs}. IAMs provide
global projections of systemic change by dividing the world into a number of
representative regions, the definition of which is distinct for each
model. Results from IAMs are integral in a number of international studies,
which notably include projections of climate and economic futures used in
analyses underpinning the Intergovenmental Panel on Climate Change (IPCC)
\TODO{cite} and the United Nations Framework Convention on Climate Unchange
(UNFCCC)'s Paris Agreement \TODO{cite}.

While IAMs are implemented in myriad ways, including simulation and
optimization, the core inputs, outputs, and purpose are similar across different
models. Modeling teams incorporate existing datasets of energy systems, world
development indicators, and emissions sources and concentrations, among other
data, in order to provide consistent historical trajectories of modeled
variables. The models then provide estimates of future trajectories of these
variables under various socioeconomic and technological assumptions as well as
proposed policy constraints, e.g., targets for future Greenhouse Gas (GHG)
emissions.

Future emissions pathways are of critical importance to both of global and local
governance concerns. For example, the United Nation's Sustainable Development
Goals (SDGs) \TODO{CITE} speak directly to the impacts of emissions in Goal 3
(Good Health and Well-Being) and Goal 13 (Climate Action). IAMs endogenously
calculate many of the emissions species of import that impact the progress
towards these SDGs by modeling the individual technologies and sectors that
contribute towards the emissions of different air pollutants and GHGs as well as
various mitigation technologies (e.g., NOx and SOx scrubbers on power plants)
that are used due to regulation or implemented policies.

In practice, IAMs calculate the total source intensity of emitting technologies,
for example the total activity of coal power plant emissions in China, and
incorporate emissions-intensity factors for individual species, for example the
quantity of sulfur emissions from coal plants per megawatt-hour of production,
in order to determine total emissions of a given gas species from a given
technology. Models are then \textit{calibrated} to historical data sources which
provide both the activity of emitting technologies as well as trajectories of
past emissions in order to determine historical emissions factors in each model
region for each technology.

With some frequency, however, historical data is updated to incorporate new
assumptions of source activity, new emissions accounting practices, or other
concerns. The global climate change community has recently updated a standard
historical emissions dataset for both anthropogenic and land-use change
emissions \TODO{cite ceds and LUC paper} as part of the 6th phase of the Coupled
Model Intercomparison Project \TODO{cite cmip6} (CMIP6). These historical
emissions trajectories in conjunction with future trajectories as provided by
IAMs are critical inputs to the climate-related modeling exercises of CMIP6.

Because the process of model calibration is quite onerous, model teams initially
incorporate updated historical datasets through the process of
\textit{harmonization} in order to guarantee consistency with new historical
datasets. Harmonization describes the process of updating model results to
incorporate new historical timeseries while maintaining critical aspects of
modeled future trajectories for variables of import. In the emissions context,
this means that each individual combination of model region, model sector, and
emissions species must be harmonized. Depending on the model, this can require
the selection of thousands to tens-of-thousands of harmonization methods.

Harmonization has been addressed in previous studies as it is a common practice
in the IAM and climate change communities. For example, \TODO{cite Malte}
describes a harmonization process using scaling routines for the 5 SRES regions
with no sectoral definition. Further, \TODO{cite Joeri} describes the impacts of
choosing various harmonization routines on future trajectories. Importantly, the
choice of harmonization method to date has been determined via expert
solicitation and has generally been applied to all trajectories for a given
class of emissions species.

In order to incoporate the variety of possible model results and given
relatively large increases in regional and sectoral dimensionality, an automated
process for determining harmonization methods is needed. The remainder of this
paper describes the methodology of the \code{aneris} codebase which implements
such a system in Section \ref{sec:meths}. The results of applying the automated
harmonization mechanism on two example IAM scenarios, one with emissions growth
and another with emissions mitigation, is presented in Section
\ref{sec:results}. Finally, the general effectiveness and potential future
improvements on the automated methodology is discussed in Section
\ref{sec:future}.

% todo: add?
% what aspects of harmonization are important (e.g., cumulative carbon emissions).

### 2017-05-23 elsfrontmatter.tex
saving work, meth and results mostly done
rough draft of results section
rough draft of abstract complete

@@
 %% Group authors per affiliation:
-\author[iiasa,uw]{Matthew J. Gidden\corref{corref}}
+\author[iiasa]{Matthew J. Gidden\corref{corref}}
 \ead{gidden@iiasa.ac.at}
 \cortext[corref]{Corresponding author}
+
+\author[iiasa,nies]{Shinichiro Fujimori}
+
 \address[iiasa]{International Institute for Applied Systems Analysis,
   Schlossplatz 1, A-2361 Laxenburg, Austria}
+\address[nies]{National Institute for Environmental Studies, Tsukuba, Japan}
 
 \begin{abstract}
-\TODO do this \cite{Dirac1953888}
+Harmonization describes the process of calibrating Integrated Assessment Model
+(IAM) results with a new source of historical emissions trajectories. To date,
+harmonization has been performed separately by individual modeling teams through
+the use of expert opinion and solicitation. As models and their results become
+more complex in both regional and sectoral dimensions, an automated approach to
+choosing harmonization methods becomes necessary. This work describes a novel
+methodology for determining such harmonization methods and an open-source Python
+code base for implementing the methodology. Results are shown for two example
+scenarios using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
+of the total emissions trajectories. \cite{Dirac1953888}
 \end{abstract}
@@
 \begin{keyword}
-\TODO do this
+Integrated Assessment Modeling, Harmonization, Climate Change, Air Pollution 
 \end{keyword}

### 2017-05-23 future.tex
adding discussion section

\section{Discussion \& Future Work}\label{sec:future}

This work presented a novel methodology and Python implementation of automated
emissions harmonization for IAMs. An in-depth explanation of the processes and
methods for determining the use of harmonization methods was provided in Section
\ref{sec:meths}. The \code{aneris} codebase was able to satisfactorily
harmonized over 98\% of the 2860 individual trajectories that were analyzed in
Section \ref{sec:results}. Of the remaining trajectories, harmonization method
overrides were applied, and the situation in which the need for overrides arose
was discussed.

The automated approach critically removes the need for expert opinion in
determining harmonization methods for each individual combination of model
region, sector, and emissions species while still providing a justified reason
for each automated choice of harmonization method based on both the historical
and future emissions trajectories. Furthermore, the automated approach continues
to scale well as models become more detailed in both the regional and sectoral
dimensions. Finally, expert opinion is still allowed to trump the automated
method as determined by the algorithm via method overrides; however, these cases
are clearly documented via the metadata provided as an output of \code{aneris}
and thus can be individually explained. This provides not only transparency but
also scientific integrity in the choice of harmonization methods.

There are a variety of avenues for future improvement of both the \code{aneris}
codebase and underlying methdology. As with any software project, additional
users will provide use cases for more robust handling of input/output concerns
and corner cases. Further configuration parameters may also be added in the
future in order to provide overrides for all gas species in a given sector or
region. Perhaps the most fruitful investigation will involve further refinement
of the defaul decision tree introduce in Section \ref{sec:meths}. A key aspect
missing from the decision tree is input from models regarding whether missing
sources are the likely cause of a harmonization year discrepancy (suggesting the
use of an offset method) or instead a discrepancy in emissions factors
(suggesting the use of a ratio method)\TODO{cite Joeri}.

However, this work provides a new direction and framework in which the IAM and
climate communities can follow in order to reduce the necessicity of expert
opinion and increase transparency and reproducibility of harmonization
exercises. Furthermore, it provides an open-source, tested, and documented code
base which can be used and improved upon by these communities. Both of these are
clear steps in a positive direction for future climate and integrated assesment
modeling exercises.

### 2017-05-23 intro.tex
adding citations

@@
 interactions between energy systems, economoic systems, land use, the climate,
-and air pollution, among other sectors \TODO{cite some IAMs}. IAMs provide
-global projections of systemic change by dividing the world into a number of
-representative regions, the definition of which is distinct for each
-model. Results from IAMs are integral in a number of international studies,
-which notably include projections of climate and economic futures used in
-analyses underpinning the Intergovenmental Panel on Climate Change (IPCC)
-\TODO{cite} and the United Nations Framework Convention on Climate Unchange
-(UNFCCC)'s Paris Agreement \TODO{cite}.
+and air pollution, among other sectors \cite{van_vuuren_energy_2017,
+  kriegler_fossil-fueled_2017, calvin_ssp4:_2017, fujimori_ssp3:_2017,
+  fricko_marker_2017}. IAMs provide global projections of systemic change by
+dividing the world into a number of representative regions, the definition of
+which is distinct for each model. Results from IAMs are integral in a number of
+international studies, which notably include projections of climate and economic
+futures used in analyses underpinning the Intergovenmental Panel on Climate
+Change (IPCC) \TODO{cite} and the United Nations Framework Convention on Climate
+Unchange (UNFCCC)'s Paris Agreement \TODO{cite}.
 
@@
 Harmonization has been addressed in previous studies as it is a common practice
-in the IAM and climate change communities. For example, \TODO{cite Malte}
-describes a harmonization process using scaling routines for the 5 SRES regions
-with no sectoral definition. Further, \TODO{cite Joeri} describes the impacts of
-choosing various harmonization routines on future trajectories. Importantly, the
-choice of harmonization method to date has been determined via expert
-solicitation and has generally been applied to all trajectories for a given
-class of emissions species.
+in the IAM and climate change communities. For example,
+\cite{meinshausen_rcp_2011} describes a harmonization process using scaling
+routines for the 5 SRES regions with no sectoral definition. Further,
+\cite{rogelj_discrepancies_2011} describes the impacts of choosing various
+harmonization routines on future trajectories. Importantly, the choice of
+harmonization method to date has been determined via expert solicitation and has
+generally been applied to all trajectories for a given class of emissions
+species.
 

### 2017-05-23 intro.tex
adding citations, history

@@
 futures used in analyses underpinning the Intergovenmental Panel on Climate
-Change (IPCC) \TODO{cite} and the United Nations Framework Convention on Climate
-Unchange (UNFCCC)'s Paris Agreement \TODO{cite}.
+Change (IPCC) and the United Nations Framework Convention on Climate Unchange
+(UNFCCC)'s Paris Agreement \TODO{cite}.
 
@@
 historical emissions dataset for both anthropogenic and land-use change
-emissions \TODO{cite ceds and LUC paper} as part of the 6th phase of the Coupled
-Model Intercomparison Project \TODO{cite cmip6} (CMIP6). These historical
-emissions trajectories in conjunction with future trajectories as provided by
-IAMs are critical inputs to the climate-related modeling exercises of CMIP6.
+emissions \cite{hoesly_historical_2017,van_marle_historic_2017} as part of the
+6th phase of the Coupled Model Intercomparison Project \TODO{cite cmip6}
+(CMIP6). These historical emissions trajectories in conjunction with future
+trajectories as provided by IAMs are critical inputs to the climate-related
+modeling exercises of CMIP6.
 

### 2017-05-24 intro.tex
intro edited

--- 2017-05-23 intro.tex
@@
 Integrated Assessment Models (IAMs) are tools used to understand the complex
-interactions between energy systems, economoic systems, land use, the climate,
-and air pollution, among other sectors \TODO{cite some IAMs}. IAMs provide
-global projections of systemic change by dividing the world into a number of
-representative regions, the definition of which is distinct for each
-model. Results from IAMs are integral in a number of international studies,
-which notably include projections of climate and economic futures used in
-analyses underpinning the Intergovenmental Panel on Climate Change (IPCC)
-\TODO{cite} and the United Nations Framework Convention on Climate Unchange
-(UNFCCC)'s Paris Agreement \TODO{cite}.
+interactions between energy systems, economic systems, land use, the climate,
+and air pollution, among other sectors . IAMs provide global projections of
+systemic change by dividing the world into a number of representative regions,
+the definition of which is distinct for each model. Results from IAMs are
+integral in a number of international studies, which notably include projections
+of climate and economic futures. Recently, flagship IAMs have developed
+scenarios based on the Shared Socioeconomic Pathways (SSPs)
+\cite{van_vuuren_energy_2017, kriegler_fossil-fueled_2017, calvin_ssp4:_2017,
+  fujimori_ssp3:_2017, fricko_marker_2017} which model a variety of potential
+global futures. The SSPs are used in analyses underpinning the Intergovernmental
+Panel on Climate Change (IPCC) and the United Nations Framework Convention on
+Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.
 
@@
 (Good Health and Well-Being) and Goal 13 (Climate Action). IAMs endogenously
-calculate many of the emissions species of import that impact the progress
-towards these SDGs by modeling the individual technologies and sectors that
-contribute towards the emissions of different air pollutants and GHGs as well as
-various mitigation technologies (e.g., NOx and SOx scrubbers on power plants)
-that are used due to regulation or implemented policies.
+calculate the trajectories many of the emissions species of import that impact
+the progress towards these SDGs by modeling the individual technologies and
+sectors that contribute towards the emissions of different air pollutants and
+GHGs as well as various mitigation technologies (e.g., NOx and SOx scrubbers on
+power plants).
 
 In practice, IAMs calculate the total source intensity of emitting technologies,
-for example the total activity of coal power plant emissions in China, and
-incorporate emissions-intensity factors for individual species, for example the
-quantity of sulfur emissions from coal plants per megawatt-hour of production,
-in order to determine total emissions of a given gas species from a given
+for example the total activity of coal power plants in China, and incorporate
+emissions-intensity factors for individual gas species, for example the quantity
+of sulfur emissions from coal plants per megawatt-hour of production, in order
+to determine total emissions of a given gas species from a given
 technology. Models are then \textit{calibrated} to historical data sources which
@@
 assumptions of source activity, new emissions accounting practices, or other
-concerns. The global climate change community has recently updated a standard
-historical emissions dataset for both anthropogenic and land-use change
-emissions \TODO{cite ceds and LUC paper} as part of the 6th phase of the Coupled
-Model Intercomparison Project \TODO{cite cmip6} (CMIP6). These historical
-emissions trajectories in conjunction with future trajectories as provided by
-IAMs are critical inputs to the climate-related modeling exercises of CMIP6.
+concerns. The global climate change community has recently developed a standard
+historical emissions dataset for both anthropogenic
+\cite{hoesly_historical_2017} and land-use change emissions
+\cite{van_marle_historic_2017} as part of the 6th phase of the Coupled Model
+Inter-comparison Project (CMIP6). These historical emissions trajectories in
+conjunction with future trajectories as provided by IAMs are critical inputs to
+the climate-related modeling exercises of CMIP6.
 
@@
 incorporate updated historical datasets through the process of
-\textit{harmonization} in order to guarantee consistency with new historical
-datasets. Harmonization describes the process of updating model results to
-incorporate new historical timeseries while maintaining critical aspects of
-modeled future trajectories for variables of import. In the emissions context,
-this means that each individual combination of model region, model sector, and
-emissions species must be harmonized. Depending on the model, this can require
-the selection of thousands to tens-of-thousands of harmonization methods.
+\textit{harmonization} in order to guarantee consistency with these new
+historical datasets. Harmonization describes the process of updating model
+results to incorporate new historical time series while maintaining critical
+aspects of modeled future trajectories impacted outputs. In the emissions
+context, this means that each individual combination of model region, model
+sector, and emissions species must be harmonized. Depending on the total number
+of model regions, sectors, and emissions species, this can require the selection
+of thousands to tens-of-thousands of harmonization methods.
 
 Harmonization has been addressed in previous studies as it is a common practice
-in the IAM and climate change communities. For example, \TODO{cite Malte}
-describes a harmonization process using scaling routines for the 5 SRES regions
-with no sectoral definition. Further, \TODO{cite Joeri} describes the impacts of
+in the IAM and climate change communities. For example,
+\cite{meinshausen_rcp_2011} describes the use of scaling routines for the 5
+regions used in the Special Report on Emissions Scenarios \TODO{CITE}; however,
+only total emissions were harmonized in the exercise, thus there is no sectoral
+dimension. Further, \cite{rogelj_discrepancies_2011} describes the impacts of
 choosing various harmonization routines on future trajectories. Importantly, the
@@
 
-In order to incoporate the variety of possible model results and given
+In order to incorporate the large variety of possible model results and given
 relatively large increases in regional and sectoral dimensionality, an automated
 process for determining harmonization methods is needed. The remainder of this
-paper describes the methodology of the \code{aneris} codebase which implements
-such a system in Section \ref{sec:meths}. The results of applying the automated
-harmonization mechanism on two example IAM scenarios, one with emissions growth
-and another with emissions mitigation, is presented in Section
-\ref{sec:results}. Finally, the general effectiveness and potential future
-improvements on the automated methodology is discussed in Section
-\ref{sec:future}.
+paper describes the methodology of the \code{aneris} code base, written in the
+Python programming language, which implements such a system in Section
+\ref{sec:meths}. The results of applying the automated harmonization mechanism
+on two example IAM scenarios, one with emissions growth and another with
+emissions mitigation, is presented in Section \ref{sec:results}. Finally, the
+general effectiveness and potential future improvements on the automated
+methodology is discussed in Section \ref{sec:future}.
 

### 2017-05-24 intro.tex
revised meths section

@@
 concerns. The global climate change community has recently developed a standard
-historical emissions dataset for both anthropogenic
-\cite{hoesly_historical_2017} and land-use change emissions
-\cite{van_marle_historic_2017} as part of the 6th phase of the Coupled Model
-Inter-comparison Project (CMIP6). These historical emissions trajectories in
-conjunction with future trajectories as provided by IAMs are critical inputs to
-the climate-related modeling exercises of CMIP6.
+historical emissions dataset for both anthropogenic emissions (i.e., the
+Community Emissions Datasystem (CEDS)) \cite{hoesly_historical_2017} and
+land-use change (LUC) emissions \cite{van_marle_historic_2017} as part of the
+6th phase of the Coupled Model Inter-comparison Project (CMIP6). These
+historical emissions trajectories in conjunction with future trajectories as
+provided by IAMs are critical inputs to the climate-related modeling exercises
+of CMIP6.
 
@@
 \cite{meinshausen_rcp_2011} describes the use of scaling routines for the 5
-regions used in the Special Report on Emissions Scenarios \TODO{CITE}; however,
-only total emissions were harmonized in the exercise, thus there is no sectoral
-dimension. Further, \cite{rogelj_discrepancies_2011} describes the impacts of
-choosing various harmonization routines on future trajectories. Importantly, the
-choice of harmonization method to date has been determined via expert
-solicitation and has generally been applied to all trajectories for a given
-class of emissions species.
+regions used in the Special Report on Emissions Scenarios (SRES)
+\cite{nakicenovic2000}; however, only total emissions were harmonized in the
+exercise, thus there is no sectoral dimension. Further,
+\cite{rogelj_discrepancies_2011} describes the impacts of choosing various
+harmonization routines on future trajectories. Importantly, the choice of
+harmonization method to date has been determined via expert solicitation and has
+generally been applied to all trajectories for a given class of emissions
+species.
 

### 2017-05-24 meth.tex
intro added
Update on Overleaf.
revised meths section

@@
-\section{Methodology \& Implementation}
+\section{Methodology \& Implementation}\label{sec:meths}
 
@@
 including, temporal (normally half decade or decade), spatial (i.e., model
-regions), gas species, and sectoral. Each individual temporal trajectory (i.e.,
-unique combinations of regions, species, and sectors) must be harmonized to the
-initial time period. Given a model trajectory, $m_{r, g, s}(t)$, historical
-trajectory, $h_{r, g, s}(t)$, and initial time period $t_i$, a harmonized
-trajectory can be calculated.
+regions), gas species, and sectoral dimensions. Each individual temporal
+trajectory, i.e., unique combinations of regions ($r$), species ($g$), and
+sectors ($s$), must be harmonized to the initial time step. Given a model
+trajectory, $m_{r, g, s}(t)$, historical trajectory, $h_{r, g, s}(t)$, and
+initial time step $t_i$, a harmonized trajectory can be calculated.
 
-In previous studies, two \textit{families} of methods have been used: those that
-operate on the ratio of base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and
-those that operate on the offset of base year values (i.e., $h(t_i) -
-m(t_i)$). The \code{aneris} code base provides a number of variations on the
-classic methods, including ratio-convergence methods shown in Equation
-\ref{eqs:ratio}, offset-convergence methods shown in Equation \ref{eqs:offset},
-and interpolation methods shown in Equation \ref{eqs:interpolate}; the
-convergence factor for associated methods is shown first in Equation
-\ref{eqs:factor}. Each is a function of time, model trajectory, historical
-trajectory, harmonization time period, and a convergence time period ($t_f$),
-where the harmonized trajectory converges to the unharmonized trajectory.
+In previous studies \cite{meinshausen_rcp_2011,rogelj_discrepancies_2011}, two
+\textit{families} of methods have been used: those that operate on the ratio of
+base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and those that operate on the
+offset of base year values (i.e., $h(t_i) - m(t_i)$). The \code{aneris} code
+base provides a number of variations on the classic methods, including
+ratio-convergence methods shown in Equation \ref{eqs:ratio}, offset-convergence
+methods shown in Equation \ref{eqs:offset}, and interpolation methods shown in
+Equation \ref{eqs:interpolate}; the convergence factor, $\beta$, for associated
+methods is shown first in Equation \ref{eqs:factor}. Note that model and
+regional indices have been dropped for clarity. Each is a function of time,
+model trajectory, historical trajectory, harmonization time step, and a
+convergence time step ($t_f$), at which point the harmonized trajectory
+converges to the unharmonized trajectory.
 
@@
 
-The \code{aneris} code base provides a family of methods to choose from for each
-of the harmonization cases. A summary of all available methods is provided below
-in Table \ref{tab:meths}.
+The \code{aneris} code base provides a number of methods to choose from for each
+of the harmonization families. A summary of all available methods is provided
+below in Table \ref{tab:meths}.
 
@@
 \hline
-Method Name                             & Harmonization Type & Convergence Year\\
+Method Name                             & Harmonization Family & Convergence Year\\
 \hline
@@
 
-While historically expert opinion has guided the determination of harmonization
-methods to choose for a given trajectory, a new, \textit{decision tree} approach
-has been implemented in this work. In order to provide reasonable
+While historically expert opinion has guided the choice of harmonization methods
+for a given trajectory, a new, \textit{decision tree} approach has been
+implemented in this work which provides a systematic and documented
+decision-making process for method determination. In order to provide reasonable
 \textit{default} methods, the historical trajectory, unharmonized model
 trajectory, and relative difference between history and model values in the
-harmonization year are analyzed. A figure of the decision tree used in this
-analysis is shown in Figure \ref{fig:decision_tree}.
+harmonization year are analyzed. The decision tree used in this analysis is
+shown graphically in Figure \ref{fig:decision_tree}.
 
@@
       The default method decision tree used in the \code{aneris} software
-      library. Decision-making threshold are described below.}
+      library. For all decisions, upper branches represent a ``yes'' response
+      and lower branches represent a ``no'' response. Decision-making thresholds
+      are described below.}
   \end{center}
@@
 harmonization year. When model and historical values are relatively close, a
-convergence method is provided in order to be as representative as possible to
-the underlying unharmonized model results. The final model-based characteristic
-of import is whether the model reports negative emissions. Such a case is
-possible for gas species which can be extracted from the environment, as is the
-case for CO2 in future scenarios with high mitigation strategies. If a model
+convergence method is chosen in order to be as representative as possible to the
+underlying unharmonized model results. The final model-based characteristic of
+import is whether the model reports negative emissions. Such a case is possible
+for gas species which can be extracted from the environment and stored, as is
+the case for CO2 in future scenarios with high climate mitigation. If a model
 provides a trajectory that transitions from positive to negative emissions, then
@@
 and model in the harmonization year, for a negative trajectory to be
-inappropriately harmonized to positive, but decreasing, trajectory.
+inappropriately harmonized to a positive, but decreasing, trajectory.
 
@@
 modeled with varying fidelity by IAMs, but in general proxies are used for
-emissions trajectories. Take for example the emissions of black carbon (BC) from
-grassland burning in Africa \TODO{check that this is a good sector}: for most
-models, the use of grass and paturelands is modeled, and the emissions of
-species from the burning of these lands is determined by applying a emissions
-factor to a proxy model of burning of these lands; in general, the more
-grassland usage occurs, the higher the emissions from burning of
-grasslands. Therefore, consistency in harmonization method is important due to
-the approaches taken in modeling these sectors by IAMs. In order to reliably
-detect land-use-like emissions, a measure of the co-variance of the historical
-trajectory is calculated as shown in Equation \ref{eqs:cov} and tested against a
-threshold. To determine this threshold, an analysis of the recent CEDS
-\TODO{cite ceds paper} historical data has been performed. Figure \ref{fig:cov}
-shows the distribution of land-use co-variances and non-land-use co-variances as
-determined for historical data aggregated to the model regions of 5 different
-flagship IAMs \TODO{cite all 5 IAMs}. A threshold value of 20 has been chosen
-based on these observations. Importantly, tails of the LUC and non-LUC overlap,
-thus there are both false positives (~10\% of non-LUC trajectories) and false
-negatives (~7\% of LUC trajectories). However, as any regional definition is
-model dependent and thus any regional aggregation is possible an automated
-detection mechanism is necessary.
+emissions trajectories. Take for example emissions from grassland burning: for
+most models, the use of grass and paturelands is endogenously modeled, and
+emissions from the burning of these lands is determined by applying a emissions
+factor; in general, the more grassland usage that occurs in a scenario, the
+higher the emissions from the burning thereof. Therefore, consistency in
+harmonization method is important because the effects are modeled similarly
+across regions and species. In order to reliably detect land-use-like emissions,
+a measure of the coefficient of variation of the first derivative of the
+historical trajectory is calculated as shown in Equation \ref{eqs:cov}, where
+$\sigma$ is the standard deviation, $\mu$ is the mean, and $c_v$ is the
+coefficient of variation.
+
+The value of $c_v$ is then tested against a threshold. To determine this
+threshold, an analysis of the recent CEDS and LUC historical data has been
+performed. Figure \ref{fig:cov} shows the distribution of LUC $c_v$s and non-LUC
+$c_v$s as determined for historical data aggregated to the model regions of 5
+different flagship IAMs. A threshold value of 20 has been chosen based on these
+observations as it optimally divides the two distributions. Importantly, tails
+of the LUC and non-LUC overlap, thus there are both false positives (~10\% of
+non-LUC trajectories) and false negatives (~7\% of LUC trajectories). However,
+as any regional definition is model dependent and thus any regional aggregation
+is possible an automated detection mechanism is necessary.
 
 \begin{equation}\label{eqs:cov}
-    \Omega =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
+    c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
 \end{equation}
@@
       \label{fig:cov}
-      The distribution of $\Omega$ values for LUC and non-LUC historical
-      trajectories is shown. CEDS historical data is used for non-LUC data and
-      \TODO{cite LUC data} is used for LUC data. All historical data has been
-      aggregated from countries to IAM model regional definitions, and all gas
-      species included in the historical datasets are included in the
-      analysis. The solid black line indicates the threshold value used by
-      default in \code{aneris}.  
-    }
+      The distribution of $c_v$ values for LUC and non-LUC historical
+      trajectories is shown. CEDS historical data \cite{hoesly_historical_2017}
+      is used for non-LUC data and \cite{van_marle_historic_2017} is used for
+      LUC data. All historical data has been aggregated from countries to IAM
+      model regional definitions, and all gas species included in the historical
+      datasets are included in the analysis. The solid black line indicates the
+      threshold value used by default in \code{aneris}.  }
   \end{center}
@@
 Finally, consideration is taken with respect to the relative difference between
-the historic and model values in the harmonization year. In order to investigate
-the possible values that these relative differences can take, the flagship model
-values used in the SSP \TODO{CITE} and (ongoing) CMIP6 inter-comparison exercises
-are used. A distribution of these differences for all models in the study is
-presented in Figure \ref{fig:dh}. Given the available data, a threshold value of
-50\% was chosen to be used as a default in \code{aneris}.
+the historic and model values in the harmonization time period. In order to
+investigate the possible values that these relative differences can take, the
+flagship IAMs values used in the SSP and (ongoing) CMIP6 inter-comparison
+exercises are used. A distribution of these differences for all models in the
+study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
+value of 50\% was chosen to be used as a default in \code{aneris}.
 
@@
 The full \code{aneris} workflow is comprised of a number of components shown
-graphically in Figure \ref{fig:workflow}. Unharmonized model data and
+graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
 run-control configuration are read in via an Excel spreadsheet. Data is assumed
@@
 \code{Scenario}, \code{Region}, \code{Variable}, and \code{Unit} columns in
-addition to column representing each modeled time period. Users are able to
-control the harmonization process via a number of controls. First, users can
-provide \textit{override} methods for any combination of region and variable. If
-override methods are provided, they are used instead of the default methods as
-determined by the default method decision tree. Further, users can set the
-above-mentioned thresholds as well as the LUC method used in the decision
-tree. Further detail of input parameters can be found at
+addition to columns representing each modeled time period. Users are able to
+control the harmonization process via a number of options. First, users can
+provide \textit{override} methods for any combination of region and variable
+(i.e., sector and gas species). If override methods are provided, they are used
+instead of the default methods as determined by the default method decision
+tree. Furthermore, users can set the above-mentioned thresholds as well as the
+LUC method used in the decision tree (see Figure
+\ref{fig:fig:decision_tree}). Further detail of input parameters can be found at
 \hyperlink{http://aneris.readthedocs.io/en/latest/}.
@@
       that can be configured with user-based input configurations are shown in
-      purple. The core harmonization process is shown in orange.  }
+      purple. The core harmonization process is shown in yellow.  }
   \end{center}
@@
 
-Input data then undergoes a cleaning operation, which adds any trajectories in
-the historical dataset but not provided by the model and detects any issues that
-would cause the harmonization process to fail. The methods used to harmonize the
-data are then determined and the harmonization process is executed. Upon
-completion of the harmonization process, aggregation of common analysis regions
-is performed (e.g., the 5 RCP regions \TODO{CITE}). Finally, any exogenous
-trajectories the user provides are added. Exogenous trajectories are normally
-provided for unmodeled gases with well-accepted scenario trajectories, e.g.,
-Chlorofloro Carbons provided by WMO \TODO{CITE}. Upon completion, the harmonized
-trajectories and metadata regarding the harmonization process are returned. A
-description of all returned metadata is provided in Table \ref{tab:metadata}.
+Input data then undergoes a cleaning operation, which adds (null) model
+trajectories that exist in the historical dataset but are not provided by the
+model input and detects any issues that would cause the harmonization process to
+fail. The methods used to harmonize the data are then determined and the
+harmonization process is executed. Upon completion of the harmonization process,
+aggregation of common analysis regions is performed (e.g., the 5 SRES
+regions). Finally, any exogenous trajectories the user provides are
+added. Exogenous trajectories are normally provided for unmodeled gases with
+well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
+\TODO{CITE}. Upon completion, the harmonized trajectories and meta data
+regarding the harmonization process are returned. A description of all returned
+meta data is provided in Table \ref{tab:metadata}.
 
@@
 \centering
-\caption{Metadata provided by the \code{aneris} harmonization routine. This metadata is provided for every combination of region, sector, and emissions species.}
+\caption{Meta data provided by the \code{aneris} harmonization routine. This meta data is provided for every combination of region, sector, and emissions species.}
 \label{tab:metadata}
@@
 \hline
-cov          & The covariance value of the historical trajectory.                           \\
+cov          & The coefficient of variation value of the historical trajectory.                           \\
 \hline

### 2017-05-24 intro.tex
results section revised

@@
 scenarios based on the Shared Socioeconomic Pathways (SSPs)
-\cite{van_vuuren_energy_2017, kriegler_fossil-fueled_2017, calvin_ssp4:_2017,
-  fujimori_ssp3:_2017, fricko_marker_2017} which model a variety of potential
-global futures. The SSPs are used in analyses underpinning the Intergovernmental
-Panel on Climate Change (IPCC) and the United Nations Framework Convention on
-Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.
+\cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
+  calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which model a variety of
+potential global futures. The SSPs are used in analyses underpinning the
+Intergovernmental Panel on Climate Change (IPCC) and the United Nations
+Framework Convention on Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.
 
@@
 process for determining harmonization methods is needed. The remainder of this
-paper describes the methodology of the \code{aneris} code base, written in the
-Python programming language, which implements such a system in Section
-\ref{sec:meths}. The results of applying the automated harmonization mechanism
-on two example IAM scenarios, one with emissions growth and another with
-emissions mitigation, is presented in Section \ref{sec:results}. Finally, the
-general effectiveness and potential future improvements on the automated
+paper describes the methodology of the \code{aneris} code base \TODO{cite},
+written in the Python programming language, which implements such a system in
+Section \ref{sec:meths}. The results of applying the automated harmonization
+mechanism on two example IAM scenarios, one with emissions growth and another
+with emissions mitigation, is presented in Section \ref{sec:results}. Finally,
+the general effectiveness and potential future improvements on the automated
 methodology is discussed in Section \ref{sec:future}.

### 2017-05-24 results.tex
intro added
rough draft of results section
results section revised

@@
-\section{Results \& Conclusions}
-
-\subsection{A Harmonization Test Case}
+\section{Results}\label{sec:results}
 
 In order to test the \code{aneris} harmonization procedure, results from the IAM
-MESSAGE-GLOBIOM \TODO{CITE} was processed. Two scenarios from the Shared Socioeconomic
-Pathways scenario library \TODO{CITE} are investigated. The reference SSP2, or
-``middle of the road'' is chosen first because MESSAGE-GLOBIOM is the marker
-scenario for this SSP. Notably, emissions from multiple sectors tend to increase
-in SSP2, thus testing the harmonization mechanism with a increasing emission
-pathway. Additionally, the SSP-45 scenario is chosen, where the ``45''
-designation refers to a scenario with end-of-century emissions resulting in
-approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$ warming effect by the end of
-the century, due to both warming from Greenhouse Gases (GHGs) and fluoridated
-gases as well as cooling from aerosols. In this scenario, mitigation
-technologies and policies are enacted causing a general reduction in pollutants
-and GHGs, including (eventual) negative CO2 emissions in some regions and
-sectors due to carbon capture and sequestration and afforestation. This scenario
-explores the harmonization mechanism's response to generally decreasing emission
+MESSAGE-GLOBIOM \cite{fricko_marker_2017} was processed. Two scenarios from the
+SSP scenario library \cite{Riahi2017153,Rao2017346} are investigated. The SSP2,
+or ``middle of the road'', scenario family is chosen to be analyzed because
+MESSAGE-GLOBIOM is the marker scenario for this SSP. Of all SSP2 scenarios, the
+reference (marker) and SSP2-45 scenarios are analyzed. Notably, emissions from
+multiple sectors tend to increase in the reference scenario, thus testing the
+harmonization mechanism with an increasing emissions pathway. The SSP-45
+scenario, where the ``45'' designation refers to a scenario with end-of-century
+emissions resulting in approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$
+radiative forcing due to both warming from GHGs and fluoridated gases as well as
+cooling from aerosols, is chosen because mitigation technologies and policies
+are enacted causing a general reduction in pollutants and GHGs, including
+(eventual) negative CO2 emissions in some regions and sectors due to carbon
+capture and sequestration and afforestation. This scenario explores the
+harmonization mechanism's response to generally decreasing emission
 trajectories.
@@
 MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
-mapped directly to the 5-region definition used in the RCPs. Historical data is
-taken from CEDS and LUC \TODO{get correct name}, which comprise 10 separate
-pollutant and GHG species and 13 sectors. Therefore, a total of 1430 distinct
-trajectories were harmonized.
+mapped directly to the 5-region definition used in the SRES. Historical data is
+taken from previously described LUC and anthropogenic sources, which comprise 10
+separate pollutant and GHG species and 13 sectors. Therefore, a total of 1430
+distinct trajectories were harmonized for each scenario.
 
 The effect of harmonization overrides is analyzed by harmonizing each scenario
-first without overrides and then with overrides. Of the 1430 trajectories,
-approximately 2\% required the use of harmonization overrides after an initial
-investigation; thus, 98\% of all trajectories were satisfactorily determined
-using the default methods as chosen by the harmonization algorithm. Figure
-\ref{fig:nox} provides one such example of a emissions species and sector in
-which all regions were satisfactorily harmonized with the default methods. As
-can be seen, results with and without overrides are identical.
+first without and then with overrides. Of the 1430 trajectories, approximately
+2\% required the use of harmonization overrides after an initial investigation;
+thus, 98\% of all trajectories were satisfactorily determined using the default
+methods as chosen by the harmonization algorithm. Figure \ref{fig:nox} provides
+one such example of a emissions species and sector in which all regions were
+satisfactorily harmonized with the default methods. As can be seen, results with
+and without overrides are identical.
 
@@
       trajectories for SSP2 and SSP2-45 are presented. SSP-45 is denoted with
-      markers. Panel \textbf{a} shows 5-Region trajectories for scenarios
+      ``x'' markers. Panel \textbf{a} shows 5-Region trajectories for scenarios
       \textit{without} any overrides. Panel \textbf{b} shows 5-Region
@@
 represent an updated historical emissions trajectory while also maintaining
-consistency with the original, unharmonized pathway. When the default methods as
-provided by the harmonization procedure distort or otherwise sufficiently
-misrepresent the underlying unharmonized results, an override method is required
-to be provided for the trajectory of the region, sector, and species in
-question. Based on the 2\% of total trajectories required to be overridden, two
-classifications were observed: regional trajectories whose \textit{magnitude}
-was overly distorted and regional trajectories whose \textit{shape} was overly
-distorted.
+consistency with the original, unharmonized pathway. As such, when the default
+methods as provided by the harmonization procedure distort or otherwise
+sufficiently misrepresent the underlying unharmonized results, an override
+method is required to be provided for the trajectory of the region, sector, and
+species in question. Based on the 2\% of total trajectories required to be
+overridden, two classifications were observed: regional trajectories whose
+\textit{magnitude} was overly distorted and regional trajectories whose
+\textit{shape} was overly distorted.
 
-Figure \ref{fig:co} presents a situation in which the magnitude of a trajectory
-is distorted. A large discrepancy (~300\% relative difference) is observed in
-the harmonization year for carbon monoxide (CO) emissions in the industrial
-sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region, which
-comprises most of the Asian subcontinent. The default method chosen
-(\code{constant_ratio} maintains model trends for the region, overall model
-results are distorted. By applying a \code{constant_offset} override, the
-regional trend and shape is maintained. With the new harmonization method for
-the SAS region, the global trajectory for industrial CO also corresponds more
-closely with the unharmonized trajectory.
+Figure \ref{fig:co} presents a case in which the magnitude of a trajectory is
+distorted. A large discrepancy ($\sim$300\% relative difference) is
+observed in the harmonization year for carbon monoxide (CO) emissions in the
+industrial sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region,
+which comprises most of the Asian subcontinent. The default method chosen
+(\code{constant_ratio}) maintains model trends for the region; however, overall
+model results are distorted. By applying a \code{constant_offset} override, the
+regional trend and magnitude is maintained. With the new harmonization method
+for the SAS region, the global trajectory for industrial CO also corresponds
+more closely with the unharmonized trajectory.
 
@@
 can affect not only the magnitude but also the shape of regional
-trajectories. Figure \ref{fig:nh3} shows emissions trajectories for ammonia
-(NH3) resulting from the agriculture sector in Asia. Again, the SAS region shows
-a large discrepancy in the harmonization year (> 150\% in this case). The
-resulting trajectory harmonized with a \code{constant_ratio} method by default
-provides a large increase after 2080 in the SSP2 reference scenario. Notably,
-the SSP2-45 scenario is not affected to the same degree. While this distortion
-affects the magnitude of the SAS trajectory, it largely affects the post-2080
-shape of the global trajectory (see Figure \ref{fig:nh3}, panel \textbf{c}). By
-using a \code{constant_offset} method as an override, this distortion is
-addressed and more accurately reflects unharmonized results both in the SAS
-region and global results for agricultural ammonia emissions.
+trajectories. Figure \ref{fig:nh3} shows an example case of emissions
+trajectories for ammonia (NH3) from the agriculture sector in Asia. Again, the
+SAS region shows a large discrepancy in the harmonization year ($>$150\% in
+this case). The resulting trajectory harmonized with the default method
+(\code{constant_ratio}) provides a large increase after 2080 in the SSP2
+reference scenario. Notably, the SSP2-45 scenario is not affected to the same
+degree. While this distortion changes the magnitude of the SAS trajectory, it
+largely affects the post-2080 shape of the global trajectory (see Figure
+\ref{fig:nh3}, panel \textbf{c}). By using a \code{constant_offset} method as an
+override, this distortion is addressed and more accurately reflects unharmonized
+results both in the SAS region and global results for agricultural ammonia
+emissions.
 
@@
       \label{fig:nh3}
-      NH3 Agriculture harmonized and unharmonized emissions are presented for
+      NH3 agricultural harmonized and unharmonized emissions are presented for
       SSP2 and SSP2-45 scenarios. Scenarios and panel layouts are identical to
@@
       a magnitude distortion, but also a shape distortion at the tail of the
-      trajectory. Override methods have been applied to correct the distortion.
-    }
+      trajectory. Additionally, global trajectories are greatly affected by the
+      harmonization method choice (there is $\sim$20 percent relative difference
+      between trajectories in the reference scenario in 2100). Override methods
+      have been applied to correct the distortion.  }
   \end{center}
 \end{figure}
-
-\subsection{Discussion \& Future Work}

### 2017-05-24 future.tex
discusison section revised

@@
 methods for determining the use of harmonization methods was provided in Section
-\ref{sec:meths}. The \code{aneris} codebase was able to satisfactorily
+\ref{sec:meths}. The \code{aneris} code base was able to satisfactorily
 harmonized over 98\% of the 2860 individual trajectories that were analyzed in
@@
 determining harmonization methods for each individual combination of model
-region, sector, and emissions species while still providing a justified reason
-for each automated choice of harmonization method based on both the historical
-and future emissions trajectories. Furthermore, the automated approach continues
-to scale well as models become more detailed in both the regional and sectoral
-dimensions. Finally, expert opinion is still allowed to trump the automated
-method as determined by the algorithm via method overrides; however, these cases
-are clearly documented via the metadata provided as an output of \code{aneris}
-and thus can be individually explained. This provides not only transparency but
-also scientific integrity in the choice of harmonization methods.
+region, sector, and emissions species while still providing a justifiable
+explanation for each automated choice of harmonization method based on both the
+historical and future emissions trajectories. Furthermore, the automated
+approach continues to scale well as models become more detailed in both the
+regional and sectoral dimensions. Finally, expert opinion is still allowed to
+trump the automated method as determined by the algorithm via method overrides;
+however, these cases are clearly documented via the meta data provided as an
+output of \code{aneris} and thus can be individually explained. This provides
+not only transparency but also scientific integrity in the choice of
+harmonization methods.
 
 There are a variety of avenues for future improvement of both the \code{aneris}
-codebase and underlying methdology. As with any software project, additional
+code base and underlying methodology. As with any software project, additional
 users will provide use cases for more robust handling of input/output concerns
@@
 region. Perhaps the most fruitful investigation will involve further refinement
-of the defaul decision tree introduce in Section \ref{sec:meths}. A key aspect
+of the default decision tree introduced in Section \ref{sec:meths}. A key aspect
 missing from the decision tree is input from models regarding whether missing
 sources are the likely cause of a harmonization year discrepancy (suggesting the
-use of an offset method) or instead a discrepancy in emissions factors
-(suggesting the use of a ratio method)\TODO{cite Joeri}.
+use of an offset method) or instead a significant difference in emissions
+factors (suggesting the use of a ratio method) \cite{rogelj_discrepancies_2011}.
 
 However, this work provides a new direction and framework in which the IAM and
-climate communities can follow in order to reduce the necessicity of expert
+climate communities can follow in order to reduce the necessity of expert
 opinion and increase transparency and reproducibility of harmonization
@@
 base which can be used and improved upon by these communities. Both of these are
-clear steps in a positive direction for future climate and integrated assesment
+clear steps in a positive direction for future climate and integrated assessment
 modeling exercises.

### 2017-05-24 future.tex
discusison section revised v2

@@
 
-However, this work provides a new direction and framework in which the IAM and
-climate communities can follow in order to reduce the necessity of expert
+However, this work provides a new direction and framework which the IAM and
+climate communities can build upon in order to reduce the necessity of expert
 opinion and increase transparency and reproducibility of harmonization

### 2017-05-30 elsfrontmatter.tex
Update on Overleaf.
adding citations
Shinichiro_30May

@@
 Harmonization describes the process of calibrating Integrated Assessment Model
-(IAM) results with a new source of historical emissions trajectories. To date,
+(IAM) GHG and air pollutants results with a new source of historical emissions trajectories. To date,
 harmonization has been performed separately by individual modeling teams through
+% * <sfujimori112256@gmail.com> 2017-05-24T19:26:26.142Z:
+% 
+% > harmonization has been performed separately by individual modeling teams 
+% I think harmonization is not done. We just report the direct model outputs.
+% 
+% One of the ways to  make the background is that earlier IPCC scenarios (SRES and RCPs) did not document the harmonization method and far from transparent.
+% 
+% ^.
 the use of expert opinion and solicitation. As models and their results become
@@
 choosing harmonization methods becomes necessary. This work describes a novel
+% * <sfujimori112256@gmail.com> 2017-05-24T20:13:08.493Z:
+% 
+% > novel
+% It would be better to stress how novel and useful.
+% 
+% ^.
 methodology for determining such harmonization methods and an open-source Python
@@
 scenarios using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
-of the total emissions trajectories. \cite{Dirac1953888}
+of the total emissions trajectories.
 \end{abstract}

### 2017-05-30 intro.tex
Shinichiro_30May

--- 2017-05-24 intro.tex
@@
 and air pollution, among other sectors . IAMs provide global projections of
-systemic change by dividing the world into a number of representative regions,
+% * <sfujimori112256@gmail.com> 2017-05-24T20:16:51.179Z:
+% 
+% Here, Volker's paper seems well fit for the citation.
+% http://onlinelibrary.wiley.com/doi/10.1002/wene.98/abstract
+% 
+% 
+% ^.
+systemic change by dividing the world into a number of representative regions (typically 10 to 30),
 the definition of which is distinct for each model. Results from IAMs are
@@
 scenarios based on the Shared Socioeconomic Pathways (SSPs)
-\cite{van_vuuren_energy_2017, kriegler_fossil-fueled_2017, calvin_ssp4:_2017,
-  fujimori_ssp3:_2017, fricko_marker_2017} which model a variety of potential
-global futures. The SSPs are used in analyses underpinning the Intergovernmental
-Panel on Climate Change (IPCC) and the United Nations Framework Convention on
-Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.
+\cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
+  calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which quantify a variety of
+potential global futures.
+The SSPs are primarily designed to being used in analyses of broad climate research which includes earth system model (ESM) simulations, climate impact, adaptation climate mitigation studies.
+They consequently underpin the
+Intergovernmental Panel on Climate Change (IPCC) across working groups and the United Nations
+% * <sfujimori112256@gmail.com> 2017-05-24T20:32:22.989Z:
+% 
+% IPCC itself collects, reviews and summarizes papers. SSPs are officially independent activity from IPCC. So, I slightly changed the context. 
+% 
+% ^.
+% * <sfujimori112256@gmail.com> 2017-05-24T20:37:11.995Z:
+% 
+% I think Paris agreement is not so relevant with SSPs so I removed.
+% 
+% ^.
 
@@
 optimization, the core inputs, outputs, and purpose are similar across different
-models. Modeling teams incorporate existing datasets of energy systems, world
-development indicators, and emissions sources and concentrations, among other
+models. Modeling teams incorporate existing datasets of energy systems, land use, economic, demographic and emissions sources and concentrations, among other
 data, in order to provide consistent historical trajectories of modeled
@@
 Future emissions pathways are of critical importance to both of global and local
+% * <sfujimori112256@gmail.com> 2017-05-24T20:47:32.377Z:
+% 
+% > Future emissions pathways are of critical importance to both of global and local
+% > governance concerns. For example, the United Nation's Sustainable Development
+% > Goals (SDGs) \TODO{CITE} speak directly to the impacts of emissions in Goal 3
+% > (Good Health and Well-Being) and Goal 13 (Climate Action). IAMs endogenously
+% > calculate the trajectories many of the emissions species of import that impact
+% > the progress towards these SDGs by modeling the individual technologies and
+% > sectors that contribute towards the emissions of different air pollutants and
+% > GHGs as well as various mitigation technologies (e.g., NOx and SOx scrubbers on
+% > power plants).
+% I am not sure whether this paragraph is needed, because if we only discuss about SDGs, the emissions harmonization is not the necessity process for us essentially.
+% 
+% I would directly keen much more on the needs of harmonization and its importance. Then, the logic would be 
+% 1) SSPs will be an basis for CMIP6 (Brian's scenarioMIP GMD paper citation)
+% 2) CMIP6 uses different model outcomes (e.g. SSP1-26 IMAGE, SSP2-45 MESSAGE), which means we need somehow harmonized data set for at least base year.
+% 3) However, current IAMs emissions inventory differ a lot, which is hardly acceptable for CMIP experiments.
+% 4) 
+% 
+% 
+% 
+% ^.
 governance concerns. For example, the United Nation's Sustainable Development
@@
 In order to incorporate the large variety of possible model results and given
+% * <sfujimori112256@gmail.com> 2017-05-24T21:09:01.445Z:
+% 
+% I think it would be better to declare what metrics we are going to use to evaluate the methodology.
+In what case we can say this method succeeds or fails?
+That would be essentially the one of the highlight of the conclusion.
+% 
+% ^.
 relatively large increases in regional and sectoral dimensionality, an automated
 process for determining harmonization methods is needed. The remainder of this
-paper describes the methodology of the \code{aneris} code base, written in the
-Python programming language, which implements such a system in Section
-\ref{sec:meths}. The results of applying the automated harmonization mechanism
-on two example IAM scenarios, one with emissions growth and another with
-emissions mitigation, is presented in Section \ref{sec:results}. Finally, the
-general effectiveness and potential future improvements on the automated
+% * <sfujimori112256@gmail.com> 2017-05-24T21:12:03.999Z:
+% 
+% > regional and sectoral dimensionality, an automated
+% > process for determining harmonization methods is needed. The remainder of this
+% 
+% I suggest to add some explanations why detailed regional and sectoral harmonization is needed.
+As you mentioned, from RCPs sectoral and regional harmonization is needed.
+Meantime, ESM (or in this case we should say climate models) have been increasing the spatial resolution overtime and much more detailed representation become available, which essentially requires detailed information.
+The obvious example is AerChemMIP.
+% 
+% ^.
+paper describes the methodology of the \code{aneris} code base \TODO{cite},
+written in the Python programming language, which implements such a system in
+Section \ref{sec:meths}. The results of applying the automated harmonization
+mechanism on two example IAM scenarios, one with emissions growth and another
+with emissions mitigation, is presented in Section \ref{sec:results}. Finally,
+the general effectiveness and potential future improvements on the automated
 methodology is discussed in Section \ref{sec:future}.

### 2017-06-06 intro.tex
addresed shinichiro's comments

--- 2017-05-24 intro.tex
@@
 and air pollution, among other sectors . IAMs provide global projections of
-systemic change by dividing the world into a number of representative regions,
-the definition of which is distinct for each model. Results from IAMs are
-integral in a number of international studies, which notably include projections
-of climate and economic futures. Recently, flagship IAMs have developed
-scenarios based on the Shared Socioeconomic Pathways (SSPs)
-\cite{van_vuuren_energy_2017, kriegler_fossil-fueled_2017, calvin_ssp4:_2017,
-  fujimori_ssp3:_2017, fricko_marker_2017} which model a variety of potential
-global futures. The SSPs are used in analyses underpinning the Intergovernmental
-Panel on Climate Change (IPCC) and the United Nations Framework Convention on
-Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.
+systemic change by dividing the world into a number of representative regions
+(typically 10 to 30), the definition of which is distinct for each model
+\cite{krey_global_2014}. Results from IAMs are integral in a number of
+international studies, which notably include projections of climate and economic
+futures. Recently, flagship IAMs have developed scenarios based on the Shared
+Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
+  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
+quantify a variety of potential global futures. The SSPs are primarily designed
+to being used in analyses of broad climate research which includes earth system
+model (ESM) simulations, climate impact, adaptation climate mitigation
+studies. They consequently underpin the United Nations' Intergovernmental Panel
+on Climate Change (IPCC) across working groups.
 
@@
 optimization, the core inputs, outputs, and purpose are similar across different
-models. Modeling teams incorporate existing datasets of energy systems, world
-development indicators, and emissions sources and concentrations, among other
+models. Modeling teams incorporate existing data sets of energy systems, land
+use, economic, demographic and emissions sources and concentrations, among other
 data, in order to provide consistent historical trajectories of modeled
@@
 
-Future emissions pathways are of critical importance to both of global and local
-governance concerns. For example, the United Nation's Sustainable Development
-Goals (SDGs) \TODO{CITE} speak directly to the impacts of emissions in Goal 3
-(Good Health and Well-Being) and Goal 13 (Climate Action). IAMs endogenously
-calculate the trajectories many of the emissions species of import that impact
-the progress towards these SDGs by modeling the individual technologies and
-sectors that contribute towards the emissions of different air pollutants and
-GHGs as well as various mitigation technologies (e.g., NOx and SOx scrubbers on
-power plants).
+The emissions trajectories calculated by IAMs are critical inputs for ongoing,
+worldwide scientific community efforts in the Coupled Model Intercomparison
+Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
+of marker SSP scenarios developed by the IAM community
+\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
+modeling the individual technologies and sectors that contribute towards the
+emissions of different air pollutants and GHGs as well as various mitigation
+technologies (e.g., NOx and SOx scrubbers on power plants). However, the
+historical emissions inventories used by different models can differ by large
+amounts depending on the region, sector, and emissions species being modeled.
 
@@
 
-In order to incorporate the large variety of possible model results and given
-relatively large increases in regional and sectoral dimensionality, an automated
+In previous studies, e.g. the Representative Concentration Pathways (RCPs), IAM
+results have been harmonized by sector and the 5 RCP global regions
+\cite{vuuren_representative_2011}. Climate modeling efforts have continued to
+progress, however, demanding increased spatial and sectoral resolution. In order
+to address the growing dimensionality of model outputs while still providing a
+consistent and scientifically rigorous harmonization procedure, an automated
 process for determining harmonization methods is needed. The remainder of this
-paper describes the methodology of the \code{aneris} code base, written in the
-Python programming language, which implements such a system in Section
-\ref{sec:meths}. The results of applying the automated harmonization mechanism
-on two example IAM scenarios, one with emissions growth and another with
-emissions mitigation, is presented in Section \ref{sec:results}. Finally, the
-general effectiveness and potential future improvements on the automated
-methodology is discussed in Section \ref{sec:future}.
+paper describes the methodology of the \code{aneris} code base
+\cite{matthew_gidden_2017_802832}, written in the Python programming language,
+which implements such a system in Section \ref{sec:meths} (documentation
+available \href{http://mattgidden.com/aneris/}{online}). The results of applying
+the automated harmonization mechanism on two example IAM scenarios, one with
+emissions growth and another with emissions mitigation, is presented in Section
+\ref{sec:results}. Finally, the general effectiveness and potential future
+improvements on the automated methodology is discussed in Section
+\ref{sec:future}.
 

### 2017-06-06 meth.tex
more meth revisions
Shinichiro_30May
addresed shinichiro's comments

@@
 trajectory, $m_{r, g, s}(t)$, historical trajectory, $h_{r, g, s}(t)$, and
-initial time step $t_i$, a harmonized trajectory can be calculated.
+initial time step $t_i$, a harmonized trajectory can be calculated. The
+``goodness of fit'' of a trajectory depends on a number of factors, most notably
+the faithful representation of the original, unharmonized, trajectory. However,
+other important factors must be considered, such the representation of negative
+trajectories (i.e., if a trajectory becomes negative, both the timing and total
+magnitude should be as close as possible) which are of critical importance for
+cumulative CO2 calculations.
 
@@
 LUC method used in the decision tree (see Figure
-\ref{fig:fig:decision_tree}). Further detail of input parameters can be found at
-\hyperlink{http://aneris.readthedocs.io/en/latest/}.
+\ref{fig:decision_tree}). Further detail of input parameters can be found at
+\hyperlink{http://aneris.readthedocs.io/en/latest/}{http://aneris.readthedocs.io/en/latest/}.
 
@@
 well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
-\TODO{CITE}. Upon completion, the harmonized trajectories and meta data
+\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
 regarding the harmonization process are returned. A description of all returned

### 2017-06-06 intro.tex
tweak doc

@@
 paper describes the methodology of the \code{aneris} code base
-\cite{matthew_gidden_2017_802832}, written in the Python programming language,
-which implements such a system in Section \ref{sec:meths} (documentation
-available \href{http://mattgidden.com/aneris/}{online}). The results of applying
-the automated harmonization mechanism on two example IAM scenarios, one with
-emissions growth and another with emissions mitigation, is presented in Section
-\ref{sec:results}. Finally, the general effectiveness and potential future
-improvements on the automated methodology is discussed in Section
+\cite{matthew_gidden_2017_802832}, written in the Python programming language
+(detailed documentation available \href{http://mattgidden.com/aneris/}{online}),
+which implements such a system in Section \ref{sec:meths}. The results of
+applying the automated harmonization mechanism on two example IAM scenarios, one
+with emissions growth and another with emissions mitigation, is presented in
+Section \ref{sec:results}. Finally, the general effectiveness and potential
+future improvements on the automated methodology is discussed in Section
 \ref{sec:future}.

### 2017-06-06 future.tex
some final edits

--- 2017-05-24 future.tex
@@
 
-However, this work provides a new direction and framework in which the IAM and
-climate communities can follow in order to reduce the necessity of expert
-opinion and increase transparency and reproducibility of harmonization
+However, this work provides a new direction and framework which the IAM and
+climate communities can build upon in order to reduce the necessity of ad-hoc
+expert opinion and increase transparency and reproducibility of harmonization
 exercises. Furthermore, it provides an open-source, tested, and documented code

### 2017-06-06 intro.tex
some final edits

@@
 quantify a variety of potential global futures. The SSPs are primarily designed
-to being used in analyses of broad climate research which includes earth system
+to be used in analyses of broad climate research which include earth system
 model (ESM) simulations, climate impact, adaptation climate mitigation
@@
 models. Modeling teams incorporate existing data sets of energy systems, land
-use, economic, demographic and emissions sources and concentrations, among other
+use, economics, demographics and emissions sources and concentrations, among other
 data, in order to provide consistent historical trajectories of modeled
@@
 Community Emissions Datasystem (CEDS)) \cite{hoesly_historical_2017} and
-land-use change (LUC) emissions \cite{van_marle_historic_2017} as part of the
-6th phase of the Coupled Model Inter-comparison Project (CMIP6). These
-historical emissions trajectories in conjunction with future trajectories as
-provided by IAMs are critical inputs to the climate-related modeling exercises
-of CMIP6.
+land-use change (LUC) emissions \cite{van_marle_historic_2017} which, in
+conjunction with the SSP IAM trajectories, will be used for climate-related
+modeling exercises of CMIP6.
 
@@
 
-In previous studies, e.g. the Representative Concentration Pathways (RCPs), IAM
+In recent studies, i.e., the Representative Concentration Pathways (RCPs), IAM
 results have been harmonized by sector and the 5 RCP global regions
 \cite{vuuren_representative_2011}. Climate modeling efforts have continued to
-progress, however, demanding increased spatial and sectoral resolution. In order
-to address the growing dimensionality of model outputs while still providing a
-consistent and scientifically rigorous harmonization procedure, an automated
-process for determining harmonization methods is needed. The remainder of this
-paper describes the methodology of the \code{aneris} code base
-\cite{matthew_gidden_2017_802832}, written in the Python programming language
-(detailed documentation available \href{http://mattgidden.com/aneris/}{online}),
-which implements such a system in Section \ref{sec:meths}. The results of
-applying the automated harmonization mechanism on two example IAM scenarios, one
-with emissions growth and another with emissions mitigation, is presented in
-Section \ref{sec:results}. Finally, the general effectiveness and potential
-future improvements on the automated methodology is discussed in Section
-\ref{sec:future}.
+progress, however, demanding increased spatial and sectoral resolution from
+IAMs. In order to address the growing dimensionality of model outputs while
+still providing a consistent and scientifically rigorous harmonization
+procedure, an automated process for determining harmonization methods is
+needed. The remainder of this paper describes the methodology of the
+\code{aneris} code base \cite{matthew_gidden_2017_802832}, written in the Python
+programming language (detailed documentation available
+\href{http://mattgidden.com/aneris/}{online}), which implements such a system in
+Section \ref{sec:meths}. The results of applying the automated harmonization
+mechanism on two example IAM scenarios, one with emissions growth and another
+with emissions mitigation, is presented in Section \ref{sec:results}. Finally,
+the general effectiveness and potential future improvements on the automated
+methodology is discussed in Section \ref{sec:future}.
 

### 2017-06-06 meth.tex
some final edits

@@
 
-IAM emission results are provided along a number of different dimensions,
-including, temporal (normally half decade or decade), spatial (i.e., model
-regions), gas species, and sectoral dimensions. Each individual temporal
-trajectory, i.e., unique combinations of regions ($r$), species ($g$), and
-sectors ($s$), must be harmonized to the initial time step. Given a model
-trajectory, $m_{r, g, s}(t)$, historical trajectory, $h_{r, g, s}(t)$, and
-initial time step $t_i$, a harmonized trajectory can be calculated. The
-``goodness of fit'' of a trajectory depends on a number of factors, most notably
-the faithful representation of the original, unharmonized, trajectory. However,
-other important factors must be considered, such the representation of negative
-trajectories (i.e., if a trajectory becomes negative, both the timing and total
-magnitude should be as close as possible) which are of critical importance for
-cumulative CO2 calculations.
+IAM emission results are provided along temporal (normally half decade or
+decade), spatial (i.e., model regions), gas species, and sectoral
+dimensions. Each individual temporal trajectory, i.e., unique combinations of
+regions ($r$), species ($g$), and sectors ($s$), must be harmonized to the
+initial time step. Given a model trajectory, $m_{r, g, s}(t)$, historical
+trajectory, $h_{r, g, s}(t)$, and initial time step $t_i$, a harmonized
+trajectory can be calculated. The ``goodness of fit'' of a trajectory depends on
+a number of factors, most notably the faithful representation of the original,
+unharmonized, trajectory. However, other important factors must be considered,
+such as the representation of negative trajectories (i.e., if a trajectory
+becomes negative, both the timing and total magnitude should be as close as
+possible) which are of critical importance for cumulative CO2 calculations.
 
@@
 Equation \ref{eqs:interpolate}; the convergence factor, $\beta$, for associated
-methods is shown first in Equation \ref{eqs:factor}. Note that model and
-regional indices have been dropped for clarity. Each is a function of time,
-model trajectory, historical trajectory, harmonization time step, and a
+methods is shown first in Equation \ref{eqs:factor}. Note that region, species,
+and sector indices have been dropped for clarity. Each equation is a function of
+time, model trajectory, historical trajectory, harmonization time step, and a
 convergence time step ($t_f$), at which point the harmonized trajectory
-converges to the unharmonized trajectory.
+converges to the unharmonized trajectory. The \code{aneris} code base provides a
+number of methods to choose from for each of the harmonization families. A
+summary of all available methods is provided in Table \ref{tab:meths}.
 
@@
 \end{equation}
-
-The \code{aneris} code base provides a number of methods to choose from for each
-of the harmonization families. A summary of all available methods is provided
-below in Table \ref{tab:meths}.
 
@@
       library. For all decisions, upper branches represent a ``yes'' response
-      and lower branches represent a ``no'' response. Decision-making thresholds
-      are described below.}
+      and lower branches represent a ``no'' response. $dH$ is defined as
+      $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and decision-making
+      thresholds are described below. Finally, default choices in \textit{green}
+      label methods that will likely closely match unharmonized results, choices
+      in \textit{yellow} will likely somewhat match unharmonized results, and
+      choices in \textit{red} can be expected to have a large relative
+      difference between harmonized and unharmonized results.}
   \end{center}
@@
 most models, the use of grass and paturelands is endogenously modeled, and
-emissions from the burning of these lands is determined by applying a emissions
+emissions from the burning of these lands is determined by applying an emissions
 factor; in general, the more grassland usage that occurs in a scenario, the
@@
 a measure of the coefficient of variation of the first derivative of the
-historical trajectory is calculated as shown in Equation \ref{eqs:cov}, where
-$\sigma$ is the standard deviation, $\mu$ is the mean, and $c_v$ is the
-coefficient of variation.
+historical trajectory, $c_v$, is calculated as shown in Equation \ref{eqs:cov},
+where $\sigma$ is the standard deviation, and $\mu$ is the mean of the
+trajectory.
 
@@
 harmonization process is executed. Upon completion of the harmonization process,
-aggregation of common analysis regions is performed (e.g., the 5 SRES
+aggregation of common analysis regions is performed (e.g., the 5 RCP
 regions). Finally, any exogenous trajectories the user provides are

### 2017-06-06 results.tex
some final edits

@@
 multiple sectors tend to increase in the reference scenario, thus testing the
-harmonization mechanism with an increasing emissions pathway. The SSP-45
+harmonization mechanism with an increasing emissions pathway. The SSP2-45
 scenario, where the ``45'' designation refers to a scenario with end-of-century
@@
 MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
-mapped directly to the 5-region definition used in the SRES. Historical data is
+mapped directly to the 5-region definition used in the RCPs. Historical data is
 taken from previously described LUC and anthropogenic sources, which comprise 10

### 2017-06-07 elsfrontmatter.tex
Shinichiro_edition_7-June-2017
Update on Overleaf.

@@
 \author[iiasa,nies]{Shinichiro Fujimori}
+\author[iiasa]{Keywan Riahi}
 
@@
 code base for implementing the methodology. Results are shown for two example
-scenarios using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
+scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
 of the total emissions trajectories.

### 2017-06-07 intro.tex
Update on Overleaf.

@@
 to be used in analyses of broad climate research which include earth system
-model (ESM) simulations, climate impact, adaptation climate mitigation
+model (ESM) simulations, climate impact, adaptation and climate mitigation
 studies. They consequently underpin the United Nations' Intergovernmental Panel
@@
 historical emissions dataset for both anthropogenic emissions (i.e., the
-Community Emissions Datasystem (CEDS)) \cite{hoesly_historical_2017} and
-land-use change (LUC) emissions \cite{van_marle_historic_2017} which, in
+Community Emissions Datasystem (CEDS) \cite{hoesly_historical_2017} and
+land-use change (LUC) emissions \cite{van_marle_historic_2017}) which, in
+% * <sfujimori112256@gmail.com> 2017-06-07T09:59:35.820Z:
+% 
+% > land-use change (LUC) emissions \cite{van_marle_historic_2017}
+% I think this one also anthropogenic
+% 
+% ^.
 conjunction with the SSP IAM trajectories, will be used for climate-related

### 2017-06-07 meth.tex
Update on Overleaf.

@@
 for gas species which can be extracted from the environment and stored, as is
-the case for CO2 in future scenarios with high climate mitigation. If a model
+the case for CO2 in future scenarios with climate mitigation. If a model
+% * <sfujimori112256@gmail.com> 2017-06-07T10:07:21.265Z:
+% 
+% > with climate mitigation
+% Not necessarily high mitigation. Land use can be negative even in baseline.
+% 
+% ^.
 provides a trajectory that transitions from positive to negative emissions, then
@@
 (i.e., sector and gas species). If override methods are provided, they are used
+% * <sfujimori112256@gmail.com> 2017-06-07T10:12:45.308Z:
+% 
+% Maybe it would be better to add an explanation why we allow this option like
+% "We allow this overriding option because the default method do not meet our criteria sometimes (but usually very few cases)"
+% 
+% ^.
 instead of the default methods as determined by the default method decision
@@
 regions). Finally, any exogenous trajectories the user provides are
+% * <sfujimori112256@gmail.com> 2017-06-07T10:17:07.935Z:
+% 
+% > (e.g., the 5 RCP
+% > regions) 
+% It might be good to show what is 5 regions (or definition) in the SM.
+% 
+% ^.
 added. Exogenous trajectories are normally provided for unmodeled gases with

### 2017-06-09 meth.tex
addressing shinichiro's comments

@@
 
-\subsection{\code{aneris} Workflow}
+\subsection{\code{aneris} Workflow}\ref{label:workflow}
 
@@
 \code{Scenario}, \code{Region}, \code{Variable}, and \code{Unit} columns in
-addition to columns representing each modeled time period. Users are able to
-control the harmonization process via a number of options. First, users can
-provide \textit{override} methods for any combination of region and variable
-(i.e., sector and gas species). If override methods are provided, they are used
-% * <sfujimori112256@gmail.com> 2017-06-07T10:12:45.308Z:
-% 
-% Maybe it would be better to add an explanation why we allow this option like
-% "We allow this overriding option because the default method do not meet our criteria sometimes (but usually very few cases)"
-% 
-% ^.
-instead of the default methods as determined by the default method decision
-tree. Furthermore, users can set the above-mentioned thresholds as well as the
-LUC method used in the decision tree (see Figure
-\ref{fig:decision_tree}). Further detail of input parameters can be found at
-\hyperlink{http://aneris.readthedocs.io/en/latest/}{http://aneris.readthedocs.io/en/latest/}.
+addition to columns representing each modeled time period.
+
+Users are able to control the harmonization process via a number of options. The
+primary mechanism by which users control the process is by providing
+\textit{override} methods for any combination of region and variable (i.e.,
+sector and gas species).  In practice, it may be possible that not all default
+methods chosen will provide robust harmonized trajectories, especially if there
+is a significant difference between historical and model values in the
+harmonization year and there is significant upward or downward movement in the
+model trajectory. In such cases, override methods can ameliorate the issues
+associated with the default method choice. In order to help identify these
+cases, harmonization \textit{diagnostics} are provided which analyze the relative
+difference between harmonized and unharmonized trajectories at their mid and
+end-points. If override methods are provided, they are used instead of the
+default methods as determined by the default method decision tree. Furthermore,
+users can set the above-mentioned thresholds as well as the LUC method used in
+the decision tree (see Figure \ref{fig:decision_tree}). Further detail of input
+parameters can be found \href{http://mattgidden.com/aneris/config.html}{online}.
 
@@
 harmonization process is executed. Upon completion of the harmonization process,
-aggregation of common analysis regions is performed (e.g., the 5 RCP
-regions). Finally, any exogenous trajectories the user provides are
-% * <sfujimori112256@gmail.com> 2017-06-07T10:17:07.935Z:
-% 
-% > (e.g., the 5 RCP
-% > regions) 
-% It might be good to show what is 5 regions (or definition) in the SM.
-% 
-% ^.
-added. Exogenous trajectories are normally provided for unmodeled gases with
+aggregation of common analysis regions is performed. A common regional
+aggregation used in the IAM community is the RCP regional definition, as shown
+in Figure \ref{fig:regions}. Finally, any exogenous trajectories the user provides
+are added. Exogenous trajectories are normally provided for unmodeled gases with
 well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
@@
 meta data is provided in Table \ref{tab:metadata}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.85\textwidth]{regions.png}
+    \caption[]{
+      \label{fig:regions}
+      The 5 regions used in the RCPs with their 11-region constituents: Asia
+      (Central Asia, South Asia, Pacific Asia), Latin Ameria, the Middle East
+      and Africa, the OECD (North America, Western Europe, and Pacfic OECD),
+      Reforming Economies (Eastern Europe and Former Soviet Union).
+    }
+  \end{center}
+\end{figure}
 

### 2017-06-09 results.tex
Update on Overleaf.
adding better regional png and species/sector table

@@
 trajectories.
+% * <sfujimori112256@gmail.com> 2017-06-07T10:22:28.473Z:
+% 
+% As a supplementary, it would be better to show the total kyoto gases emissions for these two scenarios, which gives readers rough image what kind of scenarios we choose.
+% 
+% ^.
 
@@
 taken from previously described LUC and anthropogenic sources, which comprise 10
-separate pollutant and GHG species and 13 sectors. Therefore, a total of 1430
-distinct trajectories were harmonized for each scenario.
+separate pollutant and GHG species and 13 sectors shown in
+\ref{tab:sp}. Therefore, a total of 1430 distinct trajectories were harmonized
+for each scenario.
+
+\begin{table}[]
+  \centering
+  \caption{Harmonized Species and Sectors}
+  \label{tab:sp}
+  \begin{tabular}{|l|l|}
+    \hline
+    \textbf{Emissions Species}        & \textbf{Sectors}                    \\
+    \hline
+    Black Carbon (BC)                 & Agricultural Waste Burning          \\
+    Hexafluoroethane (C2F6)           & Agriculture                         \\
+    Tetrafluoromethane (CF4)          & Aircraft                            \\
+    Methane (CH4)                     & Energy Sector                       \\
+    Carbon Dioxide (CO2)              & Forest Burning                      \\
+    Carbon Monoxide (CO)              & Grassland Burning                   \\
+    Hydrofluorocarbons (HFCs)         & Industrial Sector                   \\
+    Nitrous Oxide (N2O)               & International Shipping              \\
+    Ammonia (NH3)                     & Residential Commercial Other        \\
+    Nitrogen Oxides (NOx)             & Solvents Production and Application \\
+    Organic Carbon (OC)               & Transportation Sector               \\
+    Sulfur Hexafluoride (SF6)         & Unharmonized                        \\
+    Sulfur Oxides (SOx)               & Waste                               \\
+    Volatile Organic Compounds (VOCs) &
+    \hline
+  \end{tabular}
+\end{table}
 
@@
 overridden, two classifications were observed: regional trajectories whose
+% * <sfujimori112256@gmail.com> 2017-06-07T10:28:41.326Z:
+% 
+% > regional trajectories whose
+% > \textit{magnitude} was overly distorted and regional trajectories whose
+% > \textit{shape} was overly distorted.
+% 
+% How can we detect these results?
+% If we could have a systematic detection criteria, then it would be really worthwhile to describe.
+% Otherwise, it means that the users of this tool generally  have to see all results by their eye, which could be essentially a weak point of this tool.
+% If so, we may want to discuss that point in the discussion section,
+% 
+% ^.
 \textit{magnitude} was overly distorted and regional trajectories whose

### 2017-06-09 meth.tex
more shinichiro edits

@@
 
-\subsection{\code{aneris} Workflow}\ref{label:workflow}
+\subsection{\code{aneris} Workflow}\label{sec:workflow}
 
@@
       The 5 regions used in the RCPs with their 11-region constituents: Asia
-      (Central Asia, South Asia, Pacific Asia), Latin Ameria, the Middle East
-      and Africa, the OECD (North America, Western Europe, and Pacfic OECD),
+      (Central Asia, South Asia, Pacific Asia), Latin America, the Middle East
+      and Africa, the OECD (North America, Western Europe, and Pacific OECD),
       Reforming Economies (Eastern Europe and Former Soviet Union).

### 2017-06-09 results.tex
more shinichiro edits

--- 2017-06-06 results.tex
@@
 taken from previously described LUC and anthropogenic sources, which comprise 10
-separate pollutant and GHG species and 13 sectors. Therefore, a total of 1430
-distinct trajectories were harmonized for each scenario.
+separate pollutant and GHG species and 13 sectors shown in
+\ref{tab:sp}. Therefore, a total of 1430 distinct trajectories were harmonized
+for each scenario.
+
+\begin{table}[]
+  \centering
+  \caption{Harmonized Species and Sectors}
+  \label{tab:sp}
+  \begin{tabular}{|l|l|}
+    \hline
+    \textbf{Emissions Species}        & \textbf{Sectors}                    \\
+    \hline
+    Black Carbon (BC)                 & Agricultural Waste Burning          \\
+    Hexafluoroethane (C2F6)           & Agriculture                         \\
+    Tetrafluoromethane (CF4)          & Aircraft                            \\
+    Methane (CH4)                     & Energy Sector                       \\
+    Carbon Dioxide (CO2)              & Forest Burning                      \\
+    Carbon Monoxide (CO)              & Grassland Burning                   \\
+    Hydrofluorocarbons (HFCs)         & Industrial Sector                   \\
+    Nitrous Oxide (N2O)               & International Shipping              \\
+    Ammonia (NH3)                     & Residential Commercial Other        \\
+    Nitrogen Oxides (NOx)             & Solvents Production and Application \\
+    Organic Carbon (OC)               & Transportation Sector               \\
+    Sulfur Hexafluoride (SF6)         & Unharmonized                        \\
+    Sulfur Oxides (SOx)               & Waste                               \\
+    Volatile Organic Compounds (VOCs) &                                     \\
+    \hline
+  \end{tabular}
+\end{table}
 
@@
 first without and then with overrides. Of the 1430 trajectories, approximately
-2\% required the use of harmonization overrides after an initial investigation;
+10\% were reported as a diagnostic (see Section \ref{sec:workflow}) of which 2\%
+required the use of harmonization overrides after an initial investigation;
 thus, 98\% of all trajectories were satisfactorily determined using the default

### 2017-06-09 analysis/get_results.sh
added script to get results

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+wget http://data.ene.iiasa.ac.at/gidden/harmonization-paper-results.tar.gz
+tar xvf harmonization-paper-results.tar.gz 

### 2017-06-09 results.tex
adding kyoto gas figure

@@
 harmonization mechanism's response to generally decreasing emission
-trajectories.
-% * <sfujimori112256@gmail.com> 2017-06-07T10:22:28.473Z:
-% 
-% As a supplementary, it would be better to show the total kyoto gases emissions for these two scenarios, which gives readers rough image what kind of scenarios we choose.
-% 
-% ^.
+trajectories. The emissions of Kyoto Gases, shown in Figure \ref{fig:kyoto},
+display the relative difference in trends between scenarios.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_kyoto.pdf}
+    \caption[]{
+      \label{fig:kyoto}
+      Unharmonized Kyoto gas emissions for SSP2-Ref, a scenario with generally
+      increasing global emissions trends, and SSP2-45, a scenario with generally
+      decreasing global emissions trends.  }
+  \end{center}
+\end{figure}
 

### 2017-06-09 results.tex
still have to say table

@@
 taken from previously described LUC and anthropogenic sources, which comprise 10
-separate pollutant and GHG species and 13 sectors shown in
+separate pollutant and GHG species and 13 sectors shown in Table
 \ref{tab:sp}. Therefore, a total of 1430 distinct trajectories were harmonized

### 2017-06-14 intro.tex
incorporating keywan's comments in intro

@@
 model (ESM) simulations, climate impact, adaptation and climate mitigation
-studies. They consequently underpin the United Nations' Intergovernmental Panel
-on Climate Change (IPCC) across working groups.
+studies. They are regularly assessed by the United Nations' Intergovernmental
+Panel on Climate Change (IPCC) across working groups.
 
@@
 technologies (e.g., NOx and SOx scrubbers on power plants). However, the
-historical emissions inventories used by different models can differ by large
-amounts depending on the region, sector, and emissions species being modeled.
+historical emissions inventories calculated by different models can differ by
+large amounts depending on the region, sector, and emissions species being
+modeled.
 
@@
 technology. Models are then \textit{calibrated} to historical data sources which
-provide both the activity of emitting technologies as well as trajectories of
-past emissions in order to determine historical emissions factors in each model
-region for each technology.
+provide a combination of the activity of emitting technologies and trajectories
+of past emissions in order to determine historical emissions trajectories in
+each model region for each technology.
 
@@
 historical emissions dataset for both anthropogenic emissions (i.e., the
-Community Emissions Datasystem (CEDS) \cite{hoesly_historical_2017} and
+Community Emissions Data System (CEDS) \cite{hoesly_historical_2017} and
 land-use change (LUC) emissions \cite{van_marle_historic_2017}) which, in
@@
 
-Because the process of model calibration is quite onerous, model teams initially
-incorporate updated historical datasets through the process of
-\textit{harmonization} in order to guarantee consistency with these new
-historical datasets. Harmonization describes the process of updating model
-results to incorporate new historical time series while maintaining critical
-aspects of modeled future trajectories impacted outputs. In the emissions
-context, this means that each individual combination of model region, model
-sector, and emissions species must be harmonized. Depending on the total number
-of model regions, sectors, and emissions species, this can require the selection
-of thousands to tens-of-thousands of harmonization methods.
+When participating in intercomparison exercises in which consistency of
+historical trajectories is required (e.g., in CMIP6), model teams incorporate a
+single, common historical dataset through \textit{harmonization} in order to
+enhance comparability across models. Harmonization describes the process of
+adjusting model results to incorporate new historical time series while
+maintaining critical aspects of and providing a smooth transition to modeled
+future trajectories such that the resulting trajectories are consistent with the
+original modeled results and usable by other members of the scientific community
+involved in the exercise. In the emissions context, this means that each
+individual combination of model region, model sector, and emissions species must
+be harmonized. Depending on the total number of model regions, sectors, and
+emissions species, this can require the selection of thousands to
+tens-of-thousands of harmonization methods.
 
@@
 \cite{rogelj_discrepancies_2011} describes the impacts of choosing various
-harmonization routines on future trajectories. Importantly, the choice of
-harmonization method to date has been determined via expert solicitation and has
-generally been applied to all trajectories for a given class of emissions
-species.
+harmonization routines on future trajectories. During the evaluation of the
+Representative Concentration Pathways (RCPs), IAM results have been harmonized
+by sector and the 5 RCP global regions
+\cite{vuuren_representative_2011}. Importantly, the choice of harmonization
+method to date has been determined via expert solicitation and has generally
+been applied to all trajectories for a given class of emissions species.
 
-In recent studies, i.e., the Representative Concentration Pathways (RCPs), IAM
-results have been harmonized by sector and the 5 RCP global regions
-\cite{vuuren_representative_2011}. Climate modeling efforts have continued to
-progress, however, demanding increased spatial and sectoral resolution from
-IAMs. In order to address the growing dimensionality of model outputs while
-still providing a consistent and scientifically rigorous harmonization
-procedure, an automated process for determining harmonization methods is
-needed. The remainder of this paper describes the methodology of the
+Climate modeling efforts have continued to progress, demanding increased spatial
+and sectoral resolution from IAMs. Furthermore, a new generation of climate
+scenarios which combines aspects of both the RCPs and SSPs have been developed
+in order to incorporate both physical and socioeconomic detail. In order to
+address the growing dimensionality of model outputs and support ongoing scenario
+generation and analysis efforts while still providing a consistent and
+scientifically rigorous harmonization procedure, an automated process for
+determining harmonization methods is needed. The use of an automated,
+documented, and openly available harmonization mechanism additionally allows for
+full procedural reproducibility and for direct participation by additional
+modeling teams not involved in the original exercise.
+
+The remainder of this paper describes the methodology and implementation of the
 \code{aneris} code base \cite{matthew_gidden_2017_802832}, written in the Python
 programming language (detailed documentation available
-\href{http://mattgidden.com/aneris/}{online}), which implements such a system in
-Section \ref{sec:meths}. The results of applying the automated harmonization
-mechanism on two example IAM scenarios, one with emissions growth and another
-with emissions mitigation, is presented in Section \ref{sec:results}. Finally,
-the general effectiveness and potential future improvements on the automated
-methodology is discussed in Section \ref{sec:future}.
+\href{http://mattgidden.com/aneris/}{online}). Section \ref{sec:meths} provides
+a detailed description of the underlying mathematical components of
+\code{aneris} as well as the code workflow. The results of applying the
+automated harmonization mechanism on two example IAM scenarios, one with
+emissions growth and another with emissions mitigation, is presented in Section
+\ref{sec:results}. Finally, the general effectiveness and potential future
+improvements on the automated methodology is discussed in Section
+\ref{sec:future}.
 

### 2017-06-14 elsfrontmatter.tex
update abstract with keywan comments

@@
 \begin{abstract}
-Harmonization describes the process of calibrating Integrated Assessment Model
-(IAM) GHG and air pollutants results with a new source of historical emissions trajectories. To date,
-harmonization has been performed separately by individual modeling teams through
+Harmonization describes the process required to compare Integrated Assessment
+Model (IAM) GHG and air pollutants results against a common source of historical
+emissions trajectories. To date, harmonization has been performed separately by
+individual modeling teams through
 % * <sfujimori112256@gmail.com> 2017-05-24T19:26:26.142Z:
@@
 % ^.
-the use of expert opinion and solicitation. As models and their results become
-more complex in both regional and sectoral dimensions, an automated approach to
-choosing harmonization methods becomes necessary. This work describes a novel
+the use of expert opinion and solicitation. In order to support ongoing and
+future global climate and socioeconomic analyses, incorporate model advancements
+towards more complex regional and sectoral representations, and expand the
+capability of modeling teams to participate in intercomparison exercises, an
+automated approach to choosing harmonization methods becomes necessary. This
+work describes a novel
 % * <sfujimori112256@gmail.com> 2017-05-24T20:13:08.493Z:

### 2017-06-14 future.tex
adding future work edits

@@
 output of \code{aneris} and thus can be individually explained. This provides
-not only transparency but also scientific integrity in the choice of
-harmonization methods.
+not only transparency and reproducibility, but also scientific integrity in the
+choice of harmonization methods.
+
+The use of an open-source, automated harmonization process also provides
+benefits to the wider climate science and IAM communities. For example, other
+modeling teams are easily capable of executing identical harmonization methods
+in order to participate in ongoing and further iterations of intercomparison
+exercises and analysis. Future scenario analyses can also utilize this common
+harmonization approach such that they are consistent with prior efforts.
 

### 2017-06-14 intro.tex
adding future work edits

@@
 of past emissions in order to determine historical emissions trajectories in
-each model region for each technology.
+each model region for each technology. Due to the sometimes large uncertainties
+in historical datasets, results in the historical period may differ between
+models.
 
-With some frequency, however, historical data is updated to incorporate new
-assumptions of source activity, new emissions accounting practices, or other
-concerns. The global climate change community has recently developed a standard
-historical emissions dataset for both anthropogenic emissions (i.e., the
-Community Emissions Data System (CEDS) \cite{hoesly_historical_2017} and
-land-use change (LUC) emissions \cite{van_marle_historic_2017}) which, in
+With some frequency, historical data is updated to incorporate new assumptions
+of source activity, new emissions accounting practices, or other concerns. The
+global climate change community has recently developed a standard historical
+emissions dataset for both anthropogenic emissions (i.e., the Community
+Emissions Data System (CEDS) \cite{hoesly_historical_2017} and land-use change
+(LUC) emissions \cite{van_marle_historic_2017}) which, in
 % * <sfujimori112256@gmail.com> 2017-06-07T09:59:35.820Z:

### 2017-06-14 meth.tex
make the table not int he middle of equations
adding cases figure
update methods section with example figure

@@
 
-\begin{table}[]
+\begin{table}[h!]
 \centering
@@
 required as a ratio method would mask future emissions and erroneously harmonize
-the trajectory. In most cases, however, models do report values in the
-harmonization year. When model and historical values are relatively close, a
+the trajectory.
+
+In most cases, however, models do report values in the harmonization
+year. Figure \ref{fig:cases} displays a number of example trajectories which
+highlight the possible issues resulting from harmonizing model results in
+different contexts. When model and historical values are relatively close, a
 convergence method is chosen in order to be as representative as possible to the
-underlying unharmonized model results. The final model-based characteristic of
-import is whether the model reports negative emissions. Such a case is possible
-for gas species which can be extracted from the environment and stored, as is
-the case for CO2 in future scenarios with climate mitigation. If a model
-% * <sfujimori112256@gmail.com> 2017-06-07T10:07:21.265Z:
-% 
-% > with climate mitigation
-% Not necessarily high mitigation. Land use can be negative even in baseline.
-% 
-% ^.
-provides a trajectory that transitions from positive to negative emissions, then
-a convergence method is used in order to guarantee capture of this
-transition. Otherwise, it is possible, given the relative values between history
-and model in the harmonization year, for a negative trajectory to be
-inappropriately harmonized to a positive, but decreasing, trajectory.
+underlying unharmonized model results (Figure \ref{fig:cases}, Panel
+\code{a}). If values are not close, the constant ratio method is chosen in order
+to provide reasonable trajectories that still incorporate modeled effects
+(Figure \ref{fig:cases}, Panel \code{b}).
+
+Models can additionally report negative emissions in certain contexts which much
+be taken into account during harmonization. Such a case is possible for gas
+species which can be extracted from the environment and stored, as is the case
+for CO2 in future scenarios with climate mitigation. If a model provides a
+trajectory that transitions from positive to negative emissions and base year
+results are similar, then a convergence method is used in order to guarantee
+capture of this transition in a representative fashion (Figure \ref{fig:cases},
+Panel \code{c}). If the discrepancy in base year results is large, it is
+possible for a negative trajectory to be inappropriately harmonized to a
+positive, but decreasing, trajectory. As such, the constant ratio method is
+chosen (Figure \ref{fig:cases}, Panel \code{d}).
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{cases.pdf}
+    \caption[]{
+      \label{fig:cases}
+      The effect of different harmonization routines on model trajectories under
+      ``normal'' circumstances (Panel \code{a}), when there is a large
+      difference between historical and model values in the harmonization year
+      (Panels \code{b} and \code{d}), and when model trajectories result in
+      negative emissions by the end of the modeling time horizon (Panels
+      \code{c} and \code{d}). In each figure row (Panels \code{a}, \code{b};
+      \code{c}, \code{d}), identical model trajectories are used. In each figure
+      column, historical values are increased in the base year by an order of
+      magnitude (from 10 to 100). In each case, a subset of the potential
+      routines provide a better ``goodness of fit'' than others as described in
+      the text.  }
+  \end{center}
+\end{figure}
 

### 2017-06-17 elsfrontmatter.tex
adding all authors

@@
 
-\author[iiasa,nies]{Shinichiro Fujimori}
+\author[pbl]{Maarten van den Berg}
+\author[nies]{Shinichiro Fujimori}
+\author[pik]{David Klein}
 \author[iiasa]{Keywan Riahi}
+\author[pnnl]{Steven Smith}
+\author[pbl]{Detlef van Vuuren}
 
@@
 \address[nies]{National Institute for Environmental Studies, Tsukuba, Japan}
+\address[pbl]{PBL Netherlands Environmental Assessment Agency, Postbus 30314, 2500 GH The Hague, Netherlands}
+\address[pik]{Potsdam Institute for Climate Impact Research, Telegraphenberg A 31, 14473 Potsdam, Germany}
+\address[pnnl]{Joint Global Change Research Institute, 5825 University Research Court, Suite 3500, College Park, MD 20740}
 

### 2017-07-01 elsfrontmatter.tex
sjs Revisions

@@
 \author[iiasa]{Keywan Riahi}
-\author[pnnl]{Steven Smith}
+\author[pnnl]{Steven J. Smith}
 \author[pbl]{Detlef van Vuuren}
@@
 \begin{abstract}
-Harmonization describes the process required to compare Integrated Assessment
+Harmonization describes the process required to match Integrated Assessment
 Model (IAM) GHG and air pollutants results against a common source of historical

### 2017-07-01 intro.tex
sjs Revisions

@@
 interactions between energy systems, economic systems, land use, the climate,
-and air pollution, among other sectors . IAMs provide global projections of
+and air pollution. IAMs provide global projections of
 systemic change by dividing the world into a number of representative regions
@@
 futures. Recently, flagship IAMs have developed scenarios based on the Shared
+% * <ssmith@pnnl.gov> 2017-07-01T03:11:26.621Z:
+% 
+% > flagship
+% Suggest different word. "comprehensive" is one possibility.
+% 
+% ^ <ssmith@pnnl.gov> 2017-07-01T03:11:57.571Z.
 Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
   fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
-quantify a variety of potential global futures. The SSPs are primarily designed
-to be used in analyses of broad climate research which include earth system
+quantify a variety of potential global futures. The SSPs are  designed
+to be used in  research that include earth system
 model (ESM) simulations, climate impact, adaptation and climate mitigation
-studies. They are regularly assessed by the United Nations' Intergovernmental
-Panel on Climate Change (IPCC) across working groups.
+studies.
 
 While IAMs are implemented in myriad ways, including simulation and
-optimization, the core inputs, outputs, and purpose are similar across different
-models. Modeling teams incorporate existing data sets of energy systems, land
+optimization, the core inputs, and outputs  are similar across different
+models. Modeling teams incorporate  data on energy systems, land
 use, economics, demographics and emissions sources and concentrations, among other
-data, in order to provide consistent historical trajectories of modeled
+data, in order to provide consistent future trajectories of modeled
 variables. The models then provide estimates of future trajectories of these
@@
 technologies (e.g., NOx and SOx scrubbers on power plants). However, the
-historical emissions inventories calculated by different models can differ by
-large amounts depending on the region, sector, and emissions species being
-modeled.
+historical emissions starting points of  models can differ by
+large amounts depending on the region, sector, and emissions species.
 
@@
 emissions-intensity factors for individual gas species, for example the quantity
-of sulfur emissions from coal plants per megawatt-hour of production, in order
-to determine total emissions of a given gas species from a given
-technology. Models are then \textit{calibrated} to historical data sources which
-provide a combination of the activity of emitting technologies and trajectories
-of past emissions in order to determine historical emissions trajectories in
-each model region for each technology. Due to the sometimes large uncertainties
+of sulfur emissions from coal plants per megawatt-hour of production.
+Models are generally \textit{calibrated} to historical data sources in one or more base years.
+Due to the sometimes large uncertainties
 in historical datasets, results in the historical period may differ between
-models.
+models. Models also differ in their choice of base-year, which may lag behind available inventory data. In addition, models have varying sectoral, regional, and fuel aggregations.
 
-With some frequency, historical data is updated to incorporate new assumptions
-of source activity, new emissions accounting practices, or other concerns. The
-global climate change community has recently developed a standard historical
+The
+global climate change community has recently developed a new global historical
 emissions dataset for both anthropogenic  emissions (i.e., the Community
-Emissions Data System (CEDS) \cite{hoesly_historical_2017} and land-use change
-(LUC) emissions \cite{van_marle_historic_2017}) which, in
+Emissions Data System (CEDS) \cite{hoesly_historical_2017} and open-burning land-use and land-use change (LULUC)
+emissions \cite{van_marle_historic_2017}) which, in
 % * <sfujimori112256@gmail.com> 2017-06-07T09:59:35.820Z:
@@
 
-When participating in intercomparison exercises in which consistency of
-historical trajectories is required (e.g., in CMIP6), model teams incorporate a
-single, common historical dataset through \textit{harmonization} in order to
-enhance comparability across models. Harmonization describes the process of
-adjusting model results to incorporate new historical time series while
-maintaining critical aspects of and providing a smooth transition to modeled
-future trajectories such that the resulting trajectories are consistent with the
-original modeled results and usable by other members of the scientific community
-involved in the exercise. In the emissions context, this means that each
+When participating in intercomparison exercises in which a consistent historical starting point is required (e.g., in CMIP6), model teams incorporate a
+single, common historical dataset through \textit{harmonization}. Harmonization describes the process of
+adjusting model results to match the selected historical time series  such that the resulting future trajectories are also consistent with the
+original modeled results. In the emissions context, this means that each
 individual combination of model region, model sector, and emissions species must

### 2017-07-01 meth.tex
sjs Revisions

@@
 regions ($r$), species ($g$), and sectors ($s$), must be harmonized to the
-initial time step. Given a model trajectory, $m_{r, g, s}(t)$, historical
-trajectory, $h_{r, g, s}(t)$, and initial time step $t_i$, a harmonized
-trajectory can be calculated. The ``goodness of fit'' of a trajectory depends on
+initial starting point. Given a model trajectory, $m_{r, g, s}(t)$, historical
+values, $h_{r, g, s}(t)$, and initial starting point $t_i$, a harmonized
+trajectory needs to be calculated. The ``goodness of fit'' of a trajectory depends on
 a number of factors, most notably the faithful representation of the original,
@@
 possible) which are of critical importance for cumulative CO2 calculations.
+% * <ssmith@pnnl.gov> 2017-07-01T03:25:04.357Z:
+% 
+% Suggest moving this later where negative emissions are discussed. Its a bit of a tangent here.
+% 
+% ^ <ssmith@pnnl.gov> 2017-07-01T03:25:44.627Z.
 
@@
 
-Variability of the historical trajectory is also an important characteristic
-when considering the choice of harmonization method. Such trajectories are
-normally indicative of emissions from land use sectors. Land use sectors are
+Temporal variability of the historical trajectory is also an important characteristic
+when considering the choice of harmonization method.
+Emissions from forest and grassland fires, for example, vary from year to year due to a combination of meteorological conditions and anthropogenic drivers.
+Land use emissions are modeled by IAMs using average emission factors and do not capture conditions in a specific year.
+Ratio methods for such sectors can lead to unrealistic harmonization results.
 modeled with varying fidelity by IAMs, but in general proxies are used for
@@
 harmonization method is important because the effects are modeled similarly
-across regions and species. In order to reliably detect land-use-like emissions,
+across regions and species. In order to detect land-use-like emissions,
+% * <ssmith@pnnl.gov> 2017-07-01T03:31:52.910Z:
+% 
+% > modeled with varying fidelity by IAMs, but in general proxies are used for
+% > emissions trajectories. Take for example emissions from grassland burning: for
+% > most models, the use of grass and paturelands is endogenously modeled, and
+% > emissions from the burning of these lands is determined by applying an emissions
+% > factor; in general, the more grassland usage that occurs in a scenario, the
+% > higher the emissions from the burning thereof. Therefore, consistency in
+% > harmonization method is important because the effects are modeled similarly
+% > across regions and species. 
+% 
+% Not sure I follow the logic here. I added some new text that might help. Perhaps the intermediate part could be deleted.
+% 
+% ^ <ssmith@pnnl.gov> 2017-07-01T03:32:30.453Z.
 a measure of the coefficient of variation of the first derivative of the
@@
       values in the harmonization year is shown. The solid black line indicates
-      the threshold value used by default in \code{aneris}.  
+      the 50\% threshold value used by default in \code{aneris}.  
     }

### 2017-07-01 results.tex
sjs Revisions

@@
 trajectories. The emissions of Kyoto Gases, shown in Figure \ref{fig:kyoto},
-display the relative difference in trends between scenarios.
+shows the different trends of these two scenarios.
 
@@
 required the use of harmonization overrides after an initial investigation;
-thus, 98\% of all trajectories were satisfactorily determined using the default
-methods as chosen by the harmonization algorithm. Figure \ref{fig:nox} provides
+thus, 98\% of all trajectories were satisfactorily harmonized using the default
+methods. Figure \ref{fig:nox} provides
 one such example of a emissions species and sector in which all regions were
@@
 method is required to be provided for the trajectory of the region, sector, and
-species in question. Based on the 2\% of total trajectories required to be
-overridden, two classifications were observed: regional trajectories whose
+species in question. The 2\% of trajectories that required
+overrides clustered into two classifications: regional trajectories whose
 \textit{magnitude} was overly distorted and regional trajectories whose

### 2017-07-03 future.tex
sjs Revisions
Update on Overleaf.

@@
 harmonized over 98\% of the 2860 individual trajectories that were analyzed in
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T23:07:25.083Z:
+% 
+% > 98\%
+% Do we know to what extent does this compare to other model results?
+% 
+% ^.
 Section \ref{sec:results}. Of the remaining trajectories, harmonization method
-overrides were applied, and the situation in which the need for overrides arose
-was discussed.
+overrides were applied, and example situations in overrides were needed
+were discussed.
 
@@
 sources are the likely cause of a harmonization year discrepancy (suggesting the
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T23:15:16.273Z:
+% 
+% > missing
+% > sources
+% it might also be a difference in activity levels, which might be something different than missing sources.
+% 
+% ^.
 use of an offset method) or instead a significant difference in emissions
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T23:16:40.654Z:
+% 
+% > emissions
+% > factors (suggesting the use of a ratio method)
+% knowledge of these factors might even lead to combinations of harmonization methods for a single sector/gas/region combination as they are aggregate categories.
+% 
+% ^.
 factors (suggesting the use of a ratio method) \cite{rogelj_discrepancies_2011}.

### 2017-07-03 intro.tex
Update on Overleaf.

@@
 technologies (e.g., NOx and SOx scrubbers on power plants). However, the
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T20:34:35.330Z:
+% 
+% I think the mitigation options are broad, the example of the scrubbers seems too narrow, I would leave it out.
+% 
+% ^.
 historical emissions starting points of  models can differ by
@@
 % I think this one also anthropogenic
+% 
+% ^ <maarten.vandenberg@pbl.nl> 2017-07-03T20:46:42.206Z:
+% 
+% I think so too, however I believe the GFED also contains natural emissions.
 %
@@
 \code{aneris} code base \cite{matthew_gidden_2017_802832}, written in the Python
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:12:05.654Z:
+% 
+% > the
+% > \code{aneris} code base \cite{matthew_gidden_2017_802832
+% perhaps something like: "the harmonization code base \code{aneris}", to be explicit.
+% 
+% ^.
 programming language (detailed documentation available

### 2017-07-03 meth.tex
Update on Overleaf.

@@
 trajectory needs to be calculated. The ``goodness of fit'' of a trajectory depends on
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:16:50.244Z:
+% 
+% Assessing the "goodness of fit" is not  straightforward, there may be competing factors. Perhaps we should refer back to this in the discussion. 
+% 
+% ^.
 a number of factors, most notably the faithful representation of the original,
@@
 base provides a number of variations on the classic methods, including
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:27:39.375Z:
+% 
+% > classic methods
+% This might be interpreted as that convergence have not been previously applied. Perhaps use " implementations"  instead of "variations" .
+% 
+% ^.
 ratio-convergence methods shown in Equation \ref{eqs:ratio}, offset-convergence
@@
 summary of all available methods is provided in Table \ref{tab:meths}.
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:34:25.905Z:
+% 
+% > A
+% > summary
+% If beta would be defined as \frac{t -t_i}{t_f - t_i}, it might better match intuition of convergence: 0 value at t = t_i and 1 at full convergence (t = t_f), with added benefit the beta parameter can also be used in the interpolation option.
+% 
+% ^.
 
@@
       The default method decision tree used in the \code{aneris} software
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:40:43.868Z:
+% 
+% > The default method decision tree
+% yes and no might be added to the tree for improved clarity. 
+% 
+% ^.
       library. For all decisions, upper branches represent a ``yes'' response
@@
       thresholds are described below. Finally, default choices in \textit{green}
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:42:35.965Z:
+% 
+% for CoV and dH
+% 
+% ^.
       label methods that will likely closely match unharmonized results, choices
@@
 Ratio methods for such sectors can lead to unrealistic harmonization results.
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:59:54.238Z:
+% 
+% > Ratio methods for such sectors can lead to unrealistic harmonization results.
+% However, using for instance an offset method may also misrepresent the emission levels. Depending on the trajectory the emission factor can be subject to unintentional changes over the century.
+% 
+% ^.
 modeled with varying fidelity by IAMs, but in general proxies are used for
@@
 non-LUC trajectories) and false negatives (~7\% of LUC trajectories). However,
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:07:21.157Z:
+% 
+% >  (~10\% of
+% > non-LUC trajectories) and false negatives (~7\% of LUC trajectories)
+% visually it seems to be that the false negatives are a larger share thatn the false positives?
+% 
+% ^.
 as any regional definition is model dependent and thus any regional aggregation
@@
 \begin{equation}\label{eqs:cov}
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:08:21.230Z:
+% 
+% > \begin{equation}\label{eqs:cov}
+% >     c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
+% > \end{equation}
+% c_v is called a covariance parameter and a variance of derivatives, what is exactly calculated?
+% 
+% ^.
     c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
@@
 value of 50\% was chosen to be used as a default in \code{aneris}.
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:25:47.452Z:
+% 
+% > 50\%
+% What does this mean? I would expect something of the form low_value < dH < high_value?
+% 
+% ^.
 
@@
       the 50\% threshold value used by default in \code{aneris}.  
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:14:59.644Z:
+% 
+% > The distribution of relative differences between model and historical
+% >       values in the harmonization year is shown. The solid black line indicates
+% >       the 50\% threshold value used by defaul
+% the asymmetry here might be interesting, do the SSP scenarios consistently underestimate emission levels, or is this an effect of history up to 2015, while scenarios start earlier (might not be interesting for this paper).
+% 
+% ^.
     }

### 2017-07-11 intro.tex
incorporating edits in intro + flagship question

--- 2017-07-01 intro.tex
@@
 international studies, which notably include projections of climate and economic
-futures. Recently, flagship IAMs have developed scenarios based on the Shared
+futures. Recently, the IAM community has developed scenarios based on the Shared
 % * <ssmith@pnnl.gov> 2017-07-01T03:11:26.621Z:
@@
 While IAMs are implemented in myriad ways, including simulation and
-optimization, the core inputs, and outputs  are similar across different
+optimization, the core inputs and outputs  are similar across different
 models. Modeling teams incorporate  data on energy systems, land
 use, economics, demographics and emissions sources and concentrations, among other
-data, in order to provide consistent future trajectories of modeled
+data, in order to provide consistent existing trajectories of modeled
 variables. The models then provide estimates of future trajectories of these
@@
 emissions of different air pollutants and GHGs as well as various mitigation
-technologies (e.g., NOx and SOx scrubbers on power plants). However, the
-historical emissions starting points of  models can differ by
+technologies. However, the historical emissions starting points of  models can differ by
 large amounts depending on the region, sector, and emissions species.
@@
 The remainder of this paper describes the methodology and implementation of the
-\code{aneris} code base \cite{matthew_gidden_2017_802832}, written in the Python
+harmonization code base \code{aneris} \cite{matthew_gidden_2017_802832}, written in the Python
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:12:05.654Z:
+% 
+% > the
+% > \code{aneris} code base \cite{matthew_gidden_2017_802832
+% perhaps something like: "the harmonization code base \code{aneris}", to be explicit.
+% 
+% ^.
 programming language (detailed documentation available
@@
 \ref{sec:future}.
-
-% todo: add?
-% what aspects of harmonization are important (e.g., cumulative carbon emissions).

### 2017-07-11 meth.tex
incorporating edits in intro + flagship question

@@
 $c_v$s as determined for historical data aggregated to the model regions of 5
-different flagship IAMs. A threshold value of 20 has been chosen based on these
+different IAMs \TODO{list here or elsewhere?}. A threshold value of 20 has been chosen based on these
 observations as it optimally divides the two distributions. Importantly, tails
@@
 investigate the possible values that these relative differences can take, the
-flagship IAMs values used in the SSP and (ongoing) CMIP6 inter-comparison
+IAM values used in the SSP and (ongoing) CMIP6 inter-comparison
 exercises are used. A distribution of these differences for all models in the

### 2017-07-11 meth.tex
simple word change in meths, more interesting updates to follow here
use beta in interpolation method
update beta and caption in method tree

@@
 offset of base year values (i.e., $h(t_i) - m(t_i)$). The \code{aneris} code
-base provides a number of variations on the classic methods, including
+base provides a number of implementations on the classic methods, including
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T21:27:39.375Z:
@@
 methods shown in Equation \ref{eqs:offset}, and interpolation methods shown in
-Equation \ref{eqs:interpolate}; the convergence factor, $\beta$, for associated
-methods is shown first in Equation \ref{eqs:factor}. Note that region, species,
-and sector indices have been dropped for clarity. Each equation is a function of
-time, model trajectory, historical trajectory, harmonization time step, and a
-convergence time step ($t_f$), at which point the harmonized trajectory
-converges to the unharmonized trajectory. The \code{aneris} code base provides a
-number of methods to choose from for each of the harmonization families. A
-summary of all available methods is provided in Table \ref{tab:meths}.
+Equation \ref{eqs:interpolate}. The convergence factor, $\beta$, scales linearly
+from 1 to 0 in $[t_i, t_f)$ and is shown in Equation \ref{eqs:factor}. Note
+that region, species, and sector indices have been dropped for clarity. Each
+equation is a function of time, model trajectory, historical trajectory,
+harmonization time step, and a convergence time step ($t_f$), at which point the
+harmonized trajectory converges to the unharmonized trajectory. The
+\code{aneris} code base provides a number of methods to choose from for each of
+the harmonization families. A summary of all available methods is provided in
+Table \ref{tab:meths}.
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T21:34:25.905Z:
@@
   \begin{cases}
-    1 - \frac{t - t_i}{t_f - t_i},& \text{if } t \leq t_f\\
+    1 - \frac{t - t_i}{t_f - t_i},& \text{if } t < t_f\\
     0,                        & \text{otherwise}
@@
   \begin{cases}
-    \frac{m(t_f) - h(t_i)}{t_f - t_i}(t - t_i) + h(t_i), & \text{if } t \leq t_f\\
+    \beta(t, t_i, t_f) (h(t_i) - m(t_f)) + m(t_f), & \text{if } t < t_f\\
     m(t), & \text{otherwise}
@@
       $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and decision-making
-      thresholds are described below. Finally, default choices in \textit{green}
+      thresholds are described below. Methods labeled in \textit{green}
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T21:42:35.965Z:
@@
 % ^.
-      label methods that will likely closely match unharmonized results, choices
+      are likely closely match unharmonized results, methods
       in \textit{yellow} will likely somewhat match unharmonized results, and
-      choices in \textit{red} can be expected to have a large relative
+      methods in \textit{red} can be expected to have a large relative
       difference between harmonized and unharmonized results.}

### 2017-07-11 meth.tex
finish meth comments, update notebook for luc/non thresholds

@@
       $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and decision-making
-      thresholds are described below. Methods labeled in \textit{green}
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:42:35.965Z:
-% 
-% for CoV and dH
-% 
-% ^.
-      are likely closely match unharmonized results, methods
+      thresholds for $c_v$ and $dH$ are described below. Methods labeled in
+      \textit{green} are likely to closely match unharmonized results, methods
       in \textit{yellow} will likely somewhat match unharmonized results, and
@@
 $c_v$s as determined for historical data aggregated to the model regions of 5
-different IAMs \TODO{list here or elsewhere?}. A threshold value of 20 has been chosen based on these
-observations as it optimally divides the two distributions. Importantly, tails
-of the LUC and non-LUC overlap, thus there are both false positives (~10\% of
-non-LUC trajectories) and false negatives (~7\% of LUC trajectories). However,
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:07:21.157Z:
-% 
-% >  (~10\% of
-% > non-LUC trajectories) and false negatives (~7\% of LUC trajectories)
-% visually it seems to be that the false negatives are a larger share thatn the false positives?
-% 
-% ^.
-as any regional definition is model dependent and thus any regional aggregation
-is possible an automated detection mechanism is necessary.
+different IAMs involved in the SSP process. A threshold value of 20 has been
+chosen based on these observations as it optimally divides the two
+distributions. Importantly, tails of the LUC and non-LUC overlap, thus there are
+both false positives (~7\% of non-LUC trajectories) and false negatives (~10\%
+of LUC trajectories). However, as any regional definition is model dependent and
+thus any regional aggregation is possible an automated detection mechanism is
+necessary.
 

### 2017-07-11 meth.tex
clairifying thresholds

@@
 $c_v$s as determined for historical data aggregated to the model regions of 5
-different IAMs involved in the SSP process. A threshold value of 20 has been
+different IAMs involved in the SSP process. A threshold value of $c_v < 20$ has been
 chosen based on these observations as it optimally divides the two
@@
 study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
-value of 50\% was chosen to be used as a default in \code{aneris}.
+value of $dH < 50$\% was chosen to be used as a default in \code{aneris}.
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T22:25:47.452Z:

### 2017-07-11 results.tex
Update on Overleaf.
rm unharmonized sector

@@
 taken from previously described LUC and anthropogenic sources, which comprise 10
-separate pollutant and GHG species and 13 sectors shown in Table
+separate pollutant and GHG species and 12 sectors shown in Table
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:41:59.852Z:
+% 
+% >  which comprise 10
+% > separate pollutant and GHG species and 13 sectors shown in Table
+% Do we correctly distinguish between the regional and global values (N2O)?
+The table mentions a sector "Unharmonized", which is the sector-aggregated value I assume, which also might lead to overcounting.
+In the table I would change this sector naming.
+% 
+% ^.
 \ref{tab:sp}. Therefore, a total of 1430 distinct trajectories were harmonized
@@
     Organic Carbon (OC)               & Transportation Sector               \\
-    Sulfur Hexafluoride (SF6)         & Unharmonized                        \\
-    Sulfur Oxides (SOx)               & Waste                               \\
+    Sulfur Hexafluoride (SF6)         & Waste                               \\
+    Sulfur Oxides (SOx)               &                                     \\
     Volatile Organic Compounds (VOCs) &                                     \\
@@
 more closely with the unharmonized trajectory.
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:53:14.757Z:
+% 
+% > the global trajectory for industrial CO also corresponds
+% > more closely with the unharmonized trajectory.
+% The global trajectory suffers under the constant_ratio method from changing relative importance of underlying regions, resulting in a trend change, this effect is also prevented by using a constant_offset method.
+% 
+% ^.
 
@@
 degree. While this distortion changes the magnitude of the SAS trajectory, it
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:57:57.074Z:
+% 
+% > Notably, the SSP2-45 scenario is not affected to the same
+% > degree.
+% First, I think other effects are in play here.
+The unharmonized SSP2 -ref scenario has a discontinuity in it's derivative for the SAS region.
+This is amplified at the regional level by the constant_ratio approach and affects the 4.5 scenario in the same manner, that just happens to lack the discontinuity.
+Furthermore the trend of other regions is strengthened (a stronger decrease), leading up to an even stronger discontinuity in the global total.
+However, the ratio approach does a better job at representing the trend in the scenario.
+That trade-off is not completely clear now, also refers back to goodness of fit.
+Secondly, more problematic effects occur when applying reduce_ratio methods, that can lead to strong magnitude and trend changes; however this should only occur (significantly) when in override mode (if I understand the implementation of dH correctly).
+% 
+% ^.
 largely affects the post-2080 shape of the global trajectory (see Figure

### 2017-07-11 results.tex
some results commits

@@
 Figure \ref{fig:co} presents a case in which the magnitude of a trajectory is
-distorted. A large discrepancy ($\sim$300\% relative difference) is
-observed in the harmonization year for carbon monoxide (CO) emissions in the
-industrial sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region,
-which comprises most of the Asian subcontinent. The default method chosen
+distorted. A large discrepancy ($\sim$300\% relative difference) is observed in
+the harmonization year for carbon monoxide (CO) emissions in the industrial
+sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region, which
+comprises most of the Asian subcontinent. The default method chosen
 (\code{constant_ratio}) maintains model trends for the region; however, overall
@@
 regional trend and magnitude is maintained. With the new harmonization method
-for the SAS region, the global trajectory for industrial CO also corresponds
-more closely with the unharmonized trajectory.
+for the SAS region, the global trajectory for industrial CO also maintains the
+trends seen in the unharmonized trajectory.
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T22:53:14.757Z:

### 2017-07-12 meth.tex
clairfy diagnostics

@@
 regions ($r$), species ($g$), and sectors ($s$), must be harmonized to the
-initial starting point. Given a model trajectory, $m_{r, g, s}(t)$, historical
-values, $h_{r, g, s}(t)$, and initial starting point $t_i$, a harmonized
+initial modeling period. Given a model trajectory, $m_{r, g, s}(t)$, historical
+values, $h_{r, g, s}(t)$, and initial modeling period $t_i$, a harmonized
 trajectory needs to be calculated. The ``goodness of fit'' of a trajectory depends on
@@
 associated with the default method choice. In order to help identify these
-cases, harmonization \textit{diagnostics} are provided which analyze the relative
-difference between harmonized and unharmonized trajectories at their mid and
-end-points. If override methods are provided, they are used instead of the
-default methods as determined by the default method decision tree. Furthermore,
-users can set the above-mentioned thresholds as well as the LUC method used in
-the decision tree (see Figure \ref{fig:decision_tree}). Further detail of input
-parameters can be found \href{http://mattgidden.com/aneris/config.html}{online}.
+cases, harmonization \textit{diagnostics} are provided which analyze the
+relative difference between harmonized and unharmonized trajectories at their
+mid (if $> 400$\%) and end-points (if $> 200$\%). If override methods are
+provided, they are used instead of the default methods as determined by the
+default method decision tree. Furthermore, users can set the above-mentioned
+thresholds as well as the LUC method used in the decision tree (see Figure
+\ref{fig:decision_tree}). Further detail of input parameters can be found
+\href{http://mattgidden.com/aneris/config.html}{online}.
 

### 2017-07-12 results.tex
update xaxis for cv and an emission

@@
 methods. Figure \ref{fig:nox} provides
-one such example of a emissions species and sector in which all regions were
+one such example of an emissions species and sector in which all regions were
 satisfactorily harmonized with the default methods. As can be seen, results with

### 2017-07-12 future.tex
incorporating david's additional comments

@@
 
-The automated approach critically removes the need for expert opinion in
+The automated approach drastically reduces the need for expert opinion in
 determining harmonization methods for each individual combination of model

### 2017-07-12 meth.tex
incorporating david's additional comments

@@
 
-Models can additionally report negative emissions in certain contexts which much
+Models can additionally report negative emissions in certain contexts which must
 be taken into account during harmonization. Such a case is possible for gas
@@
 \hline
-history      & This historical value in the harmonization year.                             \\
+history      & The historical value in the harmonization year.                             \\
 \hline

### 2017-07-12 results.tex
incorporating david's additional comments

@@
 The harmonization of emissions pathways is performed in order to accurately
-represent an updated historical emissions trajectory while also maintaining
-consistency with the original, unharmonized pathway. As such, when the default
-methods as provided by the harmonization procedure distort or otherwise
-sufficiently misrepresent the underlying unharmonized results, an override
-method is required to be provided for the trajectory of the region, sector, and
-species in question. The 2\% of trajectories that required
-overrides clustered into two classifications: regional trajectories whose
+represent new or updated datasets of historical emissions inventories while also
+maintaining consistency with the original, unharmonized pathway. As such, when
+the default methods as provided by the harmonization procedure distort or
+otherwise sufficiently misrepresent the underlying unharmonized results, an
+override method is required to be provided for the trajectory of the region,
+sector, and species in question. The 2\% of trajectories that required overrides
+clustered into two classifications: regional trajectories whose
 \textit{magnitude} was overly distorted and regional trajectories whose

### 2017-07-12 future.tex
updating with maarten's comments

@@
 missing from the decision tree is input from models regarding whether missing
-sources are the likely cause of a harmonization year discrepancy (suggesting the
+sources or activity levels are the likely cause of a harmonization year
+discrepancy (suggesting the
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T23:15:16.273Z:

### 2017-07-12 results.tex
updating with maarten's comments

@@
 regional trend and magnitude is maintained. With the new harmonization method
-for the SAS region, the global trajectory for industrial CO also maintains the
-trends seen in the unharmonized trajectory.
+for the SAS region, the global trajectory for industrial CO also is
+representative the trends seen in the unharmonized trajectory and the relative
+importance of the underlying regional trajectoreis is maintained.
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T22:53:14.757Z:

### 2017-07-14 meth.tex
common form of reference for thresholds

@@
 
-The value of $c_v$ is then tested against a threshold. To determine this
-threshold, an analysis of the recent CEDS and LUC historical data has been
-performed. Figure \ref{fig:cov} shows the distribution of LUC $c_v$s and non-LUC
-$c_v$s as determined for historical data aggregated to the model regions of 5
-different IAMs involved in the SSP process. A threshold value of $c_v < 20$ has been
-chosen based on these observations as it optimally divides the two
-distributions. Importantly, tails of the LUC and non-LUC overlap, thus there are
-both false positives (~7\% of non-LUC trajectories) and false negatives (~10\%
-of LUC trajectories). However, as any regional definition is model dependent and
-thus any regional aggregation is possible an automated detection mechanism is
-necessary.
+The value of $c_v$ is then tested against a threshold, $\tau_{c_v}$. To
+determine this threshold, an analysis of the recent CEDS and LUC historical data
+has been performed. Figure \ref{fig:cov} shows the distribution of LUC $c_v$s
+and non-LUC $c_v$s as determined for historical data aggregated to the model
+regions of 5 different IAMs involved in the SSP process. A threshold value of
+$c_v < \tau_{c_v} = 20$ has been chosen based on these observations as it
+optimally divides the two distributions. Importantly, tails of the LUC and
+non-LUC overlap, thus there are both false positives (~7\% of non-LUC
+trajectories) and false negatives (~10\% of LUC trajectories). However, as any
+regional definition is model dependent and thus any regional aggregation is
+possible an automated detection mechanism is necessary.
 
@@
 study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
-value of $dH < 50$\% was chosen to be used as a default in \code{aneris}.
+value of $dH < \tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T22:25:47.452Z:

### 2017-07-17 meth.tex
update decision tree figure, update all result graphs with y axis 0

@@
       The default method decision tree used in the \code{aneris} software
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:40:43.868Z:
-% 
-% > The default method decision tree
-% yes and no might be added to the tree for improved clarity. 
-% 
-% ^.
-      library. For all decisions, upper branches represent a ``yes'' response
-      and lower branches represent a ``no'' response. $dH$ is defined as
-      $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and decision-making
-      thresholds for $c_v$ and $dH$ are described below. Methods labeled in
-      \textit{green} are likely to closely match unharmonized results, methods
-      in \textit{yellow} will likely somewhat match unharmonized results, and
-      methods in \textit{red} can be expected to have a large relative
-      difference between harmonized and unharmonized results.}
+      library. For all decisions, upper (purple) branches represent a ``yes''
+      response and lower (orange) branches represent a ``no'' response. $dH$ is
+      defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
+      decision-making thresholds for $c_v$ and $dH$ are described below. Methods
+      labeled in \textit{green} are likely to closely match unharmonized
+      results, methods in \textit{yellow} will likely somewhat match
+      unharmonized results, and methods in \textit{red} can be expected to have
+      a large relative difference between harmonized and unharmonized results.}
   \end{center}
@@
 regions of 5 different IAMs involved in the SSP process. A threshold value of
-$c_v < \tau_{c_v} = 20$ has been chosen based on these observations as it
+$\tau_{c_v} = 20$ has been chosen based on these observations as it
 optimally divides the two distributions. Importantly, tails of the LUC and
@@
 study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
-value of $dH < \tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.
+value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T22:25:47.452Z:

### 2017-07-17 elsfrontmatter.tex
Update on Overleaf.
incorporating david's additional comments
fix counting of harmonized trajectories, added footnotes to table

@@
 \address[pbl]{PBL Netherlands Environmental Assessment Agency, Postbus 30314, 2500 GH The Hague, Netherlands}
-\address[pik]{Potsdam Institute for Climate Impact Research, Telegraphenberg A 31, 14473 Potsdam, Germany}
+\address[pik]{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, D-14412 Potsdam, Germany}
 \address[pnnl]{Joint Global Change Research Institute, 5825 University Research Court, Suite 3500, College Park, MD 20740}
@@
 individual modeling teams through
+% * <maarten.vandenberg@pbl.nl> 2017-07-03T20:19:43.403Z:
+% 
+% IMAGE has provided harmonized emissions in the past, generated by a harmonization routine applying a ratio convergence method (on a higher aggregation level) (Hof et al., 2016).
+I think the contrast with these harmonization approaches and this work, which enables a documented, automated process selecting appropriate harmonization methodologies, might not be clear.
+Furthermore, later in the text cross-model harmonization methods (though less sophisticated) are mentioned.
+It is worth mentioning that previous approaches are far less sophisticated.
+% 
+% ^.
 % * <sfujimori112256@gmail.com> 2017-05-24T19:26:26.142Z:
@@
 code base for implementing the methodology. Results are shown for two example
-scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 98\%
+scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 96\%
 of the total emissions trajectories.

### 2017-07-17 results.tex
fix counting of harmonized trajectories, added footnotes to table

@@
 taken from previously described LUC and anthropogenic sources, which comprise 10
-separate pollutant and GHG species and 12 sectors shown in Table
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:41:59.852Z:
-% 
-% >  which comprise 10
-% > separate pollutant and GHG species and 13 sectors shown in Table
-% Do we correctly distinguish between the regional and global values (N2O)?
-The table mentions a sector "Unharmonized", which is the sector-aggregated value I assume, which also might lead to overcounting.
-In the table I would change this sector naming.
-% 
-% ^.
-\ref{tab:sp}. Therefore, a total of 1430 distinct trajectories were harmonized
-for each scenario.
+separate pollutant and GHG species and 12 sectors shown in Table \ref{tab:sp}. A
+total of 970 distinct trajectories were harmonized for each scenario, and
+therefore 1940 trajectories were harmonized in total.
 
-\begin{table}[]
+\begin{table}[]\begin{minipage}{\textwidth}
   \centering
@@
     Black Carbon (BC)                 & Agricultural Waste Burning          \\
-    Hexafluoroethane (C2F6)           & Agriculture                         \\
-    Tetrafluoromethane (CF4)          & Aircraft                            \\
+    Hexafluoroethane (C2F6)
+    \footnote{\label{x_glb_all}
+      Global total trajectories are harmonized due to lack of detailed 
+      historical data.}
+                                      & Agriculture                         \\
+    Tetrafluoromethane (CF4)
+    \footnoteref{x_glb_all}               & Aircraft
+    \footnote{\label{x_glb}
+      Global sectoral trajectories are harmonized due to lack of detailed 
+      historical data.}                            \\
     Methane (CH4)                     & Energy Sector                       \\
-    Carbon Dioxide (CO2)              & Forest Burning                      \\
+    Carbon Dioxide (CO2)
+    \footnote{\label{x_co2}
+      A global trajectory for land-use CO2 is used; non-land-use sectors 
+      are harmonized for each model region.}
+                                      & Forest Burning                      \\
     Carbon Monoxide (CO)              & Grassland Burning                   \\
-    Hydrofluorocarbons (HFCs)         & Industrial Sector                   \\
-    Nitrous Oxide (N2O)               & International Shipping              \\
+    Hydrofluorocarbons (HFCs)
+    \footnoteref{x_glb_all}               & Industrial Sector                   \\
+    Nitrous Oxide (N2O)
+    \footnoteref{x_glb_all}
+                                      & International Shipping\footnoteref{x_glb}              \\
     Ammonia (NH3)                     & Residential Commercial Other        \\
@@
     Organic Carbon (OC)               & Transportation Sector               \\
-    Sulfur Hexafluoride (SF6)         & Waste                               \\
+    Sulfur Hexafluoride (SF6)
+    \footnoteref{x_glb_all}               & Waste                               \\
     Sulfur Oxides (SOx)               &                                     \\
@@
   \end{tabular}
-\end{table}
+\end{minipage}\end{table}
 
 The effect of harmonization overrides is analyzed by harmonizing each scenario
-first without and then with overrides. Of the 1430 trajectories, approximately
-10\% were reported as a diagnostic (see Section \ref{sec:workflow}) of which 2\%
+first without and then with overrides. Of the 970 trajectories, approximately
+10\% were reported as a diagnostic (see Section \ref{sec:workflow}) of which 3.5\%
 required the use of harmonization overrides after an initial investigation;
-thus, 98\% of all trajectories were satisfactorily harmonized using the default
+thus, 96.5\% of all trajectories were satisfactorily harmonized using the default
 methods. Figure \ref{fig:nox} provides
@@
 override method is required to be provided for the trajectory of the region,
-sector, and species in question. The 2\% of trajectories that required overrides
+sector, and species in question. The 3.5\% of trajectories that required overrides
 clustered into two classifications: regional trajectories whose

### 2017-07-17 future.tex
fix counting of harmonized trajectories, added footnotes to table
consistent reference to aneris

@@
 methods for determining the use of harmonization methods was provided in Section
-\ref{sec:meths}. The \code{aneris} code base was able to satisfactorily
-harmonized over 98\% of the 2860 individual trajectories that were analyzed in
+\ref{sec:meths}. \code{aneris} was able to satisfactorily
+harmonized over 96\% of the 1940 individual trajectories that were analyzed in
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T23:07:25.083Z:
@@
 There are a variety of avenues for future improvement of both the \code{aneris}
-code base and underlying methodology. As with any software project, additional
+software and underlying methodology. As with any software project, additional
 users will provide use cases for more robust handling of input/output concerns
@@
 expert opinion and increase transparency and reproducibility of harmonization
-exercises. Furthermore, it provides an open-source, tested, and documented code
-base which can be used and improved upon by these communities. Both of these are
-clear steps in a positive direction for future climate and integrated assessment
-modeling exercises.
+exercises. Furthermore, it provides an open-source, tested, and documented
+software library which can be used and improved upon by these communities. Both
+of these are clear steps in a positive direction for future climate and
+integrated assessment modeling exercises.

### 2017-07-17 intro.tex
consistent reference to aneris

@@
 The remainder of this paper describes the methodology and implementation of the
-harmonization code base \code{aneris} \cite{matthew_gidden_2017_802832}, written in the Python
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:12:05.654Z:
-% 
-% > the
-% > \code{aneris} code base \cite{matthew_gidden_2017_802832
-% perhaps something like: "the harmonization code base \code{aneris}", to be explicit.
-% 
-% ^.
-programming language (detailed documentation available
+harmonization software \code{aneris} \cite{matthew_gidden_2017_802832}, written
+in the Python programming language (detailed documentation available
 \href{http://mattgidden.com/aneris/}{online}). Section \ref{sec:meths} provides

### 2017-07-17 meth.tex
decision tree as pdf
consistent reference to aneris

@@
 base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and those that operate on the
-offset of base year values (i.e., $h(t_i) - m(t_i)$). The \code{aneris} code
-base provides a number of implementations on the classic methods, including
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:27:39.375Z:
-% 
-% > classic methods
-% This might be interpreted as that convergence have not been previously applied. Perhaps use " implementations"  instead of "variations" .
-% 
-% ^.
-ratio-convergence methods shown in Equation \ref{eqs:ratio}, offset-convergence
-methods shown in Equation \ref{eqs:offset}, and interpolation methods shown in
-Equation \ref{eqs:interpolate}. The convergence factor, $\beta$, scales linearly
-from 1 to 0 in $[t_i, t_f)$ and is shown in Equation \ref{eqs:factor}. Note
-that region, species, and sector indices have been dropped for clarity. Each
-equation is a function of time, model trajectory, historical trajectory,
-harmonization time step, and a convergence time step ($t_f$), at which point the
-harmonized trajectory converges to the unharmonized trajectory. The
-\code{aneris} code base provides a number of methods to choose from for each of
-the harmonization families. A summary of all available methods is provided in
-Table \ref{tab:meths}.
+offset of base year values (i.e., $h(t_i) - m(t_i)$). \code{aneris} implements a
+number of the classic methods, including ratio-convergence shown in Equation
+\ref{eqs:ratio}, offset-convergence shown in Equation \ref{eqs:offset}, and
+interpolation shown in Equation \ref{eqs:interpolate}. The convergence factor,
+$\beta$, scales linearly from 1 to 0 in $[t_i, t_f)$ and is shown in Equation
+  \ref{eqs:factor}. Note that region, species, and sector indices have been
+  dropped for clarity. Each equation is a function of time, model trajectory,
+  historical trajectory, harmonization time step, and a convergence time step
+  ($t_f$), at which point the harmonized trajectory converges to the
+  unharmonized trajectory. \code{aneris} provides a number of methods to choose
+  from for each of the harmonization families. A summary of all available
+  methods is provided in Table \ref{tab:meths}.
 % * <maarten.vandenberg@pbl.nl> 2017-07-03T21:34:25.905Z:
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{decision_tree.png}
+    \includegraphics[width=\textwidth]{decision_tree.pdf}
     \caption[]{

### 2017-07-17 results.tex
add explanation of figure

@@
     \hline
+    \hline
     Black Carbon (BC)                 & Agricultural Waste Burning          \\
@@
 first without and then with overrides. Of the 970 trajectories, approximately
-10\% were reported as a diagnostic (see Section \ref{sec:workflow}) of which 3.5\%
-required the use of harmonization overrides after an initial investigation;
-thus, 96.5\% of all trajectories were satisfactorily harmonized using the default
-methods. Figure \ref{fig:nox} provides
-one such example of an emissions species and sector in which all regions were
-satisfactorily harmonized with the default methods. As can be seen, results with
-and without overrides are identical.
+10\% were reported as a diagnostic (see Section \ref{sec:workflow}) of which
+3.5\% required the use of harmonization overrides after an initial
+investigation; thus, 96.5\% of all trajectories were satisfactorily harmonized
+using the default methods. NOx generated from the Energy sector provides one
+such example of an emissions species and sector in which all regions were
+satisfactorily harmonized with the default methods. Figure \ref{fig:nox} shows
+the results of harmonization in Asia, and Table \ref{tab:nox} describes the
+parameters that underly the choice of method for each harmonized trajectory.
 
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{results_NOx_Energy_Sector.pdf}
+    \includegraphics[width=\textwidth]{example_NOx_Energy_Sector.pdf}
     \caption[]{
@@
       NOx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
-      trajectories for SSP2 and SSP2-45 are presented. SSP-45 is denoted with
-      ``x'' markers. Panel \textbf{a} shows 5-Region trajectories for scenarios
-      \textit{without} any overrides. Panel \textbf{b} shows 5-Region
-      trajectories for scenarios \textit{with} overrides. Panels \textbf{c} and
-      \textbf{d} show global trajectories without and with overrides,
-      respectively.  }
+      trajectories for SSP2 and SSP2-45 with historical trajectories (grey
+      lines) are presented. The SSP2 reference scenario is shown in Panels
+      \textbf{a} and \textbf{c}; the SSP-45 scenario is denoted with ``x''
+      markers in Panels \textbf{b} and \textbf{d}. The upper panels show the
+      results for endogenously modeled and harmonized regions in Asia while the
+      lower panels display the aggregate region results.  
+    }
   \end{center}
 \end{figure}
+
+\begin{table}[]
+\centering
+\caption{Key Parameters for Deciding Harmonization Methods for NOx Emissions in the Energy Sector in Asia}
+\label{tab:nox}
+\begin{tabular}{|p{1.2cm}|p{.5cm}|p{.5cm}|p{4.5cm}|p{3cm}|}
+\hline
+\textbf{Region} & \textbf{dH} & \textbf{c\_v} & \textbf{Decision Tree Traversal (Branch and Direction)} & \textbf{Method Chosen} \\ \hline
+    \hline
+CPA             & 0.35        & 2.26          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
+PAS             & 0.14        & 1.24          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
+SAS             & 0.56        & 0.58          & 1 (no), 2 (no), 3 (no), 4 (no)                          & constant\_ratio        \\ \hline
+\end{tabular}
+\end{table}
 

### 2017-07-18 intro.tex
final edits to intro

@@
 Integrated Assessment Models (IAMs) are tools used to understand the complex
-interactions between energy systems, economic systems, land use, the climate,
-and air pollution. IAMs provide global projections of
-systemic change by dividing the world into a number of representative regions
-(typically 10 to 30), the definition of which is distinct for each model
-\cite{krey_global_2014}. Results from IAMs are integral in a number of
-international studies, which notably include projections of climate and economic
-futures. Recently, the IAM community has developed scenarios based on the Shared
-% * <ssmith@pnnl.gov> 2017-07-01T03:11:26.621Z:
-% 
-% > flagship
-% Suggest different word. "comprehensive" is one possibility.
-% 
-% ^ <ssmith@pnnl.gov> 2017-07-01T03:11:57.571Z.
-Socioeconomic Pathways (SSPs) \cite{van_vuuren_energy_2017, fricko_marker_2017,
-  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
-quantify a variety of potential global futures. The SSPs are  designed
-to be used in  research that include earth system
-model (ESM) simulations, climate impact, adaptation and climate mitigation
-studies.
+interactions between energy, economy, land use, water, and climate systems. IAMs
+provide global projections of systemic change by dividing the world into a
+number of representative regions (typically 10 to 30), the definition of which
+is distinct for each model \cite{krey_global_2014}. Results from IAMs are
+integral in a number of international studies, which notably include projections
+of climate and economic futures. Recently, the IAM community has developed
+scenarios based on the Shared Socioeconomic Pathways (SSPs)
+\cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
+  calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which quantify a variety of
+potential global futures. The SSPs are designed to be used in research that
+include earth system model (ESM) simulations, climate impact, adaptation and
+climate mitigation studies.
 
@@
 emissions-intensity factors for individual gas species, for example the quantity
-of sulfur emissions from coal plants per megawatt-hour of production.
-Models are generally \textit{calibrated} to historical data sources in one or more base years.
-Due to the sometimes large uncertainties
-in historical datasets, results in the historical period may differ between
-models. Models also differ in their choice of base-year, which may lag behind available inventory data. In addition, models have varying sectoral, regional, and fuel aggregations.
+of sulfur emissions from coal plants per megawatt-hour of production. Models are
+generally \textit{calibrated} to historical data sources in one or more base
+years. Due to the sometimes large uncertainties in historical data sets, results
+in the historical period may differ between models. Models also differ in their
+choice of base-year, which may lag behind available inventory data. In addition,
+models have varying sectoral, regional, and fuel aggregations.
 
-The
-global climate change community has recently developed a new global historical
-emissions dataset for both anthropogenic  emissions (i.e., the Community
-Emissions Data System (CEDS) \cite{hoesly_historical_2017} and open-burning land-use and land-use change (LULUC)
-emissions \cite{van_marle_historic_2017}) which, in
-% * <sfujimori112256@gmail.com> 2017-06-07T09:59:35.820Z:
-% 
-% > land-use change (LUC) emissions \cite{van_marle_historic_2017}
-% I think this one also anthropogenic
-% 
-% ^ <maarten.vandenberg@pbl.nl> 2017-07-03T20:46:42.206Z:
-% 
-% I think so too, however I believe the GFED also contains natural emissions.
-%
-% ^.
-conjunction with the SSP IAM trajectories, will be used for climate-related
-modeling exercises of CMIP6.
+The global climate change community has recently developed a new global
+historical emissions data set for both anthropogenic emissions (i.e., the
+Community Emissions Data System (CEDS) \cite{hoesly_historical_2017} and
+open-burning land-use and land-use change (LULUC) emissions
+\cite{van_marle_historic_2017}) which, in conjunction with the SSP IAM
+trajectories, will be used for climate-related modeling exercises of CMIP6.
 
-When participating in intercomparison exercises in which a consistent historical starting point is required (e.g., in CMIP6), model teams incorporate a
-single, common historical dataset through \textit{harmonization}. Harmonization describes the process of
-adjusting model results to match the selected historical time series  such that the resulting future trajectories are also consistent with the
-original modeled results. In the emissions context, this means that each
-individual combination of model region, model sector, and emissions species must
-be harmonized. Depending on the total number of model regions, sectors, and
-emissions species, this can require the selection of thousands to
-tens-of-thousands of harmonization methods.
+When participating in intercomparison exercises in which a consistent historical
+starting point is required (e.g., in CMIP6), model teams incorporate a single,
+common historical data set through \textit{harmonization}. Harmonization
+describes the process of adjusting model results to match the selected
+historical time series such that the resulting future trajectories are also
+consistent with the original modeled results. In the emissions context, this
+means that each individual combination of model region, model sector, and
+emissions species must be harmonized. Depending on the total number of model
+regions, sectors, and emissions species, this can require the selection of
+thousands to tens-of-thousands of harmonization methods.
 
@@
 a detailed description of the underlying mathematical components of
-\code{aneris} as well as the code workflow. The results of applying the
+\code{aneris} as well as the procedural workflow. The results of applying the
 automated harmonization mechanism on two example IAM scenarios, one with

### 2017-07-18 meth.tex
final edits to methods

@@
 values, $h_{r, g, s}(t)$, and initial modeling period $t_i$, a harmonized
-trajectory needs to be calculated. The ``goodness of fit'' of a trajectory depends on
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:16:50.244Z:
-% 
-% Assessing the "goodness of fit" is not  straightforward, there may be competing factors. Perhaps we should refer back to this in the discussion. 
-% 
-% ^.
-a number of factors, most notably the faithful representation of the original,
-unharmonized, trajectory. However, other important factors must be considered,
-such as the representation of negative trajectories (i.e., if a trajectory
-becomes negative, both the timing and total magnitude should be as close as
-possible) which are of critical importance for cumulative CO2 calculations.
-% * <ssmith@pnnl.gov> 2017-07-01T03:25:04.357Z:
-% 
-% Suggest moving this later where negative emissions are discussed. Its a bit of a tangent here.
-% 
-% ^ <ssmith@pnnl.gov> 2017-07-01T03:25:44.627Z.
+trajectory needs to be calculated. The ``goodness of fit'' of a trajectory
+depends on a number of factors, most notably the faithful representation of the
+original, unharmonized, trajectory. However, other important factors must be
+considered, such as the representation of negative trajectories (i.e., if a
+trajectory becomes negative, both the timing and total magnitude should be as
+close as possible) which are of critical importance for cumulative \cotwo
+calculations.
 
@@
 interpolation shown in Equation \ref{eqs:interpolate}. The convergence factor,
-$\beta$, scales linearly from 1 to 0 in $[t_i, t_f)$ and is shown in Equation
-  \ref{eqs:factor}. Note that region, species, and sector indices have been
-  dropped for clarity. Each equation is a function of time, model trajectory,
-  historical trajectory, harmonization time step, and a convergence time step
-  ($t_f$), at which point the harmonized trajectory converges to the
-  unharmonized trajectory. \code{aneris} provides a number of methods to choose
-  from for each of the harmonization families. A summary of all available
-  methods is provided in Table \ref{tab:meths}.
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:34:25.905Z:
-% 
-% > A
-% > summary
-% If beta would be defined as \frac{t -t_i}{t_f - t_i}, it might better match intuition of convergence: 0 value at t = t_i and 1 at full convergence (t = t_f), with added benefit the beta parameter can also be used in the interpolation option.
-% 
-% ^.
+$\beta$, scales linearly from 1 to 0 over $[t_i, t_f)$ and is shown in Equation
+  \ref{eqs:factor}. In all equations the region, species, and sector indices
+  have been dropped for clarity. Each equation is a function of time, model
+  trajectory, historical trajectory, harmonization time step ($t_i$), and a
+  convergence time step ($t_f$), at which point the harmonized trajectory
+  converges to the unharmonized trajectory. \codeb{aneris} provides a number of
+  methods to choose from for each of the harmonization families. A summary of
+  all available methods is provided in Table \ref{tab:meths}.
 
@@
 for a given trajectory, a new, \textit{decision tree} approach has been
-implemented in this work which provides a systematic and documented
+implemented in \codeb{aneris} which provides a systematic and documented
 decision-making process for method determination. In order to provide reasonable
@@
       library. For all decisions, upper (purple) branches represent a ``yes''
-      response and lower (orange) branches represent a ``no'' response. $dH$ is
-      defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
+      response and lower (orange) branches represent a ``no'' response. The
+      coefficient of variation, $c_v$, is defined in Equation \ref{eqs:cov},
+      $dH$ is defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
       decision-making thresholds for $c_v$ and $dH$ are described below. Methods
@@
 species which can be extracted from the environment and stored, as is the case
-for CO2 in future scenarios with climate mitigation. If a model provides a
+for \cotwo in future scenarios with climate mitigation. If a model provides a
 trajectory that transitions from positive to negative emissions and base year
@@
       negative emissions by the end of the modeling time horizon (Panels
-      \code{c} and \code{d}). In each figure row (Panels \code{a}, \code{b};
-      \code{c}, \code{d}), identical model trajectories are used. In each figure
-      column, historical values are increased in the base year by an order of
-      magnitude (from 10 to 100). In each case, a subset of the potential
-      routines provide a better ``goodness of fit'' than others as described in
-      the text.  }
+      \code{c} and \code{d}). Identical model trajectories are used in each row
+      (Panels \code{a}, \code{b}; \code{c}, \code{d}). In each column,
+      historical values are increased in the base year by an order of magnitude
+      (from 10 to 100). In each Panel, a subset of the potential routines
+      provide a better ``goodness of fit'' than others as described in the text.
+    }
   \end{center}
@@
 
-Temporal variability of the historical trajectory is also an important characteristic
-when considering the choice of harmonization method.
-Emissions from forest and grassland fires, for example, vary from year to year due to a combination of meteorological conditions and anthropogenic drivers.
-Land use emissions are modeled by IAMs using average emission factors and do not capture conditions in a specific year.
-Ratio methods for such sectors can lead to unrealistic harmonization results.
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T21:59:54.238Z:
-% 
-% > Ratio methods for such sectors can lead to unrealistic harmonization results.
-% However, using for instance an offset method may also misrepresent the emission levels. Depending on the trajectory the emission factor can be subject to unintentional changes over the century.
-% 
-% ^.
-modeled with varying fidelity by IAMs, but in general proxies are used for
-emissions trajectories. Take for example emissions from grassland burning: for
-most models, the use of grass and paturelands is endogenously modeled, and
-emissions from the burning of these lands is determined by applying an emissions
-factor; in general, the more grassland usage that occurs in a scenario, the
-higher the emissions from the burning thereof. Therefore, consistency in
-harmonization method is important because the effects are modeled similarly
-across regions and species. In order to detect land-use-like emissions,
-% * <ssmith@pnnl.gov> 2017-07-01T03:31:52.910Z:
-% 
-% > modeled with varying fidelity by IAMs, but in general proxies are used for
-% > emissions trajectories. Take for example emissions from grassland burning: for
-% > most models, the use of grass and paturelands is endogenously modeled, and
-% > emissions from the burning of these lands is determined by applying an emissions
-% > factor; in general, the more grassland usage that occurs in a scenario, the
-% > higher the emissions from the burning thereof. Therefore, consistency in
-% > harmonization method is important because the effects are modeled similarly
-% > across regions and species. 
-% 
-% Not sure I follow the logic here. I added some new text that might help. Perhaps the intermediate part could be deleted.
-% 
-% ^ <ssmith@pnnl.gov> 2017-07-01T03:32:30.453Z.
-a measure of the coefficient of variation of the first derivative of the
-historical trajectory, $c_v$, is calculated as shown in Equation \ref{eqs:cov},
-where $\sigma$ is the standard deviation, and $\mu$ is the mean of the
-trajectory.
+Temporal variability of the historical trajectory is also an important
+characteristic when considering the choice of harmonization method.  Emissions
+from forest and grassland fires, for example, vary from year to year due to a
+combination of meteorological conditions and anthropogenic drivers. Land use
+emissions are modeled by IAMs using average emission factors and do not capture
+conditions in a specific year. A longer convergence horizon is thus desired in
+order to incorporate highly variable historical data with modeled results as is
+consistency in harmonization method because the effects are modeled similarly
+across regions and species. In order to detect land-use-like emissions, a
+measure of the coefficient of variation, $c_v$, of the first derivative of the
+historical trajectory is calculated as shown in Equation \ref{eqs:cov}, where
+$\sigma$ is the standard deviation, and $\mu$ is the mean.
 
@@
 \begin{equation}\label{eqs:cov}
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:08:21.230Z:
-% 
-% > \begin{equation}\label{eqs:cov}
-% >     c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
-% > \end{equation}
-% c_v is called a covariance parameter and a variance of derivatives, what is exactly calculated?
-% 
-% ^.
     c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
@@
       model regional definitions, and all gas species included in the historical
-      datasets are included in the analysis. The solid black line indicates the
-      threshold value used by default in \code{aneris}.  }
+      data sets are included in the analysis. The solid black line indicates the
+      threshold value, $\tau_{c_v}$, used by default in \code{aneris}.  }
   \end{center}
@@
 value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:25:47.452Z:
-% 
-% > 50\%
-% What does this mean? I would expect something of the form low_value < dH < high_value?
-% 
-% ^.
 
@@
       values in the harmonization year is shown. The solid black line indicates
-      the 50\% threshold value used by default in \code{aneris}.  
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:14:59.644Z:
-% 
-% > The distribution of relative differences between model and historical
-% >       values in the harmonization year is shown. The solid black line indicates
-% >       the 50\% threshold value used by defaul
-% the asymmetry here might be interesting, do the SSP scenarios consistently underestimate emission levels, or is this an effect of history up to 2015, while scenarios start earlier (might not be interesting for this paper).
-% 
-% ^.
+      the 50\% threshold value, $\tau_{dH}$, used by default in \code{aneris}.
     }
@@
 Input data then undergoes a cleaning operation, which adds (null) model
-trajectories that exist in the historical dataset but are not provided by the
+trajectories that exist in the historical data set but are not provided by the
 model input and detects any issues that would cause the harmonization process to
@@
 \hline
-Column       & Description                                                                  \\
+\textbf{Column}       & \textbf{Description}                                      \\
 \hline

### 2017-07-18 results.tex
substantive edits to results

@@
 In order to test the \code{aneris} harmonization procedure, results from the IAM
-MESSAGE-GLOBIOM \cite{fricko_marker_2017} was processed. Two scenarios from the
+MESSAGE-GLOBIOM \cite{fricko_marker_2017} were processed. Two scenarios from the
 SSP scenario library \cite{Riahi2017153,Rao2017346} are investigated. The SSP2,
@@
 total of 970 distinct trajectories were harmonized for each scenario, and
-therefore 1940 trajectories were harmonized in total.
+therefore 1940 trajectories were harmonized in total. NOx generated from the
+Energy sector provides an example of an emissions species and sector in which
+all regions were satisfactorily harmonized with the default methods. Figure
+\ref{fig:nox} shows the results of harmonization in Asia, and Table
+\ref{tab:nox} describes the parameters that underlie the choice of method for
+each harmonized trajectory.
 
@@
 
-The effect of harmonization overrides is analyzed by harmonizing each scenario
-first without and then with overrides. Of the 970 trajectories, approximately
-10\% were reported as a diagnostic (see Section \ref{sec:workflow}) of which
-3.5\% required the use of harmonization overrides after an initial
-investigation; thus, 96.5\% of all trajectories were satisfactorily harmonized
-using the default methods. NOx generated from the Energy sector provides one
-such example of an emissions species and sector in which all regions were
-satisfactorily harmonized with the default methods. Figure \ref{fig:nox} shows
-the results of harmonization in Asia, and Table \ref{tab:nox} describes the
-parameters that underly the choice of method for each harmonized trajectory.
-
 \begin{figure}
@@
 The harmonization of emissions pathways is performed in order to accurately
-represent new or updated datasets of historical emissions inventories while also
+represent new or updated data sets of historical emissions inventories while also
 maintaining consistency with the original, unharmonized pathway. As such, when
@@
 override method is required to be provided for the trajectory of the region,
-sector, and species in question. The 3.5\% of trajectories that required overrides
-clustered into two classifications: regional trajectories whose
-\textit{magnitude} was overly distorted and regional trajectories whose
-\textit{shape} was overly distorted.
+sector, and species in question. Of the 970 trajectories, approximately 10\%
+were reported as a diagnostic (see Section \ref{sec:workflow}) of which 3.5\%
+required the use of harmonization overrides after an initial investigation;
+thus, 96.5\% of all trajectories were satisfactorily harmonized using the
+default methods. The trajectories that required overrides clustered into two
+classifications: regional trajectories whose \textit{magnitude} was overly
+distorted and regional trajectories whose \textit{shape} was overly distorted.
 
@@
 sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region, which
-comprises most of the Asian subcontinent. The default method chosen
-(\code{constant_ratio}) maintains model trends for the region; however, overall
-model results are distorted. By applying a \code{constant_offset} override, the
-regional trend and magnitude is maintained. With the new harmonization method
-for the SAS region, the global trajectory for industrial CO also is
-representative the trends seen in the unharmonized trajectory and the relative
-importance of the underlying regional trajectoreis is maintained.
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T22:53:14.757Z:
-% 
-% > the global trajectory for industrial CO also corresponds
-% > more closely with the unharmonized trajectory.
-% The global trajectory suffers under the constant_ratio method from changing relative importance of underlying regions, resulting in a trend change, this effect is also prevented by using a constant_offset method.
-% 
-% ^.
+comprises most of the emissions of the Asian subcontinent. The default method
+chosen (\code{constant_ratio}) maintains model trends for the region; however,
+overall model results are distorted. By applying a \code{constant_offset}
+override, the regional trend and magnitude is maintained. With the new
+harmonization method for the SAS region, the global trajectory for industrial CO
+also is representative the trends seen in the unharmonized trajectory and the
+relative importance of the underlying regional trajectories is maintained.
 

### 2017-07-18 results.tex
add gas subscripts

@@
 total of 970 distinct trajectories were harmonized for each scenario, and
-therefore 1940 trajectories were harmonized in total. NOx generated from the
+therefore 1940 trajectories were harmonized in total. \noxx generated from the
 Energy sector provides an example of an emissions species and sector in which
@@
     Black Carbon (BC)                 & Agricultural Waste Burning          \\
-    Hexafluoroethane (C2F6)
+    Hexafluoroethane (C\textsubscript{2}F\textsubscript{6})
     \footnote{\label{x_glb_all}
@@
                                       & Agriculture                         \\
-    Tetrafluoromethane (CF4)
+    Tetrafluoromethane (CF\textsubscript{4})
     \footnoteref{x_glb_all}               & Aircraft
@@
       historical data.}                            \\
-    Methane (CH4)                     & Energy Sector                       \\
-    Carbon Dioxide (CO2)
+    Methane (CH\textsubscript{4})                     & Energy Sector                       \\
+    Carbon Dioxide (CO\textsubscript{2})
     \footnote{\label{x_co2}
-      A global trajectory for land-use CO2 is used; non-land-use sectors 
+      A global trajectory for land-use CO\textsubscript{2} is used; non-land-use sectors 
       are harmonized for each model region.}
@@
     \footnoteref{x_glb_all}               & Industrial Sector                   \\
-    Nitrous Oxide (N2O)
+    Nitrous Oxide (N\textsubscript{2}O)
     \footnoteref{x_glb_all}
                                       & International Shipping\footnoteref{x_glb}              \\
-    Ammonia (NH3)                     & Residential Commercial Other        \\
-    Nitrogen Oxides (NOx)             & Solvents Production and Application \\
+    Ammonia (\nht)                     & Residential Commercial Other        \\
+    Nitrogen Oxides (\nox)             & Solvents Production and Application \\
     Organic Carbon (OC)               & Transportation Sector               \\
-    Sulfur Hexafluoride (SF6)
+    Sulfur Hexafluoride (SF\textsubscript{6})
     \footnoteref{x_glb_all}               & Waste                               \\
-    Sulfur Oxides (SOx)               &                                     \\
+    Sulfur Oxides (SO\textsubscript{x})               &                                     \\
     Volatile Organic Compounds (VOCs) &                                     \\
@@
       \label{fig:nox}
-      NOx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
+      \noxx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
       trajectories for SSP2 and SSP2-45 with historical trajectories (grey
@@
 \centering
-\caption{Key Parameters for Deciding Harmonization Methods for NOx Emissions in the Energy Sector in Asia}
+\caption{Key Parameters for Deciding Harmonization Methods for \noxx Emissions in the Energy Sector in Asia}
 \label{tab:nox}
@@
 trajectories. Figure \ref{fig:nh3} shows an example case of emissions
-trajectories for ammonia (NH3) from the agriculture sector in Asia. Again, the
+trajectories for ammonia (\nht) from the agriculture sector in Asia. Again, the
 SAS region shows a large discrepancy in the harmonization year ($>$150\% in
@@
       \label{fig:nh3}
-      NH3 agricultural harmonized and unharmonized emissions are presented for
+      \nhtt agricultural harmonized and unharmonized emissions are presented for
       SSP2 and SSP2-45 scenarios. Scenarios and panel layouts are identical to

### 2017-07-18 results.tex
final fixes to results

@@
 \hline
-\textbf{Region} & \textbf{dH} & \textbf{c\_v} & \textbf{Decision Tree Traversal (Branch and Direction)} & \textbf{Method Chosen} \\ \hline
+\textbf{Region} & \textbf{dH} & $\mathbf{c_v}$ & \textbf{Decision Tree Traversal (Branch and Direction)} & \textbf{Default Method Chosen} \\ \hline
     \hline
@@
 % > degree.
-% First, I think other effects are in play here.
-The unharmonized SSP2 -ref scenario has a discontinuity in it's derivative for the SAS region.
-This is amplified at the regional level by the constant_ratio approach and affects the 4.5 scenario in the same manner, that just happens to lack the discontinuity.
-Furthermore the trend of other regions is strengthened (a stronger decrease), leading up to an even stronger discontinuity in the global total.
-However, the ratio approach does a better job at representing the trend in the scenario.
-That trade-off is not completely clear now, also refers back to goodness of fit.
-Secondly, more problematic effects occur when applying reduce_ratio methods, that can lead to strong magnitude and trend changes; however this should only occur (significantly) when in override mode (if I understand the implementation of dH correctly).
+% First, I think other effects are in play here. The unharmonized SSP2 -ref
+% scenario has a discontinuity in it's derivative for the SAS region. This is
+% amplified at the regional level by the constant_ratio approach and affects the
+% 4.5 scenario in the same manner, that just happens to lack the
+% discontinuity. Furthermore the trend of other regions is strengthened (a
+% stronger decrease), leading up to an even stronger discontinuity in the global
+% total. However, the ratio approach does a better job at representing the trend
+% in the scenario. That trade-off is not completely clear now, also refers back
+% to goodness of fit. Secondly, more problematic effects occur when applying
+% reduce_ratio methods, that can lead to strong magnitude and trend changes;
+% however this should only occur (significantly) when in override mode (if I
+% understand the implementation of dH correctly).
 % 
@@
 override, this distortion is addressed and more accurately reflects unharmonized
-results both in the SAS region and global results for agricultural ammonia
-emissions.
+results in the SAS region, the relative importance between regions, and global
+results for agricultural ammonia emissions, each of which contributes to a
+better ``goodness of fit'' for the harmonized SAS trajectory.
 
@@
       trajectory. Additionally, global trajectories are greatly affected by the
-      harmonization method choice (there is $\sim$20 percent relative difference
+      harmonization method choice (there is $\sim$20\% relative difference
       between trajectories in the reference scenario in 2100). Override methods

### 2017-07-18 future.tex
results review

@@
 This work presented a novel methodology and Python implementation of automated
-emissions harmonization for IAMs. An in-depth explanation of the processes and
-methods for determining the use of harmonization methods was provided in Section
-\ref{sec:meths}. \code{aneris} was able to satisfactorily
+emissions harmonization for IAMs, \code{aneris}. An in-depth explanation of the
+processes and methods for determining the use of harmonization methods was
+provided in Section \ref{sec:meths}. \code{aneris} was able to satisfactorily
 harmonized over 96\% of the 1940 individual trajectories that were analyzed in
@@
 users will provide use cases for more robust handling of input/output concerns
-and corner cases. Further configuration parameters may also be added in the
+and corner cases. Further configuration parameters can also be added in the
 future in order to provide overrides for all gas species in a given sector or
@@
 
-However, this work provides a new direction and framework which the IAM and
-climate communities can build upon in order to reduce the necessity of ad-hoc
-expert opinion and increase transparency and reproducibility of harmonization
+This work provides a new direction and framework which the IAM and climate
+communities can build upon in order to reduce the necessity of consistent expert
+opinion and increase transparency and reproducibility of harmonization
 exercises. Furthermore, it provides an open-source, tested, and documented

### 2017-07-18 elsfrontmatter.tex
proposed author order
acks, key words, nies address

@@
 
+\author[nies]{Shinichiro Fujimori}
 \author[pbl]{Maarten van den Berg}
-\author[nies]{Shinichiro Fujimori}
 \author[pik]{David Klein}
+\author[pnnl]{Steven J. Smith}
 \author[iiasa]{Keywan Riahi}
-\author[pnnl]{Steven J. Smith}
-\author[pbl]{Detlef van Vuuren}
 
@@
   Schlossplatz 1, A-2361 Laxenburg, Austria}
-\address[nies]{National Institute for Environmental Studies, Tsukuba, Japan}
+\address[nies]{Center for Social and Environmental Systems Research, National Institute for Environmental Studies, 16-2 Onogawa, Tsukuba, Ibaraki 305-8506, Japan}
 \address[pbl]{PBL Netherlands Environmental Assessment Agency, Postbus 30314, 2500 GH The Hague, Netherlands}
@@
 methodology for determining such harmonization methods and an open-source Python
-code base for implementing the methodology. Results are shown for two example
-scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 96\%
-of the total emissions trajectories.
+software library implementing the methodology. Results are shown for two example
+scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM
+that satisfactorily harmonize over 96\% of the total emissions trajectories.
 \end{abstract}
@@
 \begin{keyword}
-Integrated Assessment Modeling, Harmonization, Climate Change, Air Pollution 
+Integrated Assessment Models, Climate Change, Harmonization, Air Pollution 
 \end{keyword}

### 2017-07-18 correspondance/cover.tex
adding cover letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}
 \begin{document}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Professor D.P. Ames\\
Editor-in-Chief, Environmental Modelling \& Software}

%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\address{Matthew Gidden\\
Laimgrubengasse 17/7\\
Vienna, Austria 1060}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Professor Ames,}

Please find enclosed a manuscript entitled: ``A Methodology and Implementation
of Automated Emissions Harmonization for Use in Integrated Assessment Models''
which I am submitting for exclusive consideration of publication as a
full-length article in Environmental Modelling \& Software.  This manuscript
describes novel work in the subject area of Integrated Assessment Models (IAMs)
and Modelling, specifically related to their use in combination with Earth
Systems Models (ESMs). A new software library, \texttt{aneris}, is presented
which robustly automates the process of emissions trajectory
harmonization. Through harmonization, multiple IAMs can produce results
consistent with a single historical data set.

This capability is critical for the success of the worldwide scientific
community efforts in the Coupled Model Intercomparison Project Phase 6 (CMIP6)
in which results from IAMs are being used as inputs to ESMs in order to model
future scenarios of climate change. By producing a tested, documented, and open
source software implementation, this work

\begin{itemize}
  \item provides a mechanism by which IAMs contributing to the current process
    of CMIP6 can produce consistent results not only within a single team but
    among all teams
  \item allows for other teams in the scientific community to contribute
    consistent results to the CMIP6 process in the future
  \item and increases the transparency and scientfic integrity of that process.
\end{itemize}

Thank you for your consideration of this work. Please address all correspondence
concerning this manuscript to me and feel free to correspond with me by e-mail.

\closing{
Warm Regards,

Matthew Gidden, Ph.D.
}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}

### 2017-07-18 paper.tex
outlined sections
starting to work on methods
updating paper and meth
intro added
Update on Overleaf.
23, May from SF
Shinichiro_30May
acks, key words, nies address
add software availability

@@
 %%% ELS includes
+% * <sfujimori112256@gmail.com> 2017-05-23T14:15:19.404Z:
+% 
+% comment from SF
+% 
+% ^ <sfujimori112256@gmail.com> 2017-05-24T18:49:42.852Z.
 \documentclass[review]{elsarticle}
@@
 %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\section*{Software Availability}
+
+\texttt{aneris}, first made available in 2017, is available online at
+\url{https://github.com/gidden/aneris} as a free and open-source Python software
+library (approximately 2000 lines of code). The \texttt{aneris} software was
+developed by the lead author, Matthew J. Gidden, Ph.D., whose contact
+information is shown on the title page of this manuscript. Documentation for the
+\texttt{aneris} Python package, including software requirements, is available
+\href{http://mattgidden.com/aneris/}{online}.
 
 \newpage
-stuff
-%% \input{intro}
-%% \input{methods}
-%% \input{results}
-%% \input{conclusion}
+
+\include{intro}
+\include{meth}
+\include{results}
+% * <sfujimori112256@gmail.com> 2017-05-25T05:18:35.016Z:
 % 
+% I think the current results are basically fine (sorry I have not read in detailed though).
+What we should show and discuss are pretty much depending on how we evaluate these results.
+So, at this stage, I would like to put more comments on this section
+% 
+% 
+% ^.
+\include{future}
 
@@
 
-fill in
+The authors would like to acknowledge a number of colleagues who helped
+contribute both discussion and feedback for this work, including Drs. Detlef van
+Vuuren, Elmar Kriegler, Gunnar Luderer, and Joeri Rogelj. The authors further
+with to thank \TODO{who?} for their generous funding of this effort.
 

### 2017-07-19 paper.tex
add detlef as author
clarify acknolwedments

@@
 The authors would like to acknowledge a number of colleagues who helped
-contribute both discussion and feedback for this work, including Drs. Detlef van
-Vuuren, Elmar Kriegler, Gunnar Luderer, and Joeri Rogelj. The authors further
-with to thank \TODO{who?} for their generous funding of this effort.
+contribute both discussion and feedback for this work, including Drs. Elmar
+Kriegler, Gunnar Luderer, and Joeri Rogelj. The authors further with to thank
+the Global Environment Facility for their generous support. This project has
+received funding from the European Union’s Horizon 2020 research and innovation
+programme under grant agreement No 641816.
 

### 2017-07-19 paper.tex
typo in acks

@@
 The authors would like to acknowledge a number of colleagues who helped
-contribute both discussion and feedback for this work, including Drs. Elmar
-Kriegler, Gunnar Luderer, and Joeri Rogelj. The authors further with to thank
+contribute both discussion and feedback regarding this work including Drs. Elmar
+Kriegler, Gunnar Luderer, and Joeri Rogelj. The authors further wish to thank
 the Global Environment Facility for their generous support. This project has

### 2017-07-24 future.tex
intro edits
finished dvv edits in intro

typo in intro

methods revisions

typo in texit

some final cleaning of meths

text edits in results

added figure for rad forcing

fix acks

updating forcing fig and text

clearing up some results text

rm wm cmd

@@
 The use of an open-source, automated harmonization process also provides
-benefits to the wider climate science and IAM communities. For example, other
-modeling teams are easily capable of executing identical harmonization methods
-in order to participate in ongoing and further iterations of intercomparison
-exercises and analysis. Future scenario analyses can also utilize this common
-harmonization approach such that they are consistent with prior efforts.
+benefits to the wider climate science and IAM communities. By providing a
+standard mechanism for harmonization, the IAM community can directly provide
+input into the harmonization algorithms and rules for their default
+selection. Additionally, modeling teams are easily capable of executing
+identical harmonization procedures in order to participate in ongoing and
+further iterations of intercomparison exercises and analysis. Future scenario
+analyses can also utilize this common harmonization approach such that they are
+consistent with prior efforts.
 

### 2017-07-24 intro.tex
intro edits
finished dvv edits in intro

typo in intro

methods revisions

typo in texit

some final cleaning of meths

text edits in results

added figure for rad forcing

fix acks

updating forcing fig and text

clearing up some results text

rm wm cmd

@@
 include earth system model (ESM) simulations, climate impact, adaptation and
-climate mitigation studies.
+climate mitigation studies \cite{vuuren_new_2013}.
 
@@
 optimization, the core inputs and outputs are similar across different
-models. Modeling teams incorporate  data on energy systems, land
-use, economics, demographics and emissions sources and concentrations, among other
-data, in order to provide consistent existing trajectories of modeled
-variables. The models then provide estimates of future trajectories of these
-variables under various socioeconomic and technological assumptions as well as
-proposed policy constraints, e.g., targets for future Greenhouse Gas (GHG)
-emissions.
+models. Modeling teams incorporate data on energy systems, land use, economics,
+demographics and emissions sources and concentrations, among other data, in
+order to provide consistent existing trajectories of modeled variables. The
+models then provide estimates of future trajectories of these variables under
+various socioeconomic and technological assumptions as well as proposed policy
+constraints, e.g., targets for future Greenhouse Gas (GHG) emissions.
 
@@
 Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
-of marker SSP scenarios developed by the IAM community
+of marker SSP scenarios developed by the IAM community (ScenarioMIP)
 \cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
@@
 emissions of different air pollutants and GHGs as well as various mitigation
-technologies. However, the historical emissions starting points of  models can differ by
-large amounts depending on the region, sector, and emissions species.
+technologies. However, the historical emissions starting points of models can
+differ by large amounts depending on the region, sector, and emissions species.
 
@@
 generally \textit{calibrated} to historical data sources in one or more base
-years. Due to the sometimes large uncertainties in historical data sets, results
-in the historical period may differ between models. Models also differ in their
-choice of base-year, which may lag behind available inventory data. In addition,
-models have varying sectoral, regional, and fuel aggregations.
+years. Results in the historical period may differ between models as a result of
+the sometimes large uncertainties in historical data sets. Other sources of
+error arise because models can also differ in their choice of base-year, which
+may lag behind available inventory data. In addition, models have varying
+sectoral, regional, and fuel aggregations.
 
@@
 common historical data set through \textit{harmonization}. Harmonization
-describes the process of adjusting model results to match the selected
+refers to the process of adjusting model results to match a selected
 historical time series such that the resulting future trajectories are also
@@
 \cite{vuuren_representative_2011}. Importantly, the choice of harmonization
-method to date has been determined via expert solicitation and has generally
-been applied to all trajectories for a given class of emissions species.
+method to date has been determined by individual experts and has generally been
+applied to all trajectories for a given class of emissions species.
 
@@
 scientifically rigorous harmonization procedure, an automated process for
-determining harmonization methods is needed. The use of an automated,
+determining harmonization methods is preferred. The use of an automated,
 documented, and openly available harmonization mechanism additionally allows for

### 2017-07-24 meth.tex
intro edits
finished dvv edits in intro

typo in intro

methods revisions

typo in texit

some final cleaning of meths

text edits in results

added figure for rad forcing

fix acks

updating forcing fig and text

clearing up some results text

rm wm cmd

@@
 initial modeling period. Given a model trajectory, $m_{r, g, s}(t)$, historical
-values, $h_{r, g, s}(t)$, and initial modeling period $t_i$, a harmonized
-trajectory needs to be calculated. The ``goodness of fit'' of a trajectory
-depends on a number of factors, most notably the faithful representation of the
-original, unharmonized, trajectory. However, other important factors must be
-considered, such as the representation of negative trajectories (i.e., if a
-trajectory becomes negative, both the timing and total magnitude should be as
+values, $h_{r, g, s}(t)$, and model base year, $t_i$, a harmonized trajectory
+needs to be calculated. The \textit{harmonization quality} of a trajectory,
+i.e., how well a given harmonization algorithm performs, depends on a number of
+factors, most notably the faithful representation of the original, unharmonized,
+trajectory. This includes the representation of negative trajectories (i.e., if
+a trajectory becomes negative, both the timing and total magnitude should be as
 close as possible) which are of critical importance for cumulative \cotwo
@@
 base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and those that operate on the
-offset of base year values (i.e., $h(t_i) - m(t_i)$). \code{aneris} implements a
-number of the classic methods, including ratio-convergence shown in Equation
-\ref{eqs:ratio}, offset-convergence shown in Equation \ref{eqs:offset}, and
-interpolation shown in Equation \ref{eqs:interpolate}. The convergence factor,
-$\beta$, scales linearly from 1 to 0 over $[t_i, t_f)$ and is shown in Equation
-  \ref{eqs:factor}. In all equations the region, species, and sector indices
-  have been dropped for clarity. Each equation is a function of time, model
-  trajectory, historical trajectory, harmonization time step ($t_i$), and a
-  convergence time step ($t_f$), at which point the harmonized trajectory
-  converges to the unharmonized trajectory. \codeb{aneris} provides a number of
-  methods to choose from for each of the harmonization families. A summary of
-  all available methods is provided in Table \ref{tab:meths}.
+offset of base year values (i.e., $h(t_i) - m(t_i)$). A number of the classic
+methods are implemented in \code{aneris} including ratio-convergence shown in
+Equation \ref{eqs:ratio}, offset-convergence shown in Equation \ref{eqs:offset},
+and interpolation shown in Equation \ref{eqs:interpolate}. The convergence
+factor, $\beta$, scales linearly from 1 to 0 over $[t_i, t_f)$ and is shown in
+  Equation \ref{eqs:factor}. In all equations the region, species, and sector
+  indices have been dropped for clarity. Each equation is a function of time,
+  model trajectory, historical trajectory, base year ($t_i$), and a convergence
+  year ($t_f$), at which point the harmonized trajectory converges to the
+  unharmonized trajectory. \codeb{aneris} provides a number of methods to choose
+  from for each of the harmonization families. A summary of all available
+  methods is provided in Table \ref{tab:meths}.
 
@@
 
-While historically expert opinion has guided the choice of harmonization methods
-for a given trajectory, a new, \textit{decision tree} approach has been
-implemented in \codeb{aneris} which provides a systematic and documented
-decision-making process for method determination. In order to provide reasonable
+A \textit{decision tree} approach has been implemented in \codeb{aneris} which
+provides a systematic and documented decision-making process to determine the
+preferred harmonization algorithm. In order to provide reasonable
 \textit{default} methods, the historical trajectory, unharmonized model
 trajectory, and relative difference between history and model values in the
-harmonization year are analyzed. The decision tree used in this analysis is
-shown graphically in Figure \ref{fig:decision_tree}.
+harmonization year are analyzed. The decision tree used in this analysis is a
+result of collaborative efforts between IAM teams and is shown graphically in
+Figure \ref{fig:decision_tree}.
 
@@
       (from 10 to 100). In each Panel, a subset of the potential routines
-      provide a better ``goodness of fit'' than others as described in the text.
+      provide a better harmonization quality than others as described in the
+      text.
     }
@@
 combination of meteorological conditions and anthropogenic drivers. Land use
-emissions are modeled by IAMs using average emission factors and do not capture
-conditions in a specific year. A longer convergence horizon is thus desired in
-order to incorporate highly variable historical data with modeled results as is
-consistency in harmonization method because the effects are modeled similarly
-across regions and species. In order to detect land-use-like emissions, a
-measure of the coefficient of variation, $c_v$, of the first derivative of the
-historical trajectory is calculated as shown in Equation \ref{eqs:cov}, where
-$\sigma$ is the standard deviation, and $\mu$ is the mean.
+emissions in many IAMs are modeled using average emission factors and do not
+capture conditions in a specific year. A longer convergence horizon is thus
+desired in order to incorporate highly variable historical data with modeled
+results as is consistency in harmonization method because the effects are
+modeled similarly across regions and species. In order to detect emissions with
+a high amount of variation, a measure of the coefficient of variation, $c_v$, of
+the first derivative of the historical trajectory is calculated using the
+standard deviation, $\sigma$, and the mean, $\mu$, as shown in Equation
+\ref{eqs:cov}.
 
@@
 run-control configuration are read in via an Excel spreadsheet. Data is assumed
-to be in the \textit{IAM Consortium} format, i.e., using \code{Model},
-\code{Scenario}, \code{Region}, \code{Variable}, and \code{Unit} columns in
-addition to columns representing each modeled time period.
+to be in the IAMC format, i.e., using \code{Model}, \code{Scenario},
+\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
+representing each modeled time period.
 
@@
 is a significant difference between historical and model values in the
-harmonization year and there is significant upward or downward movement in the
-model trajectory. In such cases, override methods can ameliorate the issues
-associated with the default method choice. In order to help identify these
-cases, harmonization \textit{diagnostics} are provided which analyze the
+harmonization year, there is significant upward or downward movement in the
+model trajectory, or if there are known discrepancies in definition between the
+IAM and historical data source. In such cases, override methods can ameliorate
+the issues associated with the default method choice. In order to help identify
+these cases, harmonization \textit{diagnostics} are provided which analyze the
 relative difference between harmonized and unharmonized trajectories at their
@@
 aggregation of common analysis regions is performed. A common regional
-aggregation used in the IAM community is the RCP regional definition, as shown
-in Figure \ref{fig:regions}. Finally, any exogenous trajectories the user provides
-are added. Exogenous trajectories are normally provided for unmodeled gases with
+aggregation used in the IAM community was defined in the Representative
+Concentration Pathways (RCPs) \cite{vuuren_representative_2011}, shown in Figure
+\ref{fig:regions}. Finally, any exogenous trajectories the user provides are
+added. Exogenous trajectories are normally provided for unmodeled gases with
 well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO

### 2017-07-24 results.tex
intro edits
finished dvv edits in intro

typo in intro

methods revisions

typo in texit

some final cleaning of meths

text edits in results

added figure for rad forcing

fix acks

updating forcing fig and text

clearing up some results text

rm wm cmd

@@
 
-In order to test the \code{aneris} harmonization procedure, results from the IAM
-MESSAGE-GLOBIOM \cite{fricko_marker_2017} were processed. Two scenarios from the
-SSP scenario library \cite{Riahi2017153,Rao2017346} are investigated. The SSP2,
-or ``middle of the road'', scenario family is chosen to be analyzed because
-MESSAGE-GLOBIOM is the marker scenario for this SSP. Of all SSP2 scenarios, the
-reference (marker) and SSP2-45 scenarios are analyzed. Notably, emissions from
-multiple sectors tend to increase in the reference scenario, thus testing the
-harmonization mechanism with an increasing emissions pathway. The SSP2-45
-scenario, where the ``45'' designation refers to a scenario with end-of-century
-emissions resulting in approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$
-radiative forcing due to both warming from GHGs and fluoridated gases as well as
-cooling from aerosols, is chosen because mitigation technologies and policies
-are enacted causing a general reduction in pollutants and GHGs, including
-(eventual) negative CO2 emissions in some regions and sectors due to carbon
-capture and sequestration and afforestation. This scenario explores the
-harmonization mechanism's response to generally decreasing emission
-trajectories. The emissions of Kyoto Gases, shown in Figure \ref{fig:kyoto},
-shows the different trends of these two scenarios.
+In order to show a representative cross section of the performance of the
+\code{aneris} harmonization procedure, we focus on the harmonization of results
+of the IAM MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the SSP
+scenario library \cite{Riahi2017153,Rao2017346} are presented. The SSP2, or
+``middle of the road'', scenario (referred to as SSP2-Ref) is chosen to be
+analyzed because MESSAGE-GLOBIOM is the marker scenario for this SSP. We
+additionally present the results for the SSP2-based mitigation scenario leading
+to a radiative forcing of 4.5 $\frac{\text{W}}{\text{m}^2}$ (referred to as
+SSP2-4.5). The SSP2-45 scenario is chosen because mitigation technologies and
+policies are enacted causing a general reduction in pollutants and GHGs,
+including (eventual) negative CO2 emissions in some regions and sectors due to
+carbon capture and sequestration and afforestation. Figure \ref{fig:kyoto} shows
+the different trends of Kyoto Gases in each scenarios.
 
@@
 results for agricultural ammonia emissions, each of which contributes to a
-better ``goodness of fit'' for the harmonized SAS trajectory.
+better harmonization quality for the harmonized SAS trajectory.
 
@@
 \end{figure}
+
+We investigate the aggregate effect of harmonization with all necessary override
+methods on total anthropogenic radiative forcing projections with the simple
+carbon-cycle and climate model, MAGICC6 \cite{meinshausen2011emulating,
+  meinshausen2011rcp}, for each harmonized and unharmonized scenario
+respectively as shown in Figure \ref{fig:forcing}. We find that the change due
+to harmonization is small, ranging between 1 and 2.5\% over the model
+horizon. Relative near-term differences persist in the mitigation case (SSP-4.5)
+because differences in near-term emissions define to a larger degree the
+longer-term forcing outcome due to the cumulative nature of long-lived climate
+forcers like \cotwo.  The resulting difference in forcing in 2100 is 0.04
+$\frac{\text{W}}{\text{m}^2}$ for SSP2-4.5 and 0.01
+$\frac{\text{W}}{\text{m}^2}$ for SSP2-Ref, both of which are well within
+acceptable tolerances (e.g., 0.75 $\frac{\text{W}}{\text{m}^2}$ defined for
+ScenarioMIP \cite{oneill_scenario_2016}). Thus harmonization is considered to
+have a negligible effect on key long-term climate indicators.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_forcing.pdf}
+    \caption[]{
+      \label{fig:forcing}
+      The results of the simple climate model, MAGICC6, forced with the SSP2-Ref
+      (blue) and SSP2-4.5 (green) harmonized and unharmonized scenarios is
+      presented. The radiative forcing trajectories of harmonized and
+      unharmonized scenarios are shown in solid lines and dashed lines,
+      respectively.  }
+  \end{center}
+\end{figure}

### 2017-07-25 elsfrontmatter.tex
add detlef as author
edits to abstrct

@@
 \author[pnnl]{Steven J. Smith}
+\author[pbl]{Detlef P. van Vuuren}
 \author[iiasa]{Keywan Riahi}
@@
 \begin{abstract}
-Harmonization describes the process required to match Integrated Assessment
-Model (IAM) GHG and air pollutants results against a common source of historical
-emissions trajectories. To date, harmonization has been performed separately by
-individual modeling teams through
-% * <maarten.vandenberg@pbl.nl> 2017-07-03T20:19:43.403Z:
-% 
-% IMAGE has provided harmonized emissions in the past, generated by a harmonization routine applying a ratio convergence method (on a higher aggregation level) (Hof et al., 2016).
-I think the contrast with these harmonization approaches and this work, which enables a documented, automated process selecting appropriate harmonization methodologies, might not be clear.
-Furthermore, later in the text cross-model harmonization methods (though less sophisticated) are mentioned.
-It is worth mentioning that previous approaches are far less sophisticated.
-% 
-% ^.
-% * <sfujimori112256@gmail.com> 2017-05-24T19:26:26.142Z:
-% 
-% > harmonization has been performed separately by individual modeling teams 
-% I think harmonization is not done. We just report the direct model outputs.
-% 
-% One of the ways to  make the background is that earlier IPCC scenarios (SRES and RCPs) did not document the harmonization method and far from transparent.
-% 
-% ^.
-the use of expert opinion and solicitation. In order to support ongoing and
-future global climate and socioeconomic analyses, incorporate model advancements
-towards more complex regional and sectoral representations, and expand the
-capability of modeling teams to participate in intercomparison exercises, an
-automated approach to choosing harmonization methods becomes necessary. This
-work describes a novel
-% * <sfujimori112256@gmail.com> 2017-05-24T20:13:08.493Z:
-% 
-% > novel
-% It would be better to stress how novel and useful.
-% 
-% ^.
-methodology for determining such harmonization methods and an open-source Python
-software library implementing the methodology. Results are shown for two example
-scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM
-that satisfactorily harmonize over 96\% of the total emissions trajectories.
+Emission harmonization refers to the process used to match results from
+Integrated Assessment Model (IAM) GHG and air pollutants against a common source
+of historical emissions trajectories. To date, harmonization has been performed
+separately by individual modeling teams. For the hand-over of emission data for
+the Shared Socio-economic Pathways (SSPs) to climate model groups, a new
+automated automatized approach was developed and used based upon commonly agreed
+upon algorithms. This work describes a novel methodology for determining such
+harmonization methods and an open-source Python software library implementing
+the methodology. Results are shown for two example scenarios (with and without
+climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily
+harmonize over 96\% of the total emissions trajectories while having a
+negligible effect on key long-term climate indicators. This new capability
+enhances the comparability across different models, increases transparency and
+robustness of results, and allows other teams to easily participate in
+intercomparison exercises by using the same harmonization mechansim.
 \end{abstract}

### 2017-07-25 elsfrontmatter.tex
more edits

@@
 \begin{abstract}
-Emission harmonization refers to the process used to match results from
-Integrated Assessment Model (IAM) GHG and air pollutants against a common source
-of historical emissions trajectories. To date, harmonization has been performed
+Emission harmonization refers to the process used to match greenhouse gas (GHG)
+and air pollutant results from Integrated Assessment Models (IAMs) against a
+common source of historical emissions. To date, harmonization has been performed
 separately by individual modeling teams. For the hand-over of emission data for
 the Shared Socio-economic Pathways (SSPs) to climate model groups, a new
-automated automatized approach was developed and used based upon commonly agreed
-upon algorithms. This work describes a novel methodology for determining such
-harmonization methods and an open-source Python software library implementing
-the methodology. Results are shown for two example scenarios (with and without
-climate policy cases) using the MESSAGE-GLOBIOM IAM that satisfactorily
-harmonize over 96\% of the total emissions trajectories while having a
-negligible effect on key long-term climate indicators. This new capability
-enhances the comparability across different models, increases transparency and
-robustness of results, and allows other teams to easily participate in
-intercomparison exercises by using the same harmonization mechansim.
+automated approach based on commonly agreed upon algorithms was developed. This
+work describes the novel methodology for determining such harmonization methods
+and an open-source Python software library implementing the methodology. Results
+are shown for two example scenarios (with and without climate policy cases)
+using the MESSAGE-GLOBIOM IAM that satisfactorily harmonize over 96\% of the
+total emissions trajectories while having a negligible effect on key long-term
+climate indicators. This new capability enhances the comparability across
+different models, increases transparency and robustness of results, and allows
+other teams to easily participate in intercomparison exercises by using the same
+harmonization mechansim.
 \end{abstract}

### 2017-07-25 elsfrontmatter.tex
emissions

@@
 \begin{abstract}
-Emission harmonization refers to the process used to match greenhouse gas (GHG)
+Emissions harmonization refers to the process used to match greenhouse gas (GHG)
 and air pollutant results from Integrated Assessment Models (IAMs) against a

### 2017-07-25 elsfrontmatter.tex
openly available

@@
 different models, increases transparency and robustness of results, and allows
-other teams to easily participate in intercomparison exercises by using the same
-harmonization mechansim.
+other teams to easily participate in intercomparison exercises by using the
+same, openly available harmonization mechansim.
 \end{abstract}

### 2017-07-25 elsfrontmatter.tex
hypenate socioeconomic

--- 2017-07-25 elsfrontmatter.tex
@@
 different models, increases transparency and robustness of results, and allows
-other teams to easily participate in intercomparison exercises by using the same
-harmonization mechansim.
+other teams to easily participate in intercomparison exercises by using the
+same, openly available harmonization mechanism.
 \end{abstract}

### 2017-07-25 intro.tex
hypenate socioeconomic

@@
 of climate and economic futures. Recently, the IAM community has developed
-scenarios based on the Shared Socioeconomic Pathways (SSPs)
+scenarios based on the Shared Socio-economic Pathways (SSPs)
 \cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
@@
 models then provide estimates of future trajectories of these variables under
-various socioeconomic and technological assumptions as well as proposed policy
+various socio-economic and technological assumptions as well as proposed policy
 constraints, e.g., targets for future Greenhouse Gas (GHG) emissions.
@@
 scenarios which combines aspects of both the RCPs and SSPs have been developed
-in order to incorporate both physical and socioeconomic detail. In order to
+in order to incorporate both physical and socio-economic detail. In order to
 address the growing dimensionality of model outputs and support ongoing scenario

### 2017-07-25 meth.tex
final edits

@@
 i.e., how well a given harmonization algorithm performs, depends on a number of
-factors, most notably the faithful representation of the original, unharmonized,
-trajectory. This includes the representation of negative trajectories (i.e., if
-a trajectory becomes negative, both the timing and total magnitude should be as
-close as possible) which are of critical importance for cumulative \cotwo
-calculations.
+factors. Of chief import is the faithful representation of the original,
+unharmonized, trajectory as well as the representation of negative trajectories
+(i.e., if a trajectory becomes negative, both the timing and total magnitude
+should be as close as possible) which are of critical importance for cumulative
+\cotwo calculations.
 
@@
 is a significant difference between historical and model values in the
-harmonization year, there is significant upward or downward movement in the
-model trajectory, or if there are known discrepancies in definition between the
-IAM and historical data source. In such cases, override methods can ameliorate
-the issues associated with the default method choice. In order to help identify
-these cases, harmonization \textit{diagnostics} are provided which analyze the
-relative difference between harmonized and unharmonized trajectories at their
-mid (if $> 400$\%) and end-points (if $> 200$\%). If override methods are
-provided, they are used instead of the default methods as determined by the
+harmonization year, if there is significant upward or downward movement in the
+model trajectory, or if there are known discrepancies in sectoral definition
+between the IAM and historical data source. In such cases, override methods can
+ameliorate the issues associated with the default method choice. In order to
+help identify these cases, harmonization \textit{diagnostics} are provided which
+analyze the relative difference between harmonized and unharmonized trajectories
+at their mid (if $> 400$\%) and end-points (if $> 200$\%). If override methods
+are provided, they are used instead of the default methods as determined by the
 default method decision tree. Furthermore, users can set the above-mentioned

### 2017-08-10 meth.tex
updating workflow figure
adding much cleaner 5-region figure

@@
   \begin{center}
-    \includegraphics[height=0.85\textheight]{aneris-workflow.pdf}
+    \includegraphics[height=0.9\textheight]{aneris-workflow.pdf}
     \caption[]{
@@
   \begin{center}
-    \includegraphics[width=0.85\textwidth]{regions.png}
+    \includegraphics[width=\textwidth]{MESSAGE_11-5regions_map.pdf}
     \caption[]{
@@
       The 5 regions used in the RCPs with their 11-region constituents: Asia
-      (Central Asia, South Asia, Pacific Asia), Latin America, the Middle East
-      and Africa, the OECD (North America, Western Europe, and Pacific OECD),
-      Reforming Economies (Eastern Europe and Former Soviet Union).
-    }
+      (Central Asia, South Asia, Pacific Asia) [yellows], Latin
+      America [magenta], the Middle East and Africa [greens], the OECD (North
+      America, Western Europe, and Pacific OECD) [blues], Reforming Economies
+      (Eastern Europe and Former Soviet Union) [reds].  }
   \end{center}

### 2017-08-10 meth.tex
slight text cleanup

@@
       The 5 regions used in the RCPs with their 11-region constituents: Asia
-      (Central Asia, South Asia, Pacific Asia) [yellows], Latin
-      America [magenta], the Middle East and Africa [greens], the OECD (North
-      America, Western Europe, and Pacific OECD) [blues], Reforming Economies
+      (Central Asia, South Asia, Pacific Asia) [yellows], Latin America
+      [magenta], the Middle East and Africa [greens], the OECD (North America,
+      Western Europe, and Pacific OECD) [blues], and the Reforming Economies
       (Eastern Europe and Former Soviet Union) [reds].  }

### 2017-11-23 paper.tex
intro edits
finished dvv edits in intro

typo in intro

methods revisions

typo in texit

some final cleaning of meths

text edits in results

added figure for rad forcing

fix acks

updating forcing fig and text

clearing up some results text

rm wm cmd
so it begins

@@
-%%% ELS includes
-% * <sfujimori112256@gmail.com> 2017-05-23T14:15:19.404Z:
-% 
-% comment from SF
-% 
-% ^ <sfujimori112256@gmail.com> 2017-05-24T18:49:42.852Z.
 \documentclass[review]{elsarticle}
@@
 
-\include{intro}
-\include{meth}
-\include{results}
-% * <sfujimori112256@gmail.com> 2017-05-25T05:18:35.016Z:
-% 
-% I think the current results are basically fine (sorry I have not read in detailed though).
-What we should show and discuss are pretty much depending on how we evaluate these results.
-So, at this stage, I would like to put more comments on this section
-% 
-% 
-% ^.
-\include{future}
+\section{Introduction}
+
+Integrated Assessment Models (IAMs) are tools used to understand the complex
+interactions between energy, economy, land use, water, and climate systems. IAMs
+provide global projections of systemic change by dividing the world into a
+number of representative regions (typically 10 to 30), the definition of which
+is distinct for each model \cite{krey_global_2014}. Results from IAMs are
+integral in a number of international studies, which notably include projections
+of climate and economic futures. Recently, the IAM community has developed
+scenarios based on the Shared Socio-economic Pathways (SSPs)
+\cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
+  calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which quantify a variety of
+potential global futures. The SSPs are designed to be used in research that
+include earth system model (ESM) simulations, climate impact, adaptation and
+climate mitigation studies \cite{vuuren_new_2013}.
+
+While IAMs are implemented in myriad ways, including simulation and
+optimization, the core inputs and outputs are similar across different
+models. Modeling teams incorporate data on energy systems, land use, economics,
+demographics and emissions sources and concentrations, among other data, in
+order to provide consistent existing trajectories of modeled variables. The
+models then provide estimates of future trajectories of these variables under
+various socio-economic and technological assumptions as well as proposed policy
+constraints, e.g., targets for future Greenhouse Gas (GHG) emissions.
+
+The emissions trajectories calculated by IAMs are critical inputs for ongoing,
+worldwide scientific community efforts in the Coupled Model Intercomparison
+Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
+of marker SSP scenarios developed by the IAM community (ScenarioMIP)
+\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
+modeling the individual technologies and sectors that contribute towards the
+emissions of different air pollutants and GHGs as well as various mitigation
+technologies. However, the historical emissions starting points of models can
+differ by large amounts depending on the region, sector, and emissions species.
+
+In practice, IAMs calculate the total source intensity of emitting technologies,
+for example the total activity of coal power plants in China, and incorporate
+emissions-intensity factors for individual gas species, for example the quantity
+of sulfur emissions from coal plants per megawatt-hour of production. Models are
+generally \textit{calibrated} to historical data sources in one or more base
+years. Results in the historical period may differ between models as a result of
+the sometimes large uncertainties in historical data sets. Models can also
+differ in their choice of base-year, which may lag behind available inventory
+data. In addition, models have varying sectoral, regional, and fuel
+aggregations.
+
+The global climate change community has recently developed a new global
+historical emissions data set for both anthropogenic emissions (i.e., the
+Community Emissions Data System (CEDS) \cite{hoesly_historical_2017} and
+open-burning land-use and land-use change (LULUC) emissions
+\cite{van_marle_historic_2017}) which, in conjunction with the SSP IAM
+trajectories, will be used for climate-related modeling exercises of CMIP6.
+
+When participating in intercomparison exercises in which a consistent historical
+starting point is required (e.g., in CMIP6), model teams incorporate a single,
+common historical data set through \textit{harmonization}. Harmonization
+refers to the process of adjusting model results to match a selected
+historical time series such that the resulting future trajectories are also
+consistent with the original modeled results. In the emissions context, this
+means that each individual combination of model region, model sector, and
+emissions species must be harmonized. Depending on the total number of model
+regions, sectors, and emissions species, this can require the selection of
+thousands to tens-of-thousands of harmonization methods.
+
+Harmonization has been addressed in previous studies as it is a common practice
+in the IAM and climate change communities. For example,
+\cite{meinshausen_rcp_2011} describes the use of scaling routines for the 5
+regions used in the Special Report on Emissions Scenarios (SRES)
+\cite{nakicenovic2000}; however, only total emissions were harmonized in the
+exercise, thus there is no sectoral dimension. Further,
+\cite{rogelj_discrepancies_2011} describes the impacts of choosing various
+harmonization routines on future trajectories. During the evaluation of the
+Representative Concentration Pathways (RCPs), IAM results have been harmonized
+by sector and the 5 RCP global regions
+\cite{vuuren_representative_2011}. Importantly, the choice of harmonization
+method to date has been determined by individual experts and has generally been
+applied to all trajectories for a given class of emissions species.
+
+Climate modeling efforts have continued to progress, demanding increased spatial
+and sectoral resolution from IAMs. Furthermore, a new generation of climate
+scenarios which combines aspects of both the RCPs and SSPs have been developed
+in order to incorporate both physical and socio-economic detail. In order to
+address the growing dimensionality of model outputs and support ongoing scenario
+generation and analysis efforts while still providing a consistent and
+scientifically rigorous harmonization procedure, an automated process for
+determining harmonization methods is preferred. The use of an automated,
+documented, and openly available harmonization mechanism additionally allows for
+full procedural reproducibility and for direct participation by additional
+modeling teams not involved in the original exercise.
+
+The remainder of this paper describes the methodology and implementation of the
+harmonization software \code{aneris} \cite{matthew_gidden_2017_802832}, written
+in the Python programming language (detailed documentation available
+\href{http://mattgidden.com/aneris/}{online}). Section \ref{sec:meths} provides
+a detailed description of the underlying mathematical components of
+\code{aneris} as well as the procedural workflow. The results of applying the
+automated harmonization mechanism on two example IAM scenarios, one with
+emissions growth and another with emissions mitigation, is presented in Section
+\ref{sec:results}. Finally, the general effectiveness and potential future
+improvements on the automated methodology is discussed in Section
+\ref{sec:future}.
+\section{Methodology \& Implementation}\label{sec:meths}
+
+\subsection{Harmonization Methods}
+
+IAM emission results are provided along temporal (normally half decade or
+decade), spatial (i.e., model regions), gas species, and sectoral
+dimensions. Each individual temporal trajectory, i.e., unique combinations of
+regions ($r$), species ($g$), and sectors ($s$), must be harmonized to the
+initial modeling period. Given a model trajectory, $m_{r, g, s}(t)$, historical
+values, $h_{r, g, s}(t)$, and model base year, $t_i$, a harmonized trajectory
+needs to be calculated. The \textit{harmonization quality} of a trajectory,
+i.e., how well a given harmonization algorithm performs, depends on a number of
+factors. Of chief import is the faithful representation of the original,
+unharmonized, trajectory as well as the representation of negative trajectories
+(i.e., if a trajectory becomes negative, both the timing and total magnitude
+should be as close as possible) which are of critical importance for cumulative
+\cotwo calculations.
+
+In previous studies \cite{meinshausen_rcp_2011,rogelj_discrepancies_2011}, two
+\textit{families} of methods have been used: those that operate on the ratio of
+base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and those that operate on the
+offset of base year values (i.e., $h(t_i) - m(t_i)$). A number of the classic
+methods are implemented in \code{aneris} including ratio-convergence shown in
+Equation \ref{eqs:ratio}, offset-convergence shown in Equation \ref{eqs:offset},
+and interpolation shown in Equation \ref{eqs:interpolate}. The convergence
+factor, $\beta$, scales linearly from 1 to 0 over $[t_i, t_f)$ and is shown in
+  Equation \ref{eqs:factor}. In all equations the region, species, and sector
+  indices have been dropped for clarity. Each equation is a function of time,
+  model trajectory, historical trajectory, base year ($t_i$), and a convergence
+  year ($t_f$), at which point the harmonized trajectory converges to the
+  unharmonized trajectory. \codeb{aneris} provides a number of methods to choose
+  from for each of the harmonization families. A summary of all available
+  methods is provided in Table \ref{tab:meths}.
+
+\begin{equation}\label{eqs:factor}
+  \beta(t, t_i, t_f) =
+  \begin{cases}
+    1 - \frac{t - t_i}{t_f - t_i},& \text{if } t < t_f\\
+    0,                        & \text{otherwise}
+  \end{cases}
+\end{equation}
+
+\begin{equation}\label{eqs:ratio}
+  m^{rat}(t, m, h, t_i, t_f) = [\beta(t, t_i, t_f) (\frac{h(t_i)}{m(t_i)} - 1) + 1] m(t)
+\end{equation}
+
+\begin{equation}\label{eqs:offset}
+  m^{off}(t, m, h, t_i, t_f) = \beta(t, t_i, t_f) (h(t_i) - m(t_i)) + m(t)
+\end{equation}
+
+\begin{equation}\label{eqs:interpolate}
+  m^{int}(t, m, h, t_i, t_f) =
+  \begin{cases}
+    \beta(t, t_i, t_f) (h(t_i) - m(t_f)) + m(t_f), & \text{if } t < t_f\\
+    m(t), & \text{otherwise}
+  \end{cases}
+\end{equation}
+
+\begin{table}[h!]
+\centering
+\caption{All Harmonization Methods Provided in \code{aneris}}
+\label{tab:meths}
+\begin{tabular}{|l|l|l|}
+\hline
+Method Name                             & Harmonization Family & Convergence Year\\
+\hline
+\code{constant\_ratio}                  & ratio              & $t_f = \infty$\\
+\code{reduce\_ratio\_<year>}            & ratio              & $t_f = $\code{<year>}\\
+\code{constant\_offset}                 & offset             & $t_f = \infty$\\
+\code{reduce\_offset\_<year>}           & offset             & $t_f = $\code{<year>}\\
+\code{linear\_interpolate\_<year>}      & interpolation      & $t_f = $\code{<year>}\\
+\hline
+\end{tabular}
+\end{table}
+
+\subsection{Default Method Decision Tree}\label{sec:tree}
+
+A \textit{decision tree} approach has been implemented in \codeb{aneris} which
+provides a systematic and documented decision-making process to determine the
+preferred harmonization algorithm. In order to provide reasonable
+\textit{default} methods, the historical trajectory, unharmonized model
+trajectory, and relative difference between history and model values in the
+harmonization year are analyzed. The decision tree used in this analysis is a
+result of collaborative efforts between IAM teams and is shown graphically in
+Figure \ref{fig:decision_tree}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{decision_tree.pdf}
+    \caption[]{
+      \label{fig:decision_tree}
+      The default method decision tree used in the \code{aneris} software
+      library. For all decisions, upper (purple) branches represent a ``yes''
+      response and lower (orange) branches represent a ``no'' response. The
+      coefficient of variation, $c_v$, is defined in Equation \ref{eqs:cov},
+      $dH$ is defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
+      decision-making thresholds for $c_v$ and $dH$ are described below. Methods
+      labeled in \textit{green} are likely to closely match unharmonized
+      results, methods in \textit{yellow} will likely somewhat match
+      unharmonized results, and methods in \textit{red} can be expected to have
+      a large relative difference between harmonized and unharmonized results.}
+  \end{center}
+\end{figure}
+
+A number of characteristics impact the decision of which default method to
+select based on the effect of the characteristic on the potential harmonized
+trajectory. For example, it is possible for models to report zero values in the
+harmonization year in situations in which technologies are introduced in future
+time periods in regions or for sectors which produce an emissions species that
+is absent in the initial modeling period. In such cases, an offset method is
+required as a ratio method would mask future emissions and erroneously harmonize
+the trajectory.
+
+In most cases, however, models do report values in the harmonization
+year. Figure \ref{fig:cases} displays a number of example trajectories which
+highlight the possible issues resulting from harmonizing model results in
+different contexts. When model and historical values are relatively close, a
+convergence method is chosen in order to be as representative as possible to the
+underlying unharmonized model results (Figure \ref{fig:cases}, Panel
+\code{a}). If values are not close, the constant ratio method is chosen in order
+to provide reasonable trajectories that still incorporate modeled effects
+(Figure \ref{fig:cases}, Panel \code{b}).
+
+Models can additionally report negative emissions in certain contexts which must
+be taken into account during harmonization. Such a case is possible for gas
+species which can be extracted from the environment and stored, as is the case
+for \cotwo in future scenarios with climate mitigation. If a model provides a
+trajectory that transitions from positive to negative emissions and base year
+results are similar, then a convergence method is used in order to guarantee
+capture of this transition in a representative fashion (Figure \ref{fig:cases},
+Panel \code{c}). If the discrepancy in base year results is large, it is
+possible for a negative trajectory to be inappropriately harmonized to a
+positive, but decreasing, trajectory. As such, the constant ratio method is
+chosen (Figure \ref{fig:cases}, Panel \code{d}).
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{cases.pdf}
+    \caption[]{
+      \label{fig:cases}
+      The effect of different harmonization routines on model trajectories under
+      ``normal'' circumstances (Panel \code{a}), when there is a large
+      difference between historical and model values in the harmonization year
+      (Panels \code{b} and \code{d}), and when model trajectories result in
+      negative emissions by the end of the modeling time horizon (Panels
+      \code{c} and \code{d}). Identical model trajectories are used in each row
+      (Panels \code{a}, \code{b}; \code{c}, \code{d}). In each column,
+      historical values are increased in the base year by an order of magnitude
+      (from 10 to 100). In each Panel, a subset of the potential routines
+      provide a better harmonization quality than others as described in the
+      text.
+    }
+  \end{center}
+\end{figure}
+
+Temporal variability of the historical trajectory is also an important
+characteristic when considering the choice of harmonization method.  Emissions
+from forest and grassland fires, for example, vary from year to year due to a
+combination of meteorological conditions and anthropogenic drivers. Land use
+emissions in many IAMs are modeled using average emission factors and do not
+capture conditions in a specific year. A longer convergence horizon is thus
+desired in order to incorporate highly variable historical data with modeled
+results as is consistency in harmonization method because the effects are
+modeled similarly across regions and species. In order to detect emissions with
+a high amount of variation, a measure of the coefficient of variation, $c_v$, of
+the first derivative of the historical trajectory is calculated using the
+standard deviation, $\sigma$, and the mean, $\mu$, as shown in Equation
+\ref{eqs:cov}.
+
+The value of $c_v$ is then tested against a threshold, $\tau_{c_v}$. To
+determine this threshold, an analysis of the recent CEDS and LUC historical data
+has been performed. Figure \ref{fig:cov} shows the distribution of LUC $c_v$s
+and non-LUC $c_v$s as determined for historical data aggregated to the model
+regions of 5 different IAMs involved in the SSP process. A threshold value of
+$\tau_{c_v} = 20$ has been chosen based on these observations as it
+optimally divides the two distributions. Importantly, tails of the LUC and
+non-LUC overlap, thus there are both false positives (~7\% of non-LUC
+trajectories) and false negatives (~10\% of LUC trajectories). However, as any
+regional definition is model dependent and thus any regional aggregation is
+possible an automated detection mechanism is necessary.
+
+\begin{equation}\label{eqs:cov}
+    c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
+\end{equation}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{cov.pdf}
+    \caption[]{
+      \label{fig:cov}
+      The distribution of $c_v$ values for LUC and non-LUC historical
+      trajectories is shown. CEDS historical data \cite{hoesly_historical_2017}
+      is used for non-LUC data and \cite{van_marle_historic_2017} is used for
+      LUC data. All historical data has been aggregated from countries to IAM
+      model regional definitions, and all gas species included in the historical
+      data sets are included in the analysis. The solid black line indicates the
+      threshold value, $\tau_{c_v}$, used by default in \code{aneris}.  }
+  \end{center}
+\end{figure}
+
+Finally, consideration is taken with respect to the relative difference between
+the historic and model values in the harmonization time period. In order to
+investigate the possible values that these relative differences can take, the
+IAM values used in the SSP and (ongoing) CMIP6 inter-comparison
+exercises are used. A distribution of these differences for all models in the
+study is presented in Figure \ref{fig:dh}. Given the available data, a threshold
+value of $\tau_{dH} = 50$\% was chosen to be used as a default in \code{aneris}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{dh.pdf}
+    \caption[]{
+      \label{fig:dh}
+      The distribution of relative differences between model and historical
+      values in the harmonization year is shown. The solid black line indicates
+      the 50\% threshold value, $\tau_{dH}$, used by default in \code{aneris}.
+    }
+  \end{center}
+\end{figure}
+
+\subsection{\code{aneris} Workflow}\label{sec:workflow}
+
+The full \code{aneris} workflow is comprised of a number of components shown
+graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
+run-control configuration are read in via an Excel spreadsheet. Data is assumed
+to be in the IAMC format, i.e., using \code{Model}, \code{Scenario},
+\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
+representing each modeled time period.
+
+Users are able to control the harmonization process via a number of options. The
+primary mechanism by which users control the process is by providing
+\textit{override} methods for any combination of region and variable (i.e.,
+sector and gas species).  In practice, it may be possible that not all default
+methods chosen will provide robust harmonized trajectories, especially if there
+is a significant difference between historical and model values in the
+harmonization year, if there is significant upward or downward movement in the
+model trajectory, or if there are known discrepancies in sectoral definition
+between the IAM and historical data source. In such cases, override methods can
+ameliorate the issues associated with the default method choice. In order to
+help identify these cases, harmonization \textit{diagnostics} are provided which
+analyze the relative difference between harmonized and unharmonized trajectories
+at their mid (if $> 400$\%) and end-points (if $> 200$\%). If override methods
+are provided, they are used instead of the default methods as determined by the
+default method decision tree. Furthermore, users can set the above-mentioned
+thresholds as well as the LUC method used in the decision tree (see Figure
+\ref{fig:decision_tree}). Further detail of input parameters can be found
+\href{http://mattgidden.com/aneris/config.html}{online}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[height=0.9\textheight]{aneris-workflow.pdf}
+    \caption[]{
+      \label{fig:workflow}
+      The full harmonization process as executed by \code{aneris}. Operations
+      that can be configured with user-based input configurations are shown in
+      purple. The core harmonization process is shown in yellow.  }
+  \end{center}
+\end{figure}
+
+Input data then undergoes a cleaning operation, which adds (null) model
+trajectories that exist in the historical data set but are not provided by the
+model input and detects any issues that would cause the harmonization process to
+fail. The methods used to harmonize the data are then determined and the
+harmonization process is executed. Upon completion of the harmonization process,
+aggregation of common analysis regions is performed. A common regional
+aggregation used in the IAM community was defined in the Representative
+Concentration Pathways (RCPs) \cite{vuuren_representative_2011}, shown in Figure
+\ref{fig:regions}. Finally, any exogenous trajectories the user provides are
+added. Exogenous trajectories are normally provided for unmodeled gases with
+well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
+\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
+regarding the harmonization process are returned. A description of all returned
+meta data is provided in Table \ref{tab:metadata}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{MESSAGE_11-5regions_map.pdf}
+    \caption[]{
+      \label{fig:regions}
+      The 5 regions used in the RCPs with their 11-region constituents: Asia
+      (Central Asia, South Asia, Pacific Asia) [yellows], Latin America
+      [magenta], the Middle East and Africa [greens], the OECD (North America,
+      Western Europe, and Pacific OECD) [blues], and the Reforming Economies
+      (Eastern Europe and Former Soviet Union) [reds].  }
+  \end{center}
+\end{figure}
+
+\begin{table}[]
+\centering
+\caption{Meta data provided by the \code{aneris} harmonization routine. This meta data is provided for every combination of region, sector, and emissions species.}
+\label{tab:metadata}
+\begin{tabular}{|p{2cm}|p{8cm}|}
+\hline
+\textbf{Column}       & \textbf{Description}                                      \\
+\hline
+\hline
+method       & The harmonization method used.                                               \\
+\hline
+default      & The default harmonization method as determined by the default decision tree. \\
+\hline
+override     & The method provided as an override (if any).                                 \\
+\hline
+offset       & The offset value between history and model in the harmonization year.        \\
+\hline
+ratio        & The ratio value between history and model in the harmonization year.         \\
+\hline
+cov          & The coefficient of variation value of the historical trajectory.                           \\
+\hline
+unharmonized & The unharmonized value in the harmonization year.                            \\
+\hline
+history      & The historical value in the harmonization year.                             \\
+\hline
+harmonized   & The resulting harmonized value in the harmonization year.\\
+\hline
+\end{tabular}
+\end{table}
+\section{Results}\label{sec:results}
+
+In order to show a representative cross section of the performance of the
+\code{aneris} harmonization procedure, we focus on the harmonization of results
+of the IAM MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the SSP
+scenario library \cite{Riahi2017153,Rao2017346} are presented. The SSP2, or
+``middle of the road'', scenario (referred to as SSP2-Ref) is chosen to be
+analyzed because MESSAGE-GLOBIOM is the marker scenario for this SSP. This SSP2
+scenario lies between two RCPs, 6 and 8.5, with a radiative forcing level of
+approximately 6.5 Wm$^{-2}$. We additionally present the results for the
+SSP2-based mitigation scenario leading to a radiative forcing of 4.5 Wm$^{-2}$
+(referred to as SSP2-4.5). The SSP2-45 scenario is chosen because mitigation
+technologies and policies are enacted causing a general reduction in pollutants
+and GHGs, including (eventual) negative CO2 emissions in some regions and
+sectors due to carbon capture and sequestration and afforestation. Figure
+\ref{fig:kyoto} shows the different trends of Kyoto Gases in each scenarios.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_kyoto.pdf}
+    \caption[]{
+      \label{fig:kyoto}
+      Unharmonized Kyoto gas emissions for SSP2-Ref, a scenario with generally
+      increasing global emissions trends, and SSP2-45, a scenario with generally
+      decreasing global emissions trends.  }
+  \end{center}
+\end{figure}
+
+MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
+mapped directly to the 5-region definition used in the RCPs. Historical data is
+taken from previously described LUC and anthropogenic sources, which comprise 10
+separate pollutant and GHG species and 12 sectors shown in Table \ref{tab:sp}. A
+total of 970 distinct trajectories were harmonized for each scenario, and
+therefore 1940 trajectories were harmonized in total. \noxx generated from the
+Energy sector provides an example of an emissions species and sector in which
+all regions were satisfactorily harmonized with the default methods. Figure
+\ref{fig:nox} shows the results of harmonization in Asia, and Table
+\ref{tab:nox} describes the parameters that underlie the choice of method for
+each harmonized trajectory.
+
+\begin{table}[]\begin{minipage}{\textwidth}
+  \centering
+  \caption{Harmonized Species and Sectors}
+  \label{tab:sp}
+  \begin{tabular}{|l|l|}
+    \hline
+    \textbf{Emissions Species}        & \textbf{Sectors}                    \\
+    \hline
+    \hline
+    Black Carbon (BC)                 & Agricultural Waste Burning          \\
+    Hexafluoroethane (C\textsubscript{2}F\textsubscript{6})
+    \footnote{\label{x_glb_all}
+      Global total trajectories are harmonized due to lack of detailed 
+      historical data.}
+                                      & Agriculture                         \\
+    Tetrafluoromethane (CF\textsubscript{4})
+    \footnoteref{x_glb_all}               & Aircraft
+    \footnote{\label{x_glb}
+      Global sectoral trajectories are harmonized due to lack of detailed 
+      historical data.}                            \\
+    Methane (CH\textsubscript{4})                     & Energy Sector                       \\
+    Carbon Dioxide (CO\textsubscript{2})
+    \footnote{\label{x_co2}
+      A global trajectory for land-use CO\textsubscript{2} is used; non-land-use sectors 
+      are harmonized for each model region.}
+                                      & Forest Burning                      \\
+    Carbon Monoxide (CO)              & Grassland Burning                   \\
+    Hydrofluorocarbons (HFCs)
+    \footnoteref{x_glb_all}               & Industrial Sector                   \\
+    Nitrous Oxide (N\textsubscript{2}O)
+    \footnoteref{x_glb_all}
+                                      & International Shipping\footnoteref{x_glb}              \\
+    Ammonia (\nht)                     & Residential Commercial Other        \\
+    Nitrogen Oxides (\nox)             & Solvents Production and Application \\
+    Organic Carbon (OC)               & Transportation Sector               \\
+    Sulfur Hexafluoride (SF\textsubscript{6})
+    \footnoteref{x_glb_all}               & Waste                               \\
+    Sulfur Oxides (SO\textsubscript{x})               &                                     \\
+    Volatile Organic Compounds (VOCs) &                                     \\
+    \hline
+  \end{tabular}
+\end{minipage}\end{table}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{example_NOx_Energy_Sector.pdf}
+    \caption[]{
+      \label{fig:nox}
+      \noxx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
+      trajectories for SSP2 and SSP2-45 with historical trajectories (grey
+      lines) are presented. The SSP2 reference scenario is shown in Panels
+      \textbf{a} and \textbf{c}; the SSP-45 scenario is denoted with ``x''
+      markers in Panels \textbf{b} and \textbf{d}. The upper panels show the
+      results for endogenously modeled and harmonized regions in Asia while the
+      lower panels display the aggregate region results.  
+    }
+  \end{center}
+\end{figure}
+
+\begin{table}[]
+\centering
+\caption{Key Parameters for Deciding Harmonization Methods for \noxx Emissions in the Energy Sector in Asia}
+\label{tab:nox}
+\begin{tabular}{|p{1.2cm}|p{.5cm}|p{.5cm}|p{4.5cm}|p{3cm}|}
+\hline
+\textbf{Region} & \textbf{dH} & $\mathbf{c_v}$ & \textbf{Decision Tree Traversal (Branch and Direction)} & \textbf{Default Method Chosen} \\ \hline
+    \hline
+CPA             & 0.35        & 2.26          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
+PAS             & 0.14        & 1.24          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
+SAS             & 0.56        & 0.58          & 1 (no), 2 (no), 3 (no), 4 (no)                          & constant\_ratio        \\ \hline
+\end{tabular}
+\end{table}
+
+The harmonization of emissions pathways is performed in order to accurately
+represent new or updated data sets of historical emissions inventories while also
+maintaining consistency with the original, unharmonized pathway. As such, when
+the default methods as provided by the harmonization procedure distort or
+otherwise sufficiently misrepresent the underlying unharmonized results, an
+override method is required to be provided for the trajectory of the region,
+sector, and species in question. Of the 970 trajectories, approximately 10\%
+were reported as a diagnostic (see Section \ref{sec:workflow}) of which 3.5\%
+required the use of harmonization overrides after an initial investigation;
+thus, 96.5\% of all trajectories were satisfactorily harmonized using the
+default methods. The trajectories that required overrides clustered into two
+classifications: regional trajectories whose \textit{magnitude} was overly
+distorted and regional trajectories whose \textit{shape} was overly distorted.
+
+Figure \ref{fig:co} presents a case in which the magnitude of a trajectory is
+distorted. A large discrepancy ($\sim$300\% relative difference) is observed in
+the harmonization year for carbon monoxide (CO) emissions in the industrial
+sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region, which
+comprises most of the emissions of the Asian subcontinent. The default method
+chosen (\code{constant_ratio}) maintains model trends for the region; however,
+overall model results are distorted. By applying a \code{constant_offset}
+override, the regional trend and magnitude is maintained. With the new
+harmonization method for the SAS region, the global trajectory for industrial CO
+also is representative the trends seen in the unharmonized trajectory and the
+relative importance of the underlying regional trajectories is maintained.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_CO_Industrial_Sector.pdf}
+    \caption[]{
+      \label{fig:co}
+      CO Industrial Sector harmonized and unharmonized emissions are presented
+      for SSP2 and SSP2-45 scenarios. Scenarios as denoted identically to Figure
+      \ref{fig:nox}. Panels \textbf{a} and \textbf{b} show harmonized and
+      overridden-harmonized (respectively) regional trajectories for the 3
+      MESSAGE-GLOBIOM regions that comprise the R5ASIA region: Centrally Planned
+      Asia (CPA), Other Pacific Asia (PAS), and South Asia (SAS). Notably, the
+      SAS regional trajectory displays a distorted trajectory due to the
+      harmonization-year difference between history and model results in both
+      scenarios. The distortion is large enough to affect global results, as
+      shown in Panels \textbf{c} and \textbf{d}.  
+}
+  \end{center}
+\end{figure}
+
+In certain circumstances, the application of the default harmonization methods
+can affect not only the magnitude but also the shape of regional
+trajectories. Figure \ref{fig:nh3} shows an example case of emissions
+trajectories for ammonia (\nht) from the agriculture sector in Asia. Again, the
+SAS region shows a large discrepancy in the harmonization year ($>$150\% in this
+case). The resulting trajectory harmonized with the default method
+(\code{constant_ratio}) provides a large increase after 2080 in the SSP2
+reference scenario. Notably, the SSP2-45 scenario is not affected to the same
+degree. While this distortion changes the magnitude of the SAS trajectory, it
+largely affects the post-2080 shape of the global trajectory (see Figure
+\ref{fig:nh3}, panel \textbf{c}). By using a \code{constant_offset} method as an
+override, this distortion is addressed and more accurately reflects unharmonized
+results in the SAS region, the relative importance between regions, and global
+results for agricultural ammonia emissions, each of which contributes to a
+better harmonization quality for the harmonized SAS trajectory.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_NH3_Agriculture.pdf}
+    \caption[]{
+      \label{fig:nh3}
+      \nhtt agricultural harmonized and unharmonized emissions are presented for
+      SSP2 and SSP2-45 scenarios. Scenarios and panel layouts are identical to
+      Figure \ref{fig:co}. In this case, the SAS trajectory again shows not only
+      a magnitude distortion, but also a shape distortion at the tail of the
+      trajectory. Additionally, global trajectories are greatly affected by the
+      harmonization method choice (there is $\sim$20\% relative difference
+      between trajectories in the reference scenario in 2100). Override methods
+      have been applied to correct the distortion.  }
+  \end{center}
+\end{figure}
+
+We investigate the aggregate effect of harmonization with all necessary override
+methods on total anthropogenic radiative forcing projections with the simple
+carbon-cycle and climate model, MAGICC6 \cite{meinshausen2011emulating,
+  meinshausen2011rcp}, for each harmonized and unharmonized scenario
+respectively as shown in Figure \ref{fig:forcing}. We find that the change due
+to harmonization is small, ranging between 1 and 2.5\% over the model
+horizon. Relative near-term differences persist in the mitigation case (SSP-4.5)
+because differences in near-term emissions define to a larger degree the
+longer-term forcing outcome due to the cumulative nature of long-lived climate
+forcers like \cotwo.  The resulting difference in forcing in 2100 is 0.04
+$\frac{\text{W}}{\text{m}^2}$ for SSP2-4.5 and 0.01
+$\frac{\text{W}}{\text{m}^2}$ for SSP2-Ref, both of which are well within
+acceptable tolerances (e.g., 0.75 $\frac{\text{W}}{\text{m}^2}$ defined for
+ScenarioMIP \cite{oneill_scenario_2016}). Thus harmonization is considered to
+have a negligible effect on key long-term climate indicators.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{results_forcing.pdf}
+    \caption[]{
+      \label{fig:forcing}
+      The results of the simple climate model, MAGICC6, forced with the SSP2-Ref
+      (blue) and SSP2-4.5 (green) harmonized and unharmonized scenarios is
+      presented. The radiative forcing trajectories of harmonized and
+      unharmonized scenarios are shown in solid lines and dashed lines,
+      respectively.  }
+  \end{center}
+\end{figure}
+\section{Discussion \& Future Work}\label{sec:future}
+
+This work presented a novel methodology and Python implementation of automated
+emissions harmonization for IAMs, \code{aneris}. An in-depth explanation of the
+processes and methods for determining the use of harmonization methods was
+provided in Section \ref{sec:meths}. \code{aneris} was able to satisfactorily
+harmonized over 96\% of the 1940 individual trajectories that were analyzed in
+Section \ref{sec:results}. Of the remaining trajectories, harmonization method
+overrides were applied, and example situations in overrides were needed were
+discussed.
+
+The automated approach drastically reduces the need for expert opinion in
+determining harmonization methods for each individual combination of model
+region, sector, and emissions species while still providing a justifiable
+explanation for each automated choice of harmonization method based on both the
+historical and future emissions trajectories. Furthermore, the automated
+approach continues to scale well as models become more detailed in both the
+regional and sectoral dimensions. Finally, expert opinion is still allowed to
+trump the automated method as determined by the algorithm via method overrides;
+however, these cases are clearly documented via the meta data provided as an
+output of \code{aneris} and thus can be individually explained. This provides
+not only transparency and reproducibility, but also scientific integrity in the
+choice of harmonization methods.
+
+The use of an open-source, automated harmonization process also provides
+benefits to the wider climate science and IAM communities. By providing a
+standard mechanism for harmonization, the IAM community can directly provide
+input into the harmonization algorithms and rules for their default
+selection. Additionally, modeling teams are easily capable of executing
+identical harmonization procedures in order to participate in ongoing and
+further iterations of intercomparison exercises and analysis. Future scenario
+analyses can also utilize this common harmonization approach such that they are
+consistent with prior efforts.
+
+There are a variety of avenues for future improvement of both the \code{aneris}
+software and underlying methodology. As with any software project, additional
+users will provide use cases for more robust handling of input/output concerns
+and corner cases. Further configuration parameters can also be added in the
+future in order to provide overrides for all gas species in a given sector or
+region. Perhaps the most fruitful investigation will involve further refinement
+of the default decision tree introduced in Section \ref{sec:meths}. A key aspect
+missing from the decision tree is input from models regarding whether missing
+sources or activity levels are the likely cause of a harmonization year
+discrepancy (suggesting the use of an offset method) or instead a significant
+difference in emissions factors (suggesting the use of a ratio method)
+\cite{rogelj_discrepancies_2011}.
+
+This work provides a new direction and framework which the IAM and climate
+communities can build upon in order to reduce the necessity of consistent expert
+opinion and increase transparency and reproducibility of harmonization
+exercises. Furthermore, it provides an open-source, tested, and documented
+software library which can be used and improved upon by these communities. Both
+of these are clear steps in a positive direction for future climate and
+integrated assessment modeling exercises.
 
@@
 contribute both discussion and feedback regarding this work including Drs. Elmar
-Kriegler, Gunnar Luderer, and Joeri Rogelj. The authors further wish to thank
-the Global Environment Facility for their generous support. This project has
-received funding from the European Union’s Horizon 2020 research and innovation
-programme under grant agreement No 641816.
+Kriegler, Gunnar Luderer, and Joeri Rogelj. This project has received funding
+from the European Union’s Horizon 2020 research and innovation programme under
+grant agreement No 641816. The authors further wish to thank the Global
+Environment Facility for their generous financial support.
 

### 2017-11-23 paper.tex
further reduction to one file, start to use glossary

@@
 %%% General usepackages, definitions, etc.
-\input{preamble}
+\usepackage[acronym,toc]{glossaries}
+\usepackage{graphicx}
+\usepackage{booktabs} % nice rules for tables
+\usepackage{microtype} % if using PDF
+\usepackage{hyperref}
+\usepackage{amsmath}
+\usepackage{moreverb} % for verbatim snippets of code
+\usepackage{fancyvrb}
+\usepackage{tabularx} % for tables with line breaks
+\usepackage{threeparttable} % for tables with notes
+\usepackage[capitalize, noabbrev]{cleveref} % for reference multiple figures
+\usepackage{calc} % allows for arithmetic on latex variables
+\usepackage{float} % allows for figures to be placed explicitly
+\usepackage{algorithm2e} % for algorithms
+\usepackage{subfig}
+\usepackage{multirow} % combining rows in tables
+\usepackage{footnote}
+\usepackage{titlesec} % for using \titleformat
+\usepackage{fixltx2e} % for using \textsubscript
+\usepackage{bashful}
+\usepackage{xspace}
+\usepackage{color}
+%%%%%
 
-%%% General setup
+%%%%% Helpers
+
+\newcommand{\code}[1]{\lstinline[basicstyle=\ttfamily\color{black}]|#1|}
+\newcommand{\codeb}[1]{\texttt{#1}}
+\newcommand{\units}[1] {\:\text{#1}}%
+\newcommand{\TODO}[1]{\textbf{TODO: #1}}
+
+%%% gases
+\newcommand{\cotwo}{CO\textsubscript{2}~}
+\newcommand{\nox}{NO\textsubscript{x}}
+\newcommand{\noxx}{NO\textsubscript{x}~}
+\newcommand{\nht}{NH\textsubscript{3}}
+\newcommand{\nhtt}{NH\textsubscript{3}~}
+
+\makeatletter
+\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}
+\makeatother
+
+%%% Figure path
 \graphicspath{{./figs/}}
-%\include{acros}
+
+%%% Acronyms
+\newacronym{ssps}{SSPs}{Shared Socioeconomic Pathways}
+
 \makeglossaries
@@
 %%% ELS input
-\input{elsfrontmatter} 
+\begin{frontmatter}
+
+\title{A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models}
+
+%% Group authors per affiliation:
+\author[iiasa]{Matthew J. Gidden\corref{corref}}
+\ead{gidden@iiasa.ac.at}
+\cortext[corref]{Corresponding author}
+
+\author[nies]{Shinichiro Fujimori}
+\author[pbl]{Maarten van den Berg}
+\author[pik]{David Klein}
+\author[pnnl]{Steven J. Smith}
+\author[pbl]{Detlef P. van Vuuren}
+\author[iiasa]{Keywan Riahi}
+
+\address[iiasa]{International Institute for Applied Systems Analysis,
+  Schlossplatz 1, A-2361 Laxenburg, Austria}
+\address[nies]{Center for Social and Environmental Systems Research, National Institute for Environmental Studies, 16-2 Onogawa, Tsukuba, Ibaraki 305-8506, Japan}
+\address[pbl]{PBL Netherlands Environmental Assessment Agency, Postbus 30314, 2500 GH The Hague, Netherlands}
+\address[pik]{Potsdam Institute for Climate Impact Research (PIK), Member of the Leibniz Association, P.O. Box 60 12 03, D-14412 Potsdam, Germany}
+\address[pnnl]{Joint Global Change Research Institute, 5825 University Research Court, Suite 3500, College Park, MD 20740}
+
+\begin{abstract}
+Emissions harmonization refers to the process used to match greenhouse gas (GHG)
+and air pollutant results from Integrated Assessment Models (IAMs) against a
+common source of historical emissions. To date, harmonization has been performed
+separately by individual modeling teams. For the hand-over of emission data for
+the \gls{ssps} to climate model groups, a new automated approach based on
+commonly agreed upon algorithms was developed. This work describes the novel
+methodology for determining such harmonization methods and an open-source Python
+software library implementing the methodology. Results are shown for two example
+scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM
+that satisfactorily harmonize over 96\% of the total emissions trajectories
+while having a negligible effect on key long-term climate indicators. This new
+capability enhances the comparability across different models, increases
+transparency and robustness of results, and allows other teams to easily
+participate in intercomparison exercises by using the same, openly available
+harmonization mechanism.
+\end{abstract}
+
+\begin{keyword}
+Integrated Assessment Models, Climate Change, Harmonization, Air Pollution 
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
+
 %%%
@@
 of climate and economic futures. Recently, the IAM community has developed
-scenarios based on the Shared Socio-economic Pathways (SSPs)
+scenarios based on the \gls{ssps}
 \cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
@@
 \newpage
+%Print the glossary
+\printglossaries
+
+\newpage
 \section*{\refname}

### 2018-01-29 paper.tex
adding initial glossary

@@
 \newacronym{ssps}{SSPs}{Shared Socioeconomic Pathways}
+\newacronym{rcps}{RCPs}{Representative Concentration Pathways}
+\newacronym{cmip6}{CMIP6}{Coupled Model Intercomparison Project (Phase 6)}
+\newacronym{smip}{ScenarioMIP}{Scenario Model Intercomparison Project}
+\newacronym{ceds}{CEDS}{Community Emissions Data System}
+\newacronym{iam}{IAM}{Integrated Assessment Model}
+\newacronym{esm}{ESM}{Earth System Model}
+\newacronym{ghg}{GHG}{Greenhouse Gas}
+\newacronym{luc}{LUC}{Land-use Change}
+\newacronym{luluc}{LULUC}{Land-use and Land-use Change}
+\newacronym{weu}{WEU}{Western Europe}
+\newacronym{nam}{NAM}{North America}
+\newacronym{pao}{PAO}{Pacific OECD}
+\newacronym{eeu}{EEU}{Central and Eastern Europe}
+\newacronym{fsu}{FSU}{Former Soviet Union}
+\newacronym{pas}{PAS}{Other Pacific Asia}
+\newacronym{cpa}{CPA}{Centrally-planned Asia and China}
+\newacronym{sas}{SAS}{South Asia}
+\newacronym{mea}{MEA}{Middle East and North Africa}
+\newacronym{afr}{AFR}{Sub-Saharan Africa}
+\newacronym{lam}{LAM}{Latin America and the Caribbean}
+
+%% all model names
+%% all gases
 

### 2018-01-29 paper.tex
saving to make multifigure

@@
 \ref{sec:future}.
+
 \section{Methodology \& Implementation}\label{sec:meths}
@@
 
-\subsection{\code{aneris} Workflow}\label{sec:workflow}
+\subsection{\code{aneris} Workflow and Python Implementation}\label{sec:workflow}
 
-The full \code{aneris} workflow is comprised of a number of components shown
-graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
-run-control configuration are read in via an Excel spreadsheet. Data is assumed
-to be in the IAMC format, i.e., using \code{Model}, \code{Scenario},
-\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
-representing each modeled time period.
+We herein present \code{aneris}' conceptual workflow as well as specific details
+of the Python implementation. The workflow is depicted in Figure
+\ref{fig:workflow}, and the implementation is depicted in Figure
+\ref{fig:software}.
+
+\code{aneris} is comprised of a number of conceptual workflow and software
+components shown graphically in Figures \ref{fig:workflow} and
+\ref{fig:software}, respectively. Unharmonized model data and a run-control
+configuration are read in via an Excel spreadsheet. Data is assumed to be in the
+IAMC format, i.e., using \code{Model}, \code{Scenario}, \code{Region},
+\code{Variable}, and \code{Unit} columns in addition to columns representing
+each modeled time period.
 
@@
       purple. The core harmonization process is shown in yellow.  }
+  \end{center}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\textwidth]{implementation_schematic.pdf}
+    \caption[]{
+      \label{fig:software}
+      foo
+    }
   \end{center}

### 2018-01-30 paper.tex
saving

--- 2018-01-29 paper.tex
@@
 \usepackage{algorithm2e} % for algorithms
-\usepackage{subfig}
 \usepackage{multirow} % combining rows in tables
@@
 \usepackage{color}
+\usepackage{subfig}
+\usepackage{minted}
 %%%%%
@@
 \makeatother
+
+%%%python syntax highlighting
+% Default fixed font does not support bold face
+\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{10} % for bold
+\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{10}  % for normal
+
+% Custom colors
+\usepackage{color}
+\definecolor{deepblue}{rgb}{0,0,0.5}
+\definecolor{deepred}{rgb}{0.6,0,0}
+\definecolor{deepgreen}{rgb}{0,0.5,0}
+
+\usepackage{listings}
+
+% Python style for highlighting
+\newcommand\pythonstyle{\lstset{
+language=Python,
+basicstyle=\ttm,
+otherkeywords={self},             % Add keywords here
+keywordstyle=\ttb\color{deepblue},
+emph={MyClass,__init__},          % Custom highlighting
+emphstyle=\ttb\color{deepred},    % Custom highlighting style
+stringstyle=\color{deepgreen},
+frame=tb,                         % Any extra options here
+showstringspaces=false            % 
+}}
+
+% Python environment
+\lstnewenvironment{python}[1][]
+{
+\pythonstyle
+\lstset{#1}
+}
+{}
 
@@
 \newacronym{lam}{LAM}{Latin America and the Caribbean}
+\newacronym{cli}{CLI}{Command Line Interface}
 
@@
 
-\subsection{\code{aneris} Workflow}\label{sec:workflow}
+\subsection{\code{aneris} Python Implementation and Workflow}\label{sec:workflow}
 
-The full \code{aneris} workflow is comprised of a number of components shown
-graphically in Figure \ref{fig:workflow}. Unharmonized model data and a
-run-control configuration are read in via an Excel spreadsheet. Data is assumed
-to be in the IAMC format, i.e., using \code{Model}, \code{Scenario},
+We herein present \code{aneris}' Python implementation and conceptual
+workflow. The implementation is depicted in Figure \ref{fig:software}, and the
+conceptual workflow is depicted in Figure \ref{fig:workflow}. The software is
+composed of three primary components: the \texttt{HarmonizationDriver},
+\texttt{Harmonizer}, and data processing routines (shown in red, green, and
+blue, respectively in Figure \ref{fig:software-workflow}).
+
+The \texttt{HarmonizationDriver} is an object designed to interface with
+user-provided data and configuration files. Input data is assumed to be an Excel
+file the in the in the IAMC format, i.e., using \code{Model}, \code{Scenario},
 \code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
-representing each modeled time period.
+representing each modeled time period, which is the standard format within the
+IAM community. It is responsible for downselecting data into unique combinations
+for each model and scenario and then invoking the harmonization
+process. Finally, it adds any exogenous emissions trajectories as required by
+the user and recompiles the results and writes them to disc. The
+\texttt{HarmonizationDriver} acts the primary interface for high-level users as
+shown in Listing \ref{lst:tutorial}. Furthermore, a \gls{cli} is provided
 
-Users are able to control the harmonization process via a number of options. The
-primary mechanism by which users control the process is by providing
-\textit{override} methods for any combination of region and variable (i.e.,
-sector and gas species).  In practice, it may be possible that not all default
-methods chosen will provide robust harmonized trajectories, especially if there
-is a significant difference between historical and model values in the
-harmonization year, if there is significant upward or downward movement in the
-model trajectory, or if there are known discrepancies in sectoral definition
-between the IAM and historical data source. In such cases, override methods can
-ameliorate the issues associated with the default method choice. In order to
-help identify these cases, harmonization \textit{diagnostics} are provided which
-analyze the relative difference between harmonized and unharmonized trajectories
-at their mid (if $> 400$\%) and end-points (if $> 200$\%). If override methods
-are provided, they are used instead of the default methods as determined by the
-default method decision tree. Furthermore, users can set the above-mentioned
-thresholds as well as the LUC method used in the decision tree (see Figure
-\ref{fig:decision_tree}). Further detail of input parameters can be found
-\href{http://mattgidden.com/aneris/config.html}{online}.
+\begin{python}[caption=High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://mattgidden.com/aneris/config.html}{tutorial}, label=lst:tutorial]
+from aneris.tutorial import load_data
+model, hist, driver = load_data()
+for scenario in driver.scenarios():
+    driver.harmonize(scenario)
+harmonized, metadata = driver.harmonized_results()
+\end{python}
+
+\begin{lstlisting}[language=bash, caption=bash, label=lst:cli]
+$ aneris -h
+usage: aneris [-h] [--history HISTORY] [--regions REGIONS] [--rc RC]
+              [--output_path OUTPUT_PATH] [--output_prefix OUTPUT_PREFIX]
+              input_file
+    Harmonize historical trajectories to data in the IAMC template format.
+    Example usage:
+    aneris input.xlsx --history history.csv --regions regions.csv
+positional arguments:
+  input_file            Input data file.
+optional arguments:
+  -h, --help            show this help message and exit
+  --history HISTORY     Historical emissions in the base year.
+  --regions REGIONS     Mapping of country iso-codes to native regions.
+  --rc RC               Runcontrol YAML file (see
+                        http://mattgidden.com/aneris/config.html for
+                        examples).
+  --output_path OUTPUT_PATH
+                        Path to use for output file names.
+  --output_prefix OUTPUT_PREFIX
+                        Prefix to use for output file names.
+\end{lstlisting}
+
+The \texttt{Harmonizer} is a Python class whose concern is to harmonize various
+model value trajectories given historical data and possible user method
+\textit{overrides}, i.e., non-default methods. It is used by the
+\texttt{HarmonizationDriver}; however it is also available to the user as a
+first-class object. The \texttt{Harmonizer} requires that input data conform to
+the \code{aneris} calculation data format, which expliciltly sepearates the
+emissions species from the sector contributing the emissions (these are combined
+in the single \code{Variable} column in the IAMC format). Because the
+\texttt{Harmonizer} is configure for a single instance of a model and scenario,
+the canonical data format includes columns named \texttt{region},
+\texttt{sector}, \texttt{gas}, and \texttt{units}. Once configured with
+appropriate input data (model and history) as well as potential method
+overrides, the \texttt{Harmonizer}'s \texttt{harmonize()} method can be invoked
+which returns a \texttt{pandas.DataFrame} of harmonized data. The object can
+additionally be queried directly as to its \texttt{default\_methods()},
+\texttt{methods()} (i.e., methods used with overrides), and \texttt{metadata()}
+(i.e., methods used with all branching information along each path in the
+decision tree).
+
+There are also a variety of tools and utilities provided to users and also used
+by the \texttt{HarmonizationDriver} in order to process both input and output
+data. These include an \texttt{EmissionsAggregator} class and related routines
+used to combine historical emissions to native model regions (where historical
+data is defined at a higher spatial resolution than a model; see, e.g., Figure
+\ref{fig:regions}) and a \texttt{FormatTranslator} class which defines an
+interface for translating \texttt{pandas.DataFrame}s between the IAMC format
+expected for input and output data and the calculation format used by
+\code{aneris}' \texttt{Harmonizer}.
 
 \begin{figure}
-  \begin{center}
-    \includegraphics[height=0.9\textheight]{aneris-workflow.pdf}
-    \caption[]{
-      \label{fig:workflow}
+  \subfloat[Python Implementation\label{fig:software}]{
+    \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
+    }
+  \subfloat[Workflow\label{fig:workflow}]{
+    \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
+    }
+    \caption{
       The full harmonization process as executed by \code{aneris}. Operations
@@
       purple. The core harmonization process is shown in yellow.  }
-  \end{center}
+    \label{fig:software-workflow}
 \end{figure}
+
+Users are able to control the harmonization process via a number of options
+(with examples provided
+\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
+by which users control the process is by providing \textit{override} methods for
+any combination of region and variable (i.e., sector and gas species).  In
+practice, it may be possible that not all default methods chosen will provide
+robust harmonized trajectories, especially if there is a significant difference
+between historical and model values in the harmonization year, if there is
+significant upward or downward movement in the model trajectory, or if there are
+known discrepancies in sectoral definition between the IAM and historical data
+source. In such cases, override methods can ameliorate the issues associated
+with the default method choice. In order to help identify these cases,
+harmonization \textit{diagnostics} are provided which analyze the relative
+difference between harmonized and unharmonized trajectories at their mid (if $>
+400$\%) and end-points (if $> 200$\%). If override methods are provided, they
+are used instead of the default methods as determined by the default method
+decision tree. Furthermore, users can set the above-mentioned thresholds as well
+as the LUC method used in the decision tree (see Figure
+\ref{fig:decision_tree}).
 

### 2018-01-30 paper.tex
update implementation section, use minted for code snippets

@@
 \makeatother
-
-%%%python syntax highlighting
-% Default fixed font does not support bold face
-\DeclareFixedFont{\ttb}{T1}{txtt}{bx}{n}{10} % for bold
-\DeclareFixedFont{\ttm}{T1}{txtt}{m}{n}{10}  % for normal
-
-% Custom colors
-\usepackage{color}
-\definecolor{deepblue}{rgb}{0,0,0.5}
-\definecolor{deepred}{rgb}{0.6,0,0}
-\definecolor{deepgreen}{rgb}{0,0.5,0}
-
-\usepackage{listings}
-
-% Python style for highlighting
-\newcommand\pythonstyle{\lstset{
-language=Python,
-basicstyle=\ttm,
-otherkeywords={self},             % Add keywords here
-keywordstyle=\ttb\color{deepblue},
-emph={MyClass,__init__},          % Custom highlighting
-emphstyle=\ttb\color{deepred},    % Custom highlighting style
-stringstyle=\color{deepgreen},
-frame=tb,                         % Any extra options here
-showstringspaces=false            % 
-}}
-
-% Python environment
-\lstnewenvironment{python}[1][]
-{
-\pythonstyle
-\lstset{#1}
-}
-{}
 
@@
 \texttt{HarmonizationDriver} acts the primary interface for high-level users as
-shown in Listing \ref{lst:tutorial}. Furthermore, a \gls{cli} is provided
+shown in Listing \ref{lst:tutorial}. Furthermore, a \gls{cli} is provided to
+allow users to more easily incorporate the harmonization process in scripted
+workflows (Listing \ref{lst:cli}, Figure \ref{fig:software}).
 
-\begin{python}[caption=High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://mattgidden.com/aneris/config.html}{tutorial}, label=lst:tutorial]
+\begin{listing}[H]
+\begin{minted}{python}
 from aneris.tutorial import load_data
@@
 harmonized, metadata = driver.harmonized_results()
-\end{python}
+\end{minted}
+\caption{High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://mattgidden.com/aneris/config.html}{tutorial}}
+\label{lst:tutorial}
+\end{listing}
 
-\begin{lstlisting}[language=bash, caption=bash, label=lst:cli]
+\begin{listing}[H]
+\begin{minted}{bash}
 $ aneris -h
@@
               input_file
+
     Harmonize historical trajectories to data in the IAMC template format.
+
     Example usage:
+
     aneris input.xlsx --history history.csv --regions regions.csv
+
 positional arguments:
   input_file            Input data file.
+
 optional arguments:
@@
                         Prefix to use for output file names.
-\end{lstlisting}
+\end{minted}
+\caption{The \gls{cli} help provided by the \code{aneris} package.}
+\label{lst:cli}
+\end{listing}
 
@@
     }
-    \caption{
-      The full harmonization process as executed by \code{aneris}. Operations
-      that can be configured with user-based input configurations are shown in
-      purple. The core harmonization process is shown in yellow.  }
+    \caption{ Panel \ref{fig:software} shows the various objects in the
+      \code{aneris} codebase as well as a short description of their scope of
+      concern. Panel \ref{fig:workflow} displays the full harmonization process
+      as executed by \code{aneris} for a single instance of a model and
+      scenario. Operations that can be configured with user-based input
+      configurations are shown in purple. Operations goverened by the
+      \texttt{HarmonizationDriver} are shown in red. Data processing operations
+      are shown in blue. The core harmonization process, governed by the
+      \texttt{Harmonizer} is shown in green.  }
     \label{fig:software-workflow}
@@
 source. In such cases, override methods can ameliorate the issues associated
-with the default method choice. In order to help identify these cases,
-harmonization \textit{diagnostics} are provided which analyze the relative
-difference between harmonized and unharmonized trajectories at their mid (if $>
-400$\%) and end-points (if $> 200$\%). If override methods are provided, they
-are used instead of the default methods as determined by the default method
-decision tree. Furthermore, users can set the above-mentioned thresholds as well
-as the LUC method used in the decision tree (see Figure
-\ref{fig:decision_tree}).
+with the default method choice.
+
+In order to help identify these cases, harmonization \textit{diagnostics} are
+provided which analyze the relative difference between harmonized and
+unharmonized trajectories at their mid and end-points. These values are
+configurable by the user, but defaults of $400$\% and $200$\%, respectively, are
+provided based on experiences of using \code{aneris} to date. If override
+methods are provided, they are used instead of the default methods as determined
+by the default method decision tree. Furthermore, users can set the
+above-mentioned thresholds as well as the LUC method used in the decision tree
+(see Figure \ref{fig:decision_tree}).
 

### 2018-01-30 paper.tex
finish up impl section

@@
 
-We herein present \code{aneris}' Python implementation and conceptual
-workflow. The implementation is depicted in Figure \ref{fig:software}, and the
-conceptual workflow is depicted in Figure \ref{fig:workflow}. The software is
-composed of three primary components: the \texttt{HarmonizationDriver},
-\texttt{Harmonizer}, and data processing routines (shown in red, green, and
-blue, respectively in Figure \ref{fig:software-workflow}).
+We herein present \code{aneris}' Python implementation, depicted in Figure
+\ref{fig:software}, and conceptual, depicted in Figure \ref{fig:workflow}. The
+library is composed of a number of utilities as well as three primary
+components: the \texttt{HarmonizationDriver}, \texttt{Harmonizer}, and data
+processing routines (shown in red, green, and blue, respectively in Figure
+\ref{fig:software-workflow}).
 
 The \texttt{HarmonizationDriver} is an object designed to interface with
-user-provided data and configuration files. Input data is assumed to be an Excel
-file the in the in the IAMC format, i.e., using \code{Model}, \code{Scenario},
-\code{Region}, \code{Variable}, and \code{Unit} columns in addition to columns
-representing each modeled time period, which is the standard format within the
-IAM community. It is responsible for downselecting data into unique combinations
-for each model and scenario and then invoking the harmonization
-process. Finally, it adds any exogenous emissions trajectories as required by
-the user and recompiles the results and writes them to disc. The
+user-provided data and configuration files. Input data (i.e., unharmonized model
+results) is assumed to be an Excel file in the standard data format within the
+IAM community, i.e., with \code{Model}, \code{Scenario}, \code{Region},
+\code{Variable}, and \code{Unit} columns in addition to columns representing
+each modeled time period. It is responsible for down-selecting data into
+separate datasets for each model and scenario, invoking the harmonization
+process on each dataset, and recompiling the results. The
 \texttt{HarmonizationDriver} acts the primary interface for high-level users as
-shown in Listing \ref{lst:tutorial}. Furthermore, a \gls{cli} is provided to
-allow users to more easily incorporate the harmonization process in scripted
-workflows (Listing \ref{lst:cli}, Figure \ref{fig:software}).
+shown by the usage of the \texttt{driver} object in Listing
+\ref{lst:tutorial}. Furthermore, a \gls{cli} is provided to allow users to more
+easily incorporate the harmonization process in scripted workflows (Listing
+\ref{lst:cli}, Figure \ref{fig:software}).
 
@@
 
-The \texttt{Harmonizer} is a Python class whose concern is to harmonize various
-model value trajectories given historical data and possible user method
-\textit{overrides}, i.e., non-default methods. It is used by the
-\texttt{HarmonizationDriver}; however it is also available to the user as a
-first-class object. The \texttt{Harmonizer} requires that input data conform to
-the \code{aneris} calculation data format, which expliciltly sepearates the
-emissions species from the sector contributing the emissions (these are combined
-in the single \code{Variable} column in the IAMC format). Because the
-\texttt{Harmonizer} is configure for a single instance of a model and scenario,
-the canonical data format includes columns named \texttt{region},
-\texttt{sector}, \texttt{gas}, and \texttt{units}. Once configured with
-appropriate input data (model and history) as well as potential method
+The \texttt{Harmonizer} is a Python class whose concern is to harmonize model
+value trajectories given historical data and possible user method
+\textit{overrides}, i.e., non-default methods (described further below). It is
+used by the \texttt{HarmonizationDriver}; however it is also available to the
+user as a first-class object. The \texttt{Harmonizer} requires that input data
+conform to the \code{aneris} calculation data format, which explicitly separates
+the emissions species from the sector contributing the emissions (these are
+combined in the single \code{Variable} column in the standard IAM
+format). Because the \texttt{Harmonizer} is designed to operate on a single
+instance of a model and scenario, the canonical data format includes
+\texttt{region}, \texttt{sector}, \texttt{gas}, and \texttt{units} columns
+without extraneous metadata columns for the model and scenario. Once configured
+with appropriate input data (model and history) as well as potential method
 overrides, the \texttt{Harmonizer}'s \texttt{harmonize()} method can be invoked
@@
 data. These include an \texttt{EmissionsAggregator} class and related routines
-used to combine historical emissions to native model regions (where historical
-data is defined at a higher spatial resolution than a model; see, e.g., Figure
-\ref{fig:regions}) and a \texttt{FormatTranslator} class which defines an
-interface for translating \texttt{pandas.DataFrame}s between the IAMC format
-expected for input and output data and the calculation format used by
+used to generate sectoral emissions totals, generate regional totals, and
+combine historical emissions to native model regions (where historical data is
+defined at a higher spatial resolution than a model; see, e.g., Figure
+\ref{fig:regions}). A \texttt{FormatTranslator} class is also provided which
+defines an interface for translating \texttt{pandas.DataFrame}s between the IAM
+format expected for input and output data and the calculation format used by
 \code{aneris}' \texttt{Harmonizer}.
@@
     \caption{ Panel \ref{fig:software} shows the various objects in the
-      \code{aneris} codebase as well as a short description of their scope of
+      \code{aneris} code base as well as a short description of their scope of
       concern. Panel \ref{fig:workflow} displays the full harmonization process
@@
       scenario. Operations that can be configured with user-based input
-      configurations are shown in purple. Operations goverened by the
+      configurations are shown in purple. Operations governed by the
       \texttt{HarmonizationDriver} are shown in red. Data processing operations
@@
 
-Users are able to control the harmonization process via a number of options
-(with examples provided
-\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
-by which users control the process is by providing \textit{override} methods for
-any combination of region and variable (i.e., sector and gas species).  In
-practice, it may be possible that not all default methods chosen will provide
-robust harmonized trajectories, especially if there is a significant difference
-between historical and model values in the harmonization year, if there is
-significant upward or downward movement in the model trajectory, or if there are
-known discrepancies in sectoral definition between the IAM and historical data
-source. In such cases, override methods can ameliorate the issues associated
-with the default method choice.
-
-In order to help identify these cases, harmonization \textit{diagnostics} are
-provided which analyze the relative difference between harmonized and
-unharmonized trajectories at their mid and end-points. These values are
-configurable by the user, but defaults of $400$\% and $200$\%, respectively, are
-provided based on experiences of using \code{aneris} to date. If override
-methods are provided, they are used instead of the default methods as determined
-by the default method decision tree. Furthermore, users can set the
-above-mentioned thresholds as well as the LUC method used in the decision tree
-(see Figure \ref{fig:decision_tree}).
-
-Input data then undergoes a cleaning operation, which adds (null) model
-trajectories that exist in the historical data set but are not provided by the
-model input and detects any issues that would cause the harmonization process to
-fail. The methods used to harmonize the data are then determined and the
-harmonization process is executed. Upon completion of the harmonization process,
-aggregation of common analysis regions is performed. A common regional
-aggregation used in the IAM community was defined in the Representative
-Concentration Pathways (RCPs) \cite{vuuren_representative_2011}, shown in Figure
-\ref{fig:regions}. Finally, any exogenous trajectories the user provides are
-added. Exogenous trajectories are normally provided for unmodeled gases with
-well-accepted scenario trajectories, e.g., chlorofluorocarbons provided by WMO
-\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
-regarding the harmonization process are returned. A description of all returned
-meta data is provided in Table \ref{tab:metadata}.
+The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins by
+cleaning input data, which adds (null) model trajectories that exist in the
+historical data set but are not provided by the model input and detects any
+issues that would cause the harmonization process to fail. The methods used to
+harmonize the data are then determined and the harmonization process is
+executed. Upon completion of the harmonization process, aggregation of common
+analysis regions is performed. A common regional aggregation used in the IAM
+community was defined in the Representative Concentration Pathways (RCPs)
+\cite{vuuren_representative_2011}, shown in Figure \ref{fig:regions}. Finally,
+any exogenous trajectories the user provides are added. Exogenous trajectories
+are normally provided for unmodeled gases with well-accepted scenario
+trajectories, e.g., chlorofluorocarbons provided by WMO \cite{wmo2014}. Upon
+completion, the harmonized trajectories and meta data regarding the
+harmonization process are returned. A description of all returned meta data is
+provided in Table \ref{tab:metadata}.
 
@@
 \end{table}
+
+Users are able to control the harmonization process via a number of options
+(with examples provided
+\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
+by which users control the process is by providing \textit{override} methods for
+any combination of region and variable (i.e., sector and gas species).  In
+practice, it may be possible that not all default methods chosen will provide
+robust harmonized trajectories, especially if there is a significant difference
+between historical and model values in the harmonization year, if there is
+significant upward or downward movement in the model trajectory, or if there are
+known discrepancies in sectoral definition between the IAM and historical data
+source. In such cases, users can override default methods with their method of
+choice and both the default method and override is reported in the resulting
+metadata.
+
+In order to help identify cases where overrides may be of interest,
+harmonization \textit{diagnostics} are provided which analyze the relative
+difference between harmonized and unharmonized trajectories at their mid and
+end-points. These values are configurable by the user, but defaults of $400$\%
+and $200$\%, respectively, are provided based on experiences of the authors'
+using \code{aneris} to date.
+
 \section{Results}\label{sec:results}

### 2018-01-30 paper.tex
break the figure out into two

@@
 \usepackage{calc} % allows for arithmetic on latex variables
-\usepackage{float} % allows for figures to be placed explicitly
 \usepackage{algorithm2e} % for algorithms
@@
 components: the \texttt{HarmonizationDriver}, \texttt{Harmonizer}, and data
-processing routines (shown in red, green, and blue, respectively in Figure
-\ref{fig:software-workflow}).
+processing routines (shown in red, green, and blue, respectively in Figures
+\ref{fig:software} and \ref{fig:workflow}).
 
@@
 \begin{figure}
-  \subfloat[Python Implementation\label{fig:software}]{
-    \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
+  \begin{center}
+    \includegraphics[width=\textwidth]{implementation_schematic.pdf}
+    \caption[]{
+      \label{fig:software}
+      The various objects and their relation to one another in the \code{aneris}
+      code base as well as a short description of their scope of concern.  
     }
-  \subfloat[Workflow\label{fig:workflow}]{
-    \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
-    }
-    \caption{ Panel \ref{fig:software} shows the various objects in the
-      \code{aneris} code base as well as a short description of their scope of
-      concern. Panel \ref{fig:workflow} displays the full harmonization process
-      as executed by \code{aneris} for a single instance of a model and
-      scenario. Operations that can be configured with user-based input
-      configurations are shown in purple. Operations governed by the
-      \texttt{HarmonizationDriver} are shown in red. Data processing operations
-      are shown in blue. The core harmonization process, governed by the
+  \end{center}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[height=0.8\textheight]{aneris-workflow.pdf}
+    \caption[]{
+      \label{fig:workflow}
+The full harmonization process as executed by \code{aneris} for a single
+instance of a model and scenario. Operations that can be configured with
+user-based input configurations are shown in purple. Operations governed by the
+\texttt{HarmonizationDriver} are shown in red. Data processing operations are
+shown in blue. The core harmonization process, governed by the
 \texttt{Harmonizer} is shown in green. }
-    \label{fig:software-workflow}
+  \end{center}
 \end{figure}
+
+%% \begin{figure}
+%%   \subfloat[Python Implementation\label{fig:software}]{
+%%     \includegraphics[width=0.775\textwidth]{implementation_schematic.pdf}
+%%     }
+%%   \subfloat[Workflow\label{fig:workflow}]{
+%%     \includegraphics[width=0.225\textwidth]{aneris-workflow.pdf}
+%%     }
+%%     \caption{ Panel \ref{fig:software} shows the various objects in the
+%%       \code{aneris} code base as well as a short description of their scope of
+%%       concern. Panel \ref{fig:workflow} displays the full harmonization process
+%%       as executed by \code{aneris} for a single instance of a model and
+%%       scenario. Operations that can be configured with user-based input
+%%       configurations are shown in purple. Operations governed by the
+%%       \texttt{HarmonizationDriver} are shown in red. Data processing operations
+%%       are shown in blue. The core harmonization process, governed by the
+%%       \texttt{Harmonizer} is shown in green.  }
+%%     \label{fig:software-workflow}
+%% \end{figure}
 

### 2018-01-31 paper.tex
address most of rev 2 comments

@@
 \begin{keyword}
-Integrated Assessment Models, Climate Change, Harmonization, Air Pollution 
+Integrated Assessment Models, Harmonization, Greenhouse Gases (GHGs), Air Pollution 
 \end{keyword}
@@
 base year values (i.e., $\frac{h(t_i)}{m(t_i)}$) and those that operate on the
-offset of base year values (i.e., $h(t_i) - m(t_i)$). A number of the classic
-methods are implemented in \code{aneris} including ratio-convergence shown in
-Equation \ref{eqs:ratio}, offset-convergence shown in Equation \ref{eqs:offset},
-and interpolation shown in Equation \ref{eqs:interpolate}. The convergence
-factor, $\beta$, scales linearly from 1 to 0 over $[t_i, t_f)$ and is shown in
-  Equation \ref{eqs:factor}. In all equations the region, species, and sector
-  indices have been dropped for clarity. Each equation is a function of time,
-  model trajectory, historical trajectory, base year ($t_i$), and a convergence
-  year ($t_f$), at which point the harmonized trajectory converges to the
-  unharmonized trajectory. \codeb{aneris} provides a number of methods to choose
-  from for each of the harmonization families. A summary of all available
-  methods is provided in Table \ref{tab:meths}.
+offset of base year values (i.e., $h(t_i) - m(t_i)$). Both families of functions
+depend on a convergence factor, $\beta$, which scales linearly from 1 to 0 over
+$[t_i, t_f)$ and is shown in Equation \ref{eqs:factor}. The use of the
+  convergence factor implies that the ratio or offset applied in the base year
+  ($t_i$) decays to the unharmonized model result (i.e., the convergence factor
+  is 0) in the convergence year ($t_f$). In cases where the convergency factor
+  is applied over the entire time horizon, the convergence year is taken to be
+  $t_f = \infty$.
 
@@
 
+A number of the classic methods are implemented in \code{aneris} including
+ratio-convergence shown in Equation \ref{eqs:ratio}, offset-convergence shown in
+Equation \ref{eqs:offset}, and interpolation shown in Equation
+\ref{eqs:interpolate}. In all equations the region, species, and sector indices
+have been dropped for clarity. Each equation is a function of time, model
+trajectory, historical trajectory, base year ($t_i$), and a convergence year
+($t_f$), at which point the harmonized trajectory converges to the unharmonized
+trajectory. \codeb{aneris} provides a number of methods to choose from for each
+of the harmonization families. A summary of all available methods is provided in
+Table \ref{tab:meths}.
+
 \begin{equation}\label{eqs:ratio}
@@
 \centering
-\caption{All Harmonization Methods Provided in \code{aneris}}
+\caption{All harmonization methods provided in \code{aneris}.
+A Convergence year of $\infty$ is provided for the \code{constant\_ratio} and \code{constant\_offset} methods are listed as $\beta = 1$ for all model years in both cases.}
 \label{tab:meths}
@@
       $dH$ is defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
-      decision-making thresholds for $c_v$ and $dH$ are described below. Methods
-      labeled in \textit{green} are likely to closely match unharmonized
-      results, methods in \textit{yellow} will likely somewhat match
-      unharmonized results, and methods in \textit{red} can be expected to have
-      a large relative difference between harmonized and unharmonized results.}
+      decision-making thresholds for $c_v$ and $dH$ are described below.  Where
+      present, convergence years of default methods are provided below the
+      method name in parentheses.  Methods labeled in \textit{green} are likely
+      to closely match unharmonized results, methods in \textit{yellow} will
+      likely somewhat match unharmonized results, and methods in \textit{red}
+      can be expected to have a large relative difference between harmonized and
+      unharmonized results.}
   \end{center}
@@
 In most cases, however, models do report values in the harmonization
-year. Figure \ref{fig:cases} displays a number of example trajectories which
-highlight the possible issues resulting from harmonizing model results in
-different contexts. When model and historical values are relatively close, a
-convergence method is chosen in order to be as representative as possible to the
-underlying unharmonized model results (Figure \ref{fig:cases}, Panel
-\code{a}). If values are not close, the constant ratio method is chosen in order
-to provide reasonable trajectories that still incorporate modeled effects
-(Figure \ref{fig:cases}, Panel \code{b}).
-
-Models can additionally report negative emissions in certain contexts which must
-be taken into account during harmonization. Such a case is possible for gas
-species which can be extracted from the environment and stored, as is the case
-for \cotwo in future scenarios with climate mitigation. If a model provides a
-trajectory that transitions from positive to negative emissions and base year
-results are similar, then a convergence method is used in order to guarantee
-capture of this transition in a representative fashion (Figure \ref{fig:cases},
-Panel \code{c}). If the discrepancy in base year results is large, it is
-possible for a negative trajectory to be inappropriately harmonized to a
-positive, but decreasing, trajectory. As such, the constant ratio method is
-chosen (Figure \ref{fig:cases}, Panel \code{d}).
+year. Figure \ref{fig:cases} displays a number of illustrative example
+trajectories which highlight the possible issues resulting from harmonizing
+model results in different contexts. Panels \code{a} and \code{b} highlight
+examples where model results peak mid-century, behavior that is seen in a number
+of scenarios with general emissions mitigation effects, such as pollution
+controls applied by developing nations on transport and industry sectors. Panel
+\code{a} highlights a case where model base-year values and history are relative
+close whereas Panel \code{b} shows a situation where model values and history
+are relatively far apart. Panels \code{c} and \code{d} show similar model
+trajectories that peak mid-century but also have negative emissions. Models can
+report negative emissions in contexts in which a gas species which can be
+extracted from the environment and stored, as is the case for \cotwo in future
+scenarios with climate mitigation enacted via the deployment of carbon capture
+and sequestration technologies or direct carbon capture technologies. Again, the
+relative difference with historical values differ between the panels to explore
+harmonization method choices in each situation.
+
+When model and historical values are relatively close, a convergence method is
+chosen in order to be as representative as possible to the underlying
+unharmonized model results (Figure \ref{fig:cases}, Panel \code{a}). If values
+are not close, the constant ratio method is chosen in order to provide
+reasonable trajectories that still incorporate modeled effects (Figure
+\ref{fig:cases}, Panel \code{b}).
+
+If a model provides a trajectory that transitions from positive to negative
+emissions and base year results are similar, then a convergence method is used
+in order to guarantee capture of this transition in a representative fashion
+(Figure \ref{fig:cases}, Panel \code{c}). If the discrepancy in base year
+results is large, it is possible for a negative trajectory to be inappropriately
+harmonized to a positive, but decreasing, trajectory. As such, the constant
+ratio method is chosen (Figure \ref{fig:cases}, Panel \code{d}).
 
@@
 standard deviation, $\sigma$, and the mean, $\mu$, as shown in Equation
-\ref{eqs:cov}.
+\ref{eqs:cov}. For a single realiziation of $c_v$, the first derivative
+information of the entire historical time period is utilized.
+
+\begin{equation}\label{eqs:cov}
+    c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
+\end{equation}
 
@@
 and non-LUC $c_v$s as determined for historical data aggregated to the model
-regions of 5 different IAMs involved in the SSP process. A threshold value of
-$\tau_{c_v} = 20$ has been chosen based on these observations as it
-optimally divides the two distributions. Importantly, tails of the LUC and
-non-LUC overlap, thus there are both false positives (~7\% of non-LUC
-trajectories) and false negatives (~10\% of LUC trajectories). However, as any
-regional definition is model dependent and thus any regional aggregation is
-possible an automated detection mechanism is necessary.
-
-\begin{equation}\label{eqs:cov}
-    c_v =  \frac{\sigma(h^{\prime}(t))}{\mu(h^{\prime}(t))}
-\end{equation}
+regions of 5 different IAMs involved in the SSP process: AIM-CGE, IMAGE, GCAM4,
+MESSAGE-GLOBIOM, and REMIND-MAGPIE \TODO{add refs}, each of which have varying
+definitions of native model regions comprising different collections of
+countries. Therefore, each data point comprising Figure \ref{fig:cov} represents
+a realization of $c_v$ for a single combination of native model region, sector,
+and emissions species \footnote{A full listing of all sectors and species is
+  presented in the case-study discussion in Section \ref{sec:results}, Table
+  \ref{tab:sp}}. A threshold value of $\tau_{c_v} = 20$ has been chosen based on
+these observations as it optimally divides the two distributions. Importantly,
+tails of the LUC and non-LUC overlap, thus there are both false positives (~7\%
+of non-LUC trajectories) and false negatives (~10\% of LUC
+trajectories). However, as any regional definition is model dependent and thus
+any regional aggregation is possible an automated detection mechanism is
+necessary.
 
@@
       is used for non-LUC data and \cite{van_marle_historic_2017} is used for
-      LUC data. All historical data has been aggregated from countries to IAM
-      model regional definitions, and all gas species included in the historical
-      data sets are included in the analysis. The solid black line indicates the
-      threshold value, $\tau_{c_v}$, used by default in \code{aneris}.  }
+      LUC data. All historical data has been aggregated from their native
+      spatial resolution (i.e., individual countries) to IAM model regional
+      definitions (i.e., collections of countries), and all gas species included
+      in the historical data sets are included in the analysis. The solid black
+      line indicates the threshold value, $\tau_{c_v}$, used by default in
+      \code{aneris}.  }
   \end{center}
@@
 
-\section{Results}\label{sec:results}
+\section{Case Study: Harmonizing Results from a Global IAM}\label{sec:results}
 
@@
 technologies and policies are enacted causing a general reduction in pollutants
-and GHGs, including (eventual) negative CO2 emissions in some regions and
-sectors due to carbon capture and sequestration and afforestation. Figure
-\ref{fig:kyoto} shows the different trends of Kyoto Gases in each scenarios.
+and GHGs, including (eventual) negative \cotwo emissions in some regions and
+sectors due to carbon capture and sequestration and afforestation. A scenario in
+which negative emissions play a role in mitigation strategies is particularly
+important because of the sensitivity of key indicators, such as end-of-century
+radiative forcing (which is used to estimate mean global temperature increase),
+to the timing and magnitude of net-zero and total negative \cotwo
+emissions. Therefore, these two scenarios represent two relative extreme cases
+in the use of a harmonization approach and thus provide a case study as to its
+general applicability. Figure \ref{fig:kyoto} shows the different trends of
+Kyoto Gases in each scenarios.
 
@@
 harmonization method for the SAS region, the global trajectory for industrial CO
-also is representative the trends seen in the unharmonized trajectory and the
+also is representative of the trends seen in the unharmonized trajectory and the
 relative importance of the underlying regional trajectories is maintained.

### 2018-01-31 paper.tex
initial show for plurarl iams

@@
 
+%%% Plural gls entries
+\newacronym[
+  \glslongpluralkey={Integrated Assessment Models},
+  \glsshortpluralkey={IAMs}
+]{iam}{IAM}{Integrated Assessment Model}
+
 %%% Acronyms
@@
 \newacronym{ceds}{CEDS}{Community Emissions Data System}
-\newacronym{iam}{IAM}{Integrated Assessment Model}
 \newacronym{esm}{ESM}{Earth System Model}
@@
 
-%% all model names
-%% all gases
-
 \makeglossaries
@@
 separately by individual modeling teams. For the hand-over of emission data for
-the \gls{ssps} to climate model groups, a new automated approach based on
-commonly agreed upon algorithms was developed. This work describes the novel
-methodology for determining such harmonization methods and an open-source Python
-software library implementing the methodology. Results are shown for two example
-scenarios (with and without climate policy cases) using the MESSAGE-GLOBIOM IAM
-that satisfactorily harmonize over 96\% of the total emissions trajectories
-while having a negligible effect on key long-term climate indicators. This new
-capability enhances the comparability across different models, increases
-transparency and robustness of results, and allows other teams to easily
-participate in intercomparison exercises by using the same, openly available
-harmonization mechanism.
+the Shared Socioeconomic Pathways (SSPs) to climate model groups, a new
+automated approach based on commonly agreed upon algorithms was developed. This
+work describes the novel methodology for determining such harmonization methods
+and an open-source Python software library implementing the methodology. Results
+are shown for two example scenarios (with and without climate policy cases)
+using the IAM MESSAGE-GLOBIOM that satisfactorily harmonize over 96\% of the
+total emissions trajectories while having a negligible effect on key long-term
+climate indicators. This new capability enhances the comparability across
+different models, increases transparency and robustness of results, and allows
+other teams to easily participate in intercomparison exercises by using the
+same, openly available harmonization mechanism.
 \end{abstract}
@@
 
-Integrated Assessment Models (IAMs) are tools used to understand the complex
+\glspl{iam} are tools used to understand the complex
 interactions between energy, economy, land use, water, and climate systems. IAMs

### 2018-01-31 paper.tex
glspl for IAMs

@@
 \glspl{iam} are tools used to understand the complex
-interactions between energy, economy, land use, water, and climate systems. IAMs
+interactions between energy, economy, land use, water, and climate systems. \glspl{iam}
 provide global projections of systemic change by dividing the world into a
 number of representative regions (typically 10 to 30), the definition of which
-is distinct for each model \cite{krey_global_2014}. Results from IAMs are
+is distinct for each model \cite{krey_global_2014}. Results from \glspl{iam} are
 integral in a number of international studies, which notably include projections
@@
 
-While IAMs are implemented in myriad ways, including simulation and
+While \glspl{iam} are implemented in myriad ways, including simulation and
 optimization, the core inputs and outputs are similar across different
@@
 
-The emissions trajectories calculated by IAMs are critical inputs for ongoing,
+The emissions trajectories calculated by \glspl{iam} are critical inputs for ongoing,
 worldwide scientific community efforts in the Coupled Model Intercomparison
@@
 
-In practice, IAMs calculate the total source intensity of emitting technologies,
+In practice, \glspl{iam} calculate the total source intensity of emitting technologies,
 for example the total activity of coal power plants in China, and incorporate
@@
 Climate modeling efforts have continued to progress, demanding increased spatial
-and sectoral resolution from IAMs. Furthermore, a new generation of climate
+and sectoral resolution from \glspl{iam}. Furthermore, a new generation of climate
 scenarios which combines aspects of both the RCPs and SSPs have been developed
@@
 combination of meteorological conditions and anthropogenic drivers. Land use
-emissions in many IAMs are modeled using average emission factors and do not
+emissions in many \glspl{iam} are modeled using average emission factors and do not
 capture conditions in a specific year. A longer convergence horizon is thus
@@
 and non-LUC $c_v$s as determined for historical data aggregated to the model
-regions of 5 different IAMs involved in the SSP process: AIM-CGE, IMAGE, GCAM4,
+regions of 5 different \glspl{iam} involved in the SSP process: AIM-CGE, IMAGE, GCAM4,
 MESSAGE-GLOBIOM, and REMIND-MAGPIE \TODO{add refs}, each of which have varying
@@
 This work presented a novel methodology and Python implementation of automated
-emissions harmonization for IAMs, \code{aneris}. An in-depth explanation of the
+emissions harmonization for \glspl{iam}, \code{aneris}. An in-depth explanation of the
 processes and methods for determining the use of harmonization methods was

### 2018-01-31 paper.tex
finish acs for iam

@@
 integral in a number of international studies, which notably include projections
-of climate and economic futures. Recently, the IAM community has developed
+of climate and economic futures. Recently, the \gls{iam} community has developed
 scenarios based on the \gls{ssps}
@@
 Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
-of marker SSP scenarios developed by the IAM community (ScenarioMIP)
+of marker SSP scenarios developed by the \gls{iam} community (ScenarioMIP)
 \cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
@@
 Harmonization has been addressed in previous studies as it is a common practice
-in the IAM and climate change communities. For example,
+in the \gls{iam} and climate change communities. For example,
 \cite{meinshausen_rcp_2011} describes the use of scaling routines for the 5
@@
 harmonization routines on future trajectories. During the evaluation of the
-Representative Concentration Pathways (RCPs), IAM results have been harmonized
+Representative Concentration Pathways (RCPs), \gls{iam} results have been harmonized
 by sector and the 5 RCP global regions
@@
 \code{aneris} as well as the procedural workflow. The results of applying the
-automated harmonization mechanism on two example IAM scenarios, one with
+automated harmonization mechanism on two example \gls{iam} scenarios, one with
 emissions growth and another with emissions mitigation, is presented in Section
@@
 
-IAM emission results are provided along temporal (normally half decade or
+\gls{iam} emission results are provided along temporal (normally half decade or
 decade), spatial (i.e., model regions), gas species, and sectoral
@@
 harmonization year are analyzed. The decision tree used in this analysis is a
-result of collaborative efforts between IAM teams and is shown graphically in
+result of collaborative efforts between \gls{iam} teams and is shown graphically in
 Figure \ref{fig:decision_tree}.
@@
       LUC data. All historical data has been aggregated from their native
-      spatial resolution (i.e., individual countries) to IAM model regional
+      spatial resolution (i.e., individual countries) to \gls{iam} model regional
       definitions (i.e., collections of countries), and all gas species included
@@
 investigate the possible values that these relative differences can take, the
-IAM values used in the SSP and (ongoing) CMIP6 inter-comparison
+\gls{iam} values used in the SSP and (ongoing) CMIP6 inter-comparison
 exercises are used. A distribution of these differences for all models in the
@@
 results) is assumed to be an Excel file in the standard data format within the
-IAM community, i.e., with \code{Model}, \code{Scenario}, \code{Region},
+\gls{iam} community, i.e., with \code{Model}, \code{Scenario}, \code{Region},
 \code{Variable}, and \code{Unit} columns in addition to columns representing
@@
 
-    Harmonize historical trajectories to data in the IAMC template format.
+    Harmonize historical trajectories to data in the IAM template format.
 
@@
 the emissions species from the sector contributing the emissions (these are
-combined in the single \code{Variable} column in the standard IAM
+combined in the single \code{Variable} column in the standard \gls{iam}
 format). Because the \texttt{Harmonizer} is designed to operate on a single
@@
 \ref{fig:regions}). A \texttt{FormatTranslator} class is also provided which
-defines an interface for translating \texttt{pandas.DataFrame}s between the IAM
+defines an interface for translating \texttt{pandas.DataFrame}s between the \gls{iam}
 format expected for input and output data and the calculation format used by
@@
 executed. Upon completion of the harmonization process, aggregation of common
-analysis regions is performed. A common regional aggregation used in the IAM
+analysis regions is performed. A common regional aggregation used in the \gls{iam}
 community was defined in the Representative Concentration Pathways (RCPs)
@@
 significant upward or downward movement in the model trajectory, or if there are
-known discrepancies in sectoral definition between the IAM and historical data
+known discrepancies in sectoral definition between the \gls{iam} and historical data
 source. In such cases, users can override default methods with their method of
@@
 \code{aneris} harmonization procedure, we focus on the harmonization of results
-of the IAM MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the SSP
+of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the SSP
 scenario library \cite{Riahi2017153,Rao2017346} are presented. The SSP2, or
@@
 The use of an open-source, automated harmonization process also provides
-benefits to the wider climate science and IAM communities. By providing a
-standard mechanism for harmonization, the IAM community can directly provide
+benefits to the wider climate science and \gls{iam} communities. By providing a
+standard mechanism for harmonization, the \gls{iam} community can directly provide
 input into the harmonization algorithms and rules for their default
@@
 
-This work provides a new direction and framework which the IAM and climate
+This work provides a new direction and framework which the \gls{iam} and climate
 communities can build upon in order to reduce the necessity of consistent expert

### 2018-01-31 paper.tex
transferred all acronyms

@@
 ]{iam}{IAM}{Integrated Assessment Model}
+\newacronym[
+  \glslongpluralkey={Shared Socioeconomic Pathways},
+  \glsshortpluralkey={SSPs}
+]{ssp}{SSP}{Shared Socioeconomic Pathway}
+\newacronym[
+  \glslongpluralkey={Representative Concentration Pathways},
+  \glsshortpluralkey={RCPs}
+]{rcp}{RCP}{Representative Concentration Pathway}
+\newacronym[
+  \glslongpluralkey={Greenhouse Gases},
+  \glsshortpluralkey={GHGs}
+]{ghg}{GHG}{Greenhouse Gas}
 
 %%% Acronyms
-\newacronym{ssps}{SSPs}{Shared Socioeconomic Pathways}
-\newacronym{rcps}{RCPs}{Representative Concentration Pathways}
+\newacronym{sres}{SRES}{Special Report on Emissions Scenarios}
 \newacronym{cmip6}{CMIP6}{Coupled Model Intercomparison Project (Phase 6)}
@@
 \newacronym{esm}{ESM}{Earth System Model}
-\newacronym{ghg}{GHG}{Greenhouse Gas}
 \newacronym{luc}{LUC}{Land-use Change}
@@
 of climate and economic futures. Recently, the \gls{iam} community has developed
-scenarios based on the \gls{ssps}
+scenarios based on the \glspl{ssp}
 \cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
@@
 potential global futures. The SSPs are designed to be used in research that
-include earth system model (ESM) simulations, climate impact, adaptation and
+include \gls{esm} simulations, climate impact, adaptation and
 climate mitigation studies \cite{vuuren_new_2013}.
@@
 
-The emissions trajectories calculated by \glspl{iam} are critical inputs for ongoing,
-worldwide scientific community efforts in the Coupled Model Intercomparison
-Project Phase 6 (CMIP6) \cite{eyring_overview_2016}, which is utilizing a number
-of marker SSP scenarios developed by the \gls{iam} community (ScenarioMIP)
+The emissions trajectories calculated by \glspl{iam} are critical inputs for
+ongoing, worldwide scientific community efforts in the \gls{cmip6}
+\cite{eyring_overview_2016}, which is utilizing a number of marker SSP scenarios
+developed by the \gls{iam} community (\gls{smip})
 \cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
@@
 historical emissions data set for both anthropogenic emissions (i.e., the
-Community Emissions Data System (CEDS) \cite{hoesly_historical_2017} and
-open-burning land-use and land-use change (LULUC) emissions
+\gls{ceds} \cite{hoesly_historical_2017} and
+open-burning \gls{luluc} emissions
 \cite{van_marle_historic_2017}) which, in conjunction with the SSP IAM
-trajectories, will be used for climate-related modeling exercises of CMIP6.
+trajectories, will be used for climate-related modeling exercises of \gls{cmip6}.
 
 When participating in intercomparison exercises in which a consistent historical
-starting point is required (e.g., in CMIP6), model teams incorporate a single,
+starting point is required (e.g., in \gls{cmip6}), model teams incorporate a single,
 common historical data set through \textit{harmonization}. Harmonization
@@
 \cite{meinshausen_rcp_2011} describes the use of scaling routines for the 5
-regions used in the Special Report on Emissions Scenarios (SRES)
-\cite{nakicenovic2000}; however, only total emissions were harmonized in the
-exercise, thus there is no sectoral dimension. Further,
-\cite{rogelj_discrepancies_2011} describes the impacts of choosing various
-harmonization routines on future trajectories. During the evaluation of the
-Representative Concentration Pathways (RCPs), \gls{iam} results have been harmonized
-by sector and the 5 RCP global regions
+regions used in the IPCC \gls{sres} \cite{nakicenovic2000}; however, only total
+emissions were harmonized in the exercise, thus there is no sectoral
+dimension. Further, \cite{rogelj_discrepancies_2011} describes the impacts of
+choosing various harmonization routines on future trajectories. During the
+evaluation of the Representative Concentration Pathways (RCPs), \gls{iam}
+results have been harmonized by sector and the 5 RCP global regions
 \cite{vuuren_representative_2011}. Importantly, the choice of harmonization
@@
 The value of $c_v$ is then tested against a threshold, $\tau_{c_v}$. To
-determine this threshold, an analysis of the recent CEDS and LUC historical data
-has been performed. Figure \ref{fig:cov} shows the distribution of LUC $c_v$s
-and non-LUC $c_v$s as determined for historical data aggregated to the model
+determine this threshold, an analysis of the recent \gls{ceds} and \gls{luc} historical data
+has been performed. Figure \ref{fig:cov} shows the distribution of \gls{luc} $c_v$s
+and non-\gls{luc} $c_v$s as determined for historical data aggregated to the model
 regions of 5 different \glspl{iam} involved in the SSP process: AIM-CGE, IMAGE, GCAM4,
@@
 these observations as it optimally divides the two distributions. Importantly,
-tails of the LUC and non-LUC overlap, thus there are both false positives (~7\%
-of non-LUC trajectories) and false negatives (~10\% of LUC
+tails of the \gls{luc} and non-\gls{luc} overlap, thus there are both false positives (~7\%
+of non-\gls{luc} trajectories) and false negatives (~10\% of \gls{luc}
 trajectories). However, as any regional definition is model dependent and thus
@@
       \label{fig:cov}
-      The distribution of $c_v$ values for LUC and non-LUC historical
-      trajectories is shown. CEDS historical data \cite{hoesly_historical_2017}
-      is used for non-LUC data and \cite{van_marle_historic_2017} is used for
-      LUC data. All historical data has been aggregated from their native
+      The distribution of $c_v$ values for \gls{luc} and non-\gls{luc} historical
+      trajectories is shown. \gls{ceds} historical data \cite{hoesly_historical_2017}
+      is used for non-\gls{luc} data and \cite{van_marle_historic_2017} is used for
+      \gls{luc} data. All historical data has been aggregated from their native
       spatial resolution (i.e., individual countries) to \gls{iam} model regional
@@
 investigate the possible values that these relative differences can take, the
-\gls{iam} values used in the SSP and (ongoing) CMIP6 inter-comparison
+\gls{iam} values used in the SSP and (ongoing) \gls{cmip6} inter-comparison
 exercises are used. A distribution of these differences for all models in the
@@
       The 5 regions used in the RCPs with their 11-region constituents: Asia
-      (Central Asia, South Asia, Pacific Asia) [yellows], Latin America
-      [magenta], the Middle East and Africa [greens], the OECD (North America,
-      Western Europe, and Pacific OECD) [blues], and the Reforming Economies
-      (Eastern Europe and Former Soviet Union) [reds].  }
+      (\gls{cpa}, \gls{sas}, \gls{pas}) [yellows], \gls{lam}
+      [magenta], the \gls{mea} and \gls{afr} [greens], the OECD (\gls{nam},
+      \gls{weu}, and \gls{pao}) [blues], and the Reforming Economies
+      (\gls{eeu} and \gls{fsu}) [reds].  }
   \end{center}
@@
 mapped directly to the 5-region definition used in the RCPs. Historical data is
-taken from previously described LUC and anthropogenic sources, which comprise 10
+taken from previously described \gls{luc} and anthropogenic sources, which comprise 10
 separate pollutant and GHG species and 12 sectors shown in Table \ref{tab:sp}. A
@@
 acceptable tolerances (e.g., 0.75 $\frac{\text{W}}{\text{m}^2}$ defined for
-ScenarioMIP \cite{oneill_scenario_2016}). Thus harmonization is considered to
+\gls{smip} \cite{oneill_scenario_2016}). Thus harmonization is considered to
 have a negligible effect on key long-term climate indicators.

### 2018-01-31 paper.tex
rcps updated

@@
 choosing various harmonization routines on future trajectories. During the
-evaluation of the Representative Concentration Pathways (RCPs), \gls{iam}
-results have been harmonized by sector and the 5 RCP global regions
+evaluation of the \glspl{rcp}, \gls{iam}
+results have been harmonized by sector and the 5 \gls{rcp} global regions
 \cite{vuuren_representative_2011}. Importantly, the choice of harmonization
@@
 and sectoral resolution from \glspl{iam}. Furthermore, a new generation of climate
-scenarios which combines aspects of both the RCPs and SSPs have been developed
+scenarios which combines aspects of both the \glspl{rcp} and SSPs have been developed
 in order to incorporate both physical and socio-economic detail. In order to
@@
 
-The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins by
-cleaning input data, which adds (null) model trajectories that exist in the
+The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins
+by cleaning input data, which adds (null) model trajectories that exist in the
 historical data set but are not provided by the model input and detects any
@@
 executed. Upon completion of the harmonization process, aggregation of common
-analysis regions is performed. A common regional aggregation used in the \gls{iam}
-community was defined in the Representative Concentration Pathways (RCPs)
+analysis regions is performed. A common regional aggregation used in the
+\gls{iam} community was defined in the \glspl{rcp}
 \cite{vuuren_representative_2011}, shown in Figure \ref{fig:regions}. Finally,
@@
       \label{fig:regions}
-      The 5 regions used in the RCPs with their 11-region constituents: Asia
+      The 5 regions used in the \glspl{rcp} with their 11-region constituents: Asia
       (\gls{cpa}, \gls{sas}, \gls{pas}) [yellows], \gls{lam}
@@
 analyzed because MESSAGE-GLOBIOM is the marker scenario for this SSP. This SSP2
-scenario lies between two RCPs, 6 and 8.5, with a radiative forcing level of
+scenario lies between two \glspl{rcp}, 6 and 8.5, with a radiative forcing level of
 approximately 6.5 Wm$^{-2}$. We additionally present the results for the
@@
 MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
-mapped directly to the 5-region definition used in the RCPs. Historical data is
+mapped directly to the 5-region definition used in the \glspl{rcp}. Historical data is
 taken from previously described \gls{luc} and anthropogenic sources, which comprise 10

### 2018-01-31 paper.tex
acs for ssps

@@
 
-\glspl{iam} are tools used to understand the complex
-interactions between energy, economy, land use, water, and climate systems. \glspl{iam}
-provide global projections of systemic change by dividing the world into a
-number of representative regions (typically 10 to 30), the definition of which
-is distinct for each model \cite{krey_global_2014}. Results from \glspl{iam} are
-integral in a number of international studies, which notably include projections
-of climate and economic futures. Recently, the \gls{iam} community has developed
-scenarios based on the \glspl{ssp}
-\cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
-  calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which quantify a variety of
-potential global futures. The SSPs are designed to be used in research that
-include \gls{esm} simulations, climate impact, adaptation and
-climate mitigation studies \cite{vuuren_new_2013}.
+\glspl{iam} are tools used to understand the complex interactions between
+energy, economy, land use, water, and climate systems. \glspl{iam} provide
+global projections of systemic change by dividing the world into a number of
+representative regions (typically 10 to 30), the definition of which is distinct
+for each model \cite{krey_global_2014}. Results from \glspl{iam} are integral in
+a number of international studies, which notably include projections of climate
+and economic futures. Recently, the \gls{iam} community has developed scenarios
+based on the \glspl{ssp} \cite{van_vuuren_energy_2017, fricko_marker_2017,
+  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
+quantify a variety of potential global futures. The \glspl{ssp} are designed to
+be used in research that include \gls{esm} simulations, climate impact,
+adaptation and climate mitigation studies \cite{vuuren_new_2013}.
 
@@
 various socio-economic and technological assumptions as well as proposed policy
-constraints, e.g., targets for future Greenhouse Gas (GHG) emissions.
+constraints, e.g., targets for future \gls{ghg} emissions.
 
@@
 ongoing, worldwide scientific community efforts in the \gls{cmip6}
-\cite{eyring_overview_2016}, which is utilizing a number of marker SSP scenarios
+\cite{eyring_overview_2016}, which is utilizing a number of marker \gls{ssp} scenarios
 developed by the \gls{iam} community (\gls{smip})
@@
 modeling the individual technologies and sectors that contribute towards the
-emissions of different air pollutants and GHGs as well as various mitigation
-technologies. However, the historical emissions starting points of models can
-differ by large amounts depending on the region, sector, and emissions species.
-
-In practice, \glspl{iam} calculate the total source intensity of emitting technologies,
-for example the total activity of coal power plants in China, and incorporate
-emissions-intensity factors for individual gas species, for example the quantity
-of sulfur emissions from coal plants per megawatt-hour of production. Models are
-generally \textit{calibrated} to historical data sources in one or more base
-years. Results in the historical period may differ between models as a result of
-the sometimes large uncertainties in historical data sets. Models can also
-differ in their choice of base-year, which may lag behind available inventory
-data. In addition, models have varying sectoral, regional, and fuel
-aggregations.
+emissions of different air pollutants and \glspl{ghg} as well as various
+mitigation technologies. However, the historical emissions starting points of
+models can differ by large amounts depending on the region, sector, and
+emissions species.
+
+In practice, \glspl{iam} calculate the total source intensity of emitting
+technologies, for example the total activity of coal power plants in China, and
+incorporate emissions-intensity factors for individual gas species, for example
+the quantity of sulfur emissions from coal plants per megawatt-hour of
+production. Models are generally \textit{calibrated} to historical data sources
+in one or more base years. Results in the historical period may differ between
+models as a result of the sometimes large uncertainties in historical data
+sets. Models can also differ in their choice of base-year, which may lag behind
+available inventory data. In addition, models have varying sectoral, regional,
+and fuel aggregations.
 
@@
 historical emissions data set for both anthropogenic emissions (i.e., the
-\gls{ceds} \cite{hoesly_historical_2017} and
-open-burning \gls{luluc} emissions
-\cite{van_marle_historic_2017}) which, in conjunction with the SSP IAM
-trajectories, will be used for climate-related modeling exercises of \gls{cmip6}.
+\gls{ceds} \cite{hoesly_historical_2017} and open-burning \gls{luluc} emissions
+\cite{van_marle_historic_2017}) which, in conjunction with the \gls{ssp}
+\gls{iam} trajectories, will be used for climate-related modeling exercises of
+\gls{cmip6}.
 
@@
 and sectoral resolution from \glspl{iam}. Furthermore, a new generation of climate
-scenarios which combines aspects of both the \glspl{rcp} and SSPs have been developed
+scenarios which combines aspects of both the \glspl{rcp} and \glspl{ssp} have been developed
 in order to incorporate both physical and socio-economic detail. In order to
@@
 and non-\gls{luc} $c_v$s as determined for historical data aggregated to the model
-regions of 5 different \glspl{iam} involved in the SSP process: AIM-CGE, IMAGE, GCAM4,
+regions of 5 different \glspl{iam} involved in the \gls{ssp} process: AIM-CGE, IMAGE, GCAM4,
 MESSAGE-GLOBIOM, and REMIND-MAGPIE \TODO{add refs}, each of which have varying
@@
 investigate the possible values that these relative differences can take, the
-\gls{iam} values used in the SSP and (ongoing) \gls{cmip6} inter-comparison
+\gls{iam} values used in the \gls{ssp} and (ongoing) \gls{cmip6} inter-comparison
 exercises are used. A distribution of these differences for all models in the
@@
 \code{aneris} harmonization procedure, we focus on the harmonization of results
-of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the SSP
+of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the \gls{ssp}
 scenario library \cite{Riahi2017153,Rao2017346} are presented. The SSP2, or
 ``middle of the road'', scenario (referred to as SSP2-Ref) is chosen to be
-analyzed because MESSAGE-GLOBIOM is the marker scenario for this SSP. This SSP2
+analyzed because MESSAGE-GLOBIOM is the marker scenario for this \gls{ssp}. This SSP2
 scenario lies between two \glspl{rcp}, 6 and 8.5, with a radiative forcing level of
@@
 technologies and policies are enacted causing a general reduction in pollutants
-and GHGs, including (eventual) negative \cotwo emissions in some regions and
+and \glspl{ghg}, including (eventual) negative \cotwo emissions in some regions and
 sectors due to carbon capture and sequestration and afforestation. A scenario in
@@
 taken from previously described \gls{luc} and anthropogenic sources, which comprise 10
-separate pollutant and GHG species and 12 sectors shown in Table \ref{tab:sp}. A
+separate pollutant and \gls{ghg} species and 12 sectors shown in Table \ref{tab:sp}. A
 total of 970 distinct trajectories were harmonized for each scenario, and

### 2018-01-31 paper.tex
add wmo

@@
 \newacronym{cli}{CLI}{Command Line Interface}
+\newacronym{wmo}{WMO}{World Meteorological Organization}
 
@@
 The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins
-by cleaning input data, which adds (null) model trajectories that exist in the
-historical data set but are not provided by the model input and detects any
+by cleaning input data, which adds (0-valued) model trajectories that exist in
+the historical data set but are not provided by the model input and detects any
 issues that would cause the harmonization process to fail. The methods used to
@@
 are normally provided for unmodeled gases with well-accepted scenario
-trajectories, e.g., chlorofluorocarbons provided by WMO \cite{wmo2014}. Upon
-completion, the harmonized trajectories and meta data regarding the
-harmonization process are returned. A description of all returned meta data is
-provided in Table \ref{tab:metadata}.
+trajectories, e.g., chlorofluorocarbons provided by the \gls{wmo}
+\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
+regarding the harmonization process are returned. A description of all returned
+meta data is provided in Table \ref{tab:metadata}.
 

### 2018-01-31 paper.tex
finished most of reviewer 3 changes

@@
 \code{aneris} harmonization procedure, we focus on the harmonization of results
-of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from the \gls{ssp}
-scenario library \cite{Riahi2017153,Rao2017346} are presented. The SSP2, or
-``middle of the road'', scenario (referred to as SSP2-Ref) is chosen to be
-analyzed because MESSAGE-GLOBIOM is the marker scenario for this \gls{ssp}. This SSP2
-scenario lies between two \glspl{rcp}, 6 and 8.5, with a radiative forcing level of
-approximately 6.5 Wm$^{-2}$. We additionally present the results for the
-SSP2-based mitigation scenario leading to a radiative forcing of 4.5 Wm$^{-2}$
-(referred to as SSP2-4.5). The SSP2-45 scenario is chosen because mitigation
-technologies and policies are enacted causing a general reduction in pollutants
-and \glspl{ghg}, including (eventual) negative \cotwo emissions in some regions and
-sectors due to carbon capture and sequestration and afforestation. A scenario in
-which negative emissions play a role in mitigation strategies is particularly
-important because of the sensitivity of key indicators, such as end-of-century
-radiative forcing (which is used to estimate mean global temperature increase),
-to the timing and magnitude of net-zero and total negative \cotwo
-emissions. Therefore, these two scenarios represent two relative extreme cases
-in the use of a harmonization approach and thus provide a case study as to its
-general applicability. Figure \ref{fig:kyoto} shows the different trends of
-Kyoto Gases in each scenarios.
+of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from
+the \gls{ssp} scenario library\footnote{We refer the reader to the broad
+  literature discussing the SSPs, e.g.,
+  \cite{vuuren_new_2013,oneill_scenario_2016, oneill_new_2013} for a more in
+  depth discussion of the scenario architecture and design.} are presented. The
+SSP2 \cite{Riahi2017153,Rao2017346}, or ``middle of the road'', scenario
+(referred to as SSP2-Ref) is chosen to be analyzed because MESSAGE-GLOBIOM is
+the marker scenario\footnote{The concept of ``marker'' scenarios are used to
+  illustrate distinctive scenario types within the SSP scenario framework. See
+  \cite{vuuren_new_2013} and \cite{Riahi2017153} for a more lengthy
+  description.} for this \gls{ssp}. This SSP2 scenario lies between two
+\glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing is a
+  measurement of the capacity for different emissions species to affect
+  tropopausal irradiance and has a relationship to global mean equilibrium
+  temperature change. We refer the reader to \cite{IPCCWG1PhysicalStocker2013}
+  for a more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
+additionally present the results for the SSP2-based mitigation scenario leading
+to a radiative forcing of 4.5 Wm$^{-2}$ (referred to as SSP2-4.5). The SSP2-45
+scenario is chosen because mitigation technologies and policies are enacted
+causing a general reduction in pollutants and \glspl{ghg}, including (eventual)
+negative \cotwo emissions in some regions and sectors due to carbon capture and
+sequestration and afforestation. A scenario in which negative emissions play a
+role in mitigation strategies is particularly important because of the
+sensitivity of key indicators, such as end-of-century radiative forcing (which
+is used to estimate mean global temperature increase), to the timing and
+magnitude of net-zero and total negative \cotwo emissions. Therefore, these two
+scenarios represent two relative extreme cases in the use of a harmonization
+approach and thus provide a case study as to its general applicability. Figure
+\ref{fig:kyoto} shows the different trends of Kyoto Gases in each scenarios.
 
@@
 methods on total anthropogenic radiative forcing projections with the simple
-carbon-cycle and climate model, MAGICC6 \cite{meinshausen2011emulating,
+carbon-cycle and climate model, MAGICC\footnote{MAGICC is a reduced complexity
+  climate model which incorporates future trajectories of forcing agents (i.e.,
+  emissions) to estimate future radiative forcing and mean global temperature
+  response. We refer the reader to the
+  \href{http://wiki.magicc.org/index.php?title=Model_Description}{MAGICC wiki}
+  for a more in-depth description.} \cite{meinshausen2011emulating,
   meinshausen2011rcp}, for each harmonized and unharmonized scenario
 respectively as shown in Figure \ref{fig:forcing}. We find that the change due
-to harmonization is small, ranging between 1 and 2.5\% over the model
-horizon. Relative near-term differences persist in the mitigation case (SSP-4.5)
+to harmonization is small, ranging between 1 and 2.5\% over the modeled time
+period. Relative near-term differences persist in the mitigation case (SSP-4.5)
 because differences in near-term emissions define to a larger degree the
@@
 \end{figure}
+
 \section{Discussion \& Future Work}\label{sec:future}
@@
 This work presented a novel methodology and Python implementation of automated
-emissions harmonization for \glspl{iam}, \code{aneris}. An in-depth explanation of the
-processes and methods for determining the use of harmonization methods was
-provided in Section \ref{sec:meths}. \code{aneris} was able to satisfactorily
-harmonized over 96\% of the 1940 individual trajectories that were analyzed in
-Section \ref{sec:results}. Of the remaining trajectories, harmonization method
-overrides were applied, and example situations in overrides were needed were
-discussed.
+emissions harmonization for \glspl{iam}, \code{aneris}. An in-depth explanation
+of the processes and methods for determining the use of harmonization methods
+was provided in Section \ref{sec:meths}. \code{aneris} was able to
+satisfactorily harmonized over 96\% of the 1940 individual trajectories that
+were analyzed in Section \ref{sec:results}. Of the remaining trajectories,
+harmonization method overrides were applied, and a discussion was provided
+detailing why overrides were deemed necessary.
 

### 2018-01-31 paper.tex
addressed all text reviewer comments

@@
 and economic futures. Recently, the \gls{iam} community has developed scenarios
-based on the \glspl{ssp} \cite{van_vuuren_energy_2017, fricko_marker_2017,
-  fujimori_ssp3:_2017, calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which
+based on the \glspl{ssp} \cite{fujimori_ssp3:_2017, van_vuuren_energy_2017,
+  calvin_ssp4:_2017, fricko_marker_2017, kriegler_fossil-fueled_2017} which
 quantify a variety of potential global futures. The \glspl{ssp} are designed to
@@
 
-While \glspl{iam} are implemented in myriad ways, including simulation and
-optimization, the core inputs and outputs are similar across different
-models. Modeling teams incorporate data on energy systems, land use, economics,
-demographics and emissions sources and concentrations, among other data, in
-order to provide consistent existing trajectories of modeled variables. The
-models then provide estimates of future trajectories of these variables under
-various socio-economic and technological assumptions as well as proposed policy
-constraints, e.g., targets for future \gls{ghg} emissions.
+While \glspl{iam} are implemented in myriad ways\footnote{IAM models are
+  numerous and have a long history in the scientific literature. Various IAMs
+  have collaborated to produce
+  \href{http://themasites.pbl.nl/models/advance/index.php/ADVANCE_wiki}{community
+    IAM documentation} which readers can access for a full treatment of model
+  implementation and features.}, including simulation and optimization, the core
+inputs and outputs are similar across different models. Modeling teams
+incorporate data on energy systems, land use, economics, demographics and
+emissions sources and concentrations, among other data, in order to provide
+consistent existing trajectories of modeled variables. The models then provide
+estimates of future trajectories of these variables under various socio-economic
+and technological assumptions as well as proposed policy constraints, e.g.,
+targets for future \gls{ghg} emissions.
 
@@
 
+\subsection{The Conceptual Basis for Choosing Harmonization Methods}\label{sec:concept}
+
+The goals of any harmonization exercise is threefold: aligning model results in
+the harmonization year to a common historical data source, faithfully
+representing the original \gls{iam}’s internal consistency between the driver
+of emissions (e.g. energy use) and emissions, and maintaining critical
+parameters from the original scenario design. All harmonization methods achieve
+the first goal by design. If the difference between the model base year and
+historical values are small, considering the second and third goals leads to a
+\code{reduce\_ratio} method that converges prior to the final model year as the
+natural best choice. It preserves the ratio of \gls{iam} output and emissions
+inventory in the base year while also matching the original model output at some
+point in the modeled time period. It furthermore maintains the consistency of
+the model's usage of energy technology, volume of agricultural activities, and
+abatement options with harmonized emissions trajectory.
+
+However, other concerns may lead to a better-informed choice than using a
+blanket \code{reduce\_ratio} method for all emissions trajectories. For example,
+emissions from \gls{luluc} are known to have high year-to-year variation, and
+therefore historical data may change drastically depending on the base year
+considered. In such a situation, a \code{reduce\_ratio} that converges at a year
+past the modeled time period is a better choice in order to smooth out
+discrepancies between the historical data used to develop model and the new data
+source being used for harmonization.
+
+Separately, if there are large discrepancies between the model results in the
+base year and the historic data used for harmonization, convergence methods can
+result in harmonized trajectories that do not faithfully represent the
+underlying drivers of emissions. Furthermore, if models report negative
+emissions, as is possible in scenarios designed to depict the deployment of
+climate mitigation policies with large \cotwo reductions and eventual storage,
+then end-of-century emissions characteristics can be of chief import in order to
+faithfully match the design parameters of the original scenarios, such as global
+mean temperature and other climate metrics.
+
+Accordingly, we have developed a \textit{decision tree} approach to
+harmonization method choice, discussed in Section \ref{sec:tree}, in order to
+balance each of these concerns and use cases in a robust and automated manner.
+
 \subsection{Default Method Decision Tree}\label{sec:tree}
@@
       present, convergence years of default methods are provided below the
-      method name in parentheses.  Methods labeled in \textit{green} are likely
-      to closely match unharmonized results, methods in \textit{yellow} will
-      likely somewhat match unharmonized results, and methods in \textit{red}
-      can be expected to have a large relative difference between harmonized and
+      method name in parentheses.  Convergence years are chosen in order to
+      balance the three harmonization goals discussed in Section
+      \ref{sec:concept}. Methods labeled in \textit{green} are likely to closely
+      match unharmonized results, methods in \textit{yellow} will likely
+      somewhat match unharmonized results, and methods in \textit{red} can be
+      expected to have a large relative difference between harmonized and
       unharmonized results.}
@@
 select based on the effect of the characteristic on the potential harmonized
-trajectory. For example, it is possible for models to report zero values in the
-harmonization year in situations in which technologies are introduced in future
-time periods in regions or for sectors which produce an emissions species that
-is absent in the initial modeling period. In such cases, an offset method is
-required as a ratio method would mask future emissions and erroneously harmonize
-the trajectory.
+trajectory. For example, it is rare but possible for models to report zero
+values in the harmonization year in situations in which technologies are
+introduced in future time periods in regions or for sectors which produce an
+emissions species that is absent in the initial modeling period. In such cases,
+an offset method is required as a ratio method would mask future emissions and
+erroneously harmonize the trajectory.
 
 In most cases, however, models do report values in the harmonization
-year. Figure \ref{fig:cases} displays a number of illustrative example
-trajectories which highlight the possible issues resulting from harmonizing
-model results in different contexts. Panels \code{a} and \code{b} highlight
-examples where model results peak mid-century, behavior that is seen in a number
-of scenarios with general emissions mitigation effects, such as pollution
-controls applied by developing nations on transport and industry sectors. Panel
-\code{a} highlights a case where model base-year values and history are relative
-close whereas Panel \code{b} shows a situation where model values and history
-are relatively far apart. Panels \code{c} and \code{d} show similar model
-trajectories that peak mid-century but also have negative emissions. Models can
-report negative emissions in contexts in which a gas species which can be
-extracted from the environment and stored, as is the case for \cotwo in future
-scenarios with climate mitigation enacted via the deployment of carbon capture
-and sequestration technologies or direct carbon capture technologies. Again, the
+year. Figure \ref{fig:cases} displays a number of example trajectories which
+highlight the possible issues resulting from harmonizing model results in
+different contexts. These trajectories do not correspond to specific model
+results; rather they serve as illustrative examples of the kinds of trajectories
+observed in practice. Panels \code{a} and \code{b} highlight examples where
+model results peak mid-century, behavior that is seen in a number of scenarios
+with general emissions mitigation effects, such as pollution controls applied by
+developing nations on transport and industry sectors. Panel \code{a} highlights
+a case where model base-year values and history are relative close whereas Panel
+\code{b} shows a situation where model values and history are relatively far
+apart. Panels \code{c} and \code{d} show similar model trajectories that peak
+mid-century but also have negative emissions. Models can report negative
+emissions in contexts in which a gas species which can be extracted from the
+environment and stored, as is the case for \cotwo in future scenarios with
+climate mitigation enacted via the deployment of carbon capture and
+sequestration technologies or direct carbon capture technologies. Again, the
 relative difference with historical values differ between the panels to explore
@@
       \label{fig:cases}
-      The effect of different harmonization routines on model trajectories under
-      ``normal'' circumstances (Panel \code{a}), when there is a large
-      difference between historical and model values in the harmonization year
-      (Panels \code{b} and \code{d}), and when model trajectories result in
-      negative emissions by the end of the modeling time horizon (Panels
-      \code{c} and \code{d}). Identical model trajectories are used in each row
-      (Panels \code{a}, \code{b}; \code{c}, \code{d}). In each column,
-      historical values are increased in the base year by an order of magnitude
-      (from 10 to 100). In each Panel, a subset of the potential routines
-      provide a better harmonization quality than others as described in the
-      text.
-    }
+      We present here a number of illustrative examples of the effect of
+      different harmonization routines on model trajectories under ``normal''
+      circumstances (Panel \code{a}), when there is a large difference between
+      historical and model values in the harmonization year (Panels \code{b} and
+      \code{d}), and when model trajectories result in negative emissions by the
+      end of the modeling time horizon (Panels \code{c} and \code{d}). Identical
+      model trajectories are used in each row (Panels \code{a}, \code{b};
+      \code{c}, \code{d}). In each column, historical values are increased in
+      the base year by an order of magnitude (from 10 to 100). In each Panel, a
+      subset of the potential routines provide a better harmonization quality
+      than others as described in the text.  }
   \end{center}
@@
 The value of $c_v$ is then tested against a threshold, $\tau_{c_v}$. To
-determine this threshold, an analysis of the recent \gls{ceds} and \gls{luc} historical data
-has been performed. Figure \ref{fig:cov} shows the distribution of \gls{luc} $c_v$s
-and non-\gls{luc} $c_v$s as determined for historical data aggregated to the model
-regions of 5 different \glspl{iam} involved in the \gls{ssp} process: AIM-CGE, IMAGE, GCAM4,
-MESSAGE-GLOBIOM, and REMIND-MAGPIE \TODO{add refs}, each of which have varying
+determine this threshold, an analysis of the recent \gls{ceds} and \gls{luc}
+historical data has been performed. Figure \ref{fig:cov} shows the distribution
+of \gls{luc} $c_v$s and non-\gls{luc} $c_v$s as determined for historical data
+aggregated to the model regions of 5 different \glspl{iam} involved in the
+\gls{ssp} process: AIM-CGE \cite{fujimori_ssp3:_2017},
+IMAGE\cite{van_vuuren_energy_2017}, GCAM4\cite{calvin_ssp4:_2017},
+MESSAGE-GLOBIOM\cite{fricko_marker_2017}, and
+REMIND-MAGPIE\cite{kriegler_fossil-fueled_2017}, each of which have varying
 definitions of native model regions comprising different collections of
@@
 these observations as it optimally divides the two distributions. Importantly,
-tails of the \gls{luc} and non-\gls{luc} overlap, thus there are both false positives (~7\%
-of non-\gls{luc} trajectories) and false negatives (~10\% of \gls{luc}
-trajectories). However, as any regional definition is model dependent and thus
-any regional aggregation is possible an automated detection mechanism is
-necessary.
+tails of the \gls{luc} and non-\gls{luc} overlap, thus there are both false
+positives (~7\% of non-\gls{luc} trajectories) and false negatives (~10\% of
+\gls{luc} trajectories). However, as any regional definition is model dependent
+and thus any regional aggregation is possible an automated detection mechanism
+is necessary.
 

### 2018-01-31 paper.tex
spell check

@@
   ($t_i$) decays to the unharmonized model result (i.e., the convergence factor
-  is 0) in the convergence year ($t_f$). In cases where the convergency factor
+  is 0) in the convergence year ($t_f$). In cases where the convergence factor
   is applied over the entire time horizon, the convergence year is taken to be
@@
 standard deviation, $\sigma$, and the mean, $\mu$, as shown in Equation
-\ref{eqs:cov}. For a single realiziation of $c_v$, the first derivative
+\ref{eqs:cov}. For a single realization of $c_v$, the first derivative
 information of the entire historical time period is utilized.
@@
 \texttt{region}, \texttt{sector}, \texttt{gas}, and \texttt{units} columns
-without extraneous metadata columns for the model and scenario. Once configured
+without extraneous meta-data columns for the model and scenario. Once configured
 with appropriate input data (model and history) as well as potential method
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 

### 2018-02-04 paper.tex
updated figures and results discussion

@@
     \hline
-CPA             & 0.35        & 2.26          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
-PAS             & 0.14        & 1.24          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
-SAS             & 0.56        & 0.58          & 1 (no), 2 (no), 3 (no), 4 (no)                          & constant\_ratio        \\ \hline
+\gls{cpa}             & 0.35        & 2.26          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
+\gls{pas}             & 0.14        & 1.24          & 1 (no), 2 (no), 3 (yes)                                 & reduce\_ratio\_2080    \\ \hline
+\gls{sas}             & 0.56        & 0.58          & 1 (no), 2 (no), 3 (no), 4 (no)                          & constant\_ratio        \\ \hline
 \end{tabular}
@@
 the harmonization year for carbon monoxide (CO) emissions in the industrial
-sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region, which
-comprises most of the emissions of the Asian subcontinent. The default method
-chosen (\code{constant_ratio}) maintains model trends for the region; however,
-overall model results are distorted. By applying a \code{constant_offset}
-override, the regional trend and magnitude is maintained. With the new
-harmonization method for the SAS region, the global trajectory for industrial CO
-also is representative of the trends seen in the unharmonized trajectory and the
-relative importance of the underlying regional trajectories is maintained.
+sector specifically for the \gls{sas} and \gls{afr} MESSAGE-GLOBIOM
+regions. Furthermore, emissions in both regions see relatively large expansion
+or contraction, depending on the scenario; therefore, both regions are
+candidates for choosing harmonization method overrides. The default method
+chosen in both cases (\code{constant_ratio}) maintains model trends for the
+region; however, overall model results are distorted. By applying a
+\code{constant_offset} overrides, the regional trends and magnitudes are
+maintained. Through the use of harmonization overrides also ameliorates issues
+seen in the harmonized global trajectory of industrial CO (Figure \ref{fig:co},
+Panel \textbf{c}). Not only does the global trajectory with overrides more
+accurately match the original unharmonized model behavior and magnitude of
+emissions, but the relative importance of the underlying regional trajectories
+is also maintained.
 
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{results_CO_Industrial_Sector.pdf}
+    \includegraphics[height=0.8\textheight]{sep_regions_results_CO_Industrial_Sector.pdf}
     \caption[]{
       \label{fig:co}
-      CO Industrial Sector harmonized and unharmonized emissions are presented
-      for SSP2 and SSP2-45 scenarios. Scenarios as denoted identically to Figure
-      \ref{fig:nox}. Panels \textbf{a} and \textbf{b} show harmonized and
-      overridden-harmonized (respectively) regional trajectories for the 3
-      MESSAGE-GLOBIOM regions that comprise the R5ASIA region: Centrally Planned
-      Asia (CPA), Other Pacific Asia (PAS), and South Asia (SAS). Notably, the
-      SAS regional trajectory displays a distorted trajectory due to the
-      harmonization-year difference between history and model results in both
-      scenarios. The distortion is large enough to affect global results, as
-      shown in Panels \textbf{c} and \textbf{d}.  
-}
+      CO Industrial Sector harmonized (without markers) and unharmonized (with
+      markers) emissions are presented for SSP2 (purple lines) and SSP2-45 (red
+      lines) scenarios. Panel \textbf{a} shows harmonized and
+      overridden-harmonized (respectively) trajectories for the \gls{sas} region
+      while panels \textbf{b} and \textbf{c} show similar trajectories for the
+      \gls{afr} region and (total) global emissions, respectively. Notably, the
+      \gls{sas} and \gls{afr} regional trajectory displays a distorted
+      trajectory due to the harmonization-year difference between history and
+      model results in both scenarios. The distortion is large enough to affect
+      global results, as shown in Panel \textbf{c}.  }
   \end{center}
@@
 trajectories. Figure \ref{fig:nh3} shows an example case of emissions
-trajectories for ammonia (\nht) from the agriculture sector in Asia. Again, the
-SAS region shows a large discrepancy in the harmonization year ($>$150\% in this
-case). The resulting trajectory harmonized with the default method
+trajectories for ammonia (\nht) from the agriculture sector. Again, the
+\gls{sas} region shows a large discrepancy in the harmonization year ($>$150\%
+in this case). The resulting trajectory harmonized with the default method
 (\code{constant_ratio}) provides a large increase after 2080 in the SSP2
 reference scenario. Notably, the SSP2-45 scenario is not affected to the same
-degree. While this distortion changes the magnitude of the SAS trajectory, it
-largely affects the post-2080 shape of the global trajectory (see Figure
-\ref{fig:nh3}, panel \textbf{c}). By using a \code{constant_offset} method as an
-override, this distortion is addressed and more accurately reflects unharmonized
-results in the SAS region, the relative importance between regions, and global
-results for agricultural ammonia emissions, each of which contributes to a
-better harmonization quality for the harmonized SAS trajectory.
+degree. While this distortion changes the magnitude of the \gls{sas} trajectory,
+it largely affects the post-2080 shape of the global trajectory (Figure
+\ref{fig:nh3}, Panel \textbf{b}). For example, in the original model result,
+\gls{SAS} \nht agricultural emissions contribute $\sim$30\% of total global
+emissions, whereas in the unharmonized case with default methods, \gls{SAS}
+comprises $\sim$50\% of global emissions. By using a \code{constant_offset}
+method as an override, this distortion is addressed and more accurately reflects
+unharmonized results in the \gls{sas} region, the relative importance between
+regions, and global results for agricultural ammonia emissions, each of which
+contributes to a better harmonization quality for the harmonized \gls{sas}
+trajectory. The harmonized emissions using overrides result in a contribution of
+$\sim$32\% of total global emissions by the \gls{SAS} region which aligns
+closely with the unharmonized model results.
 
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{results_NH3_Agriculture.pdf}
+    \includegraphics[width=\textwidth]{sep_regions_results_NH3_Agriculture.pdf}
     \caption[]{
@@
       \nhtt agricultural harmonized and unharmonized emissions are presented for
-      SSP2 and SSP2-45 scenarios. Scenarios and panel layouts are identical to
-      Figure \ref{fig:co}. In this case, the SAS trajectory again shows not only
-      a magnitude distortion, but also a shape distortion at the tail of the
-      trajectory. Additionally, global trajectories are greatly affected by the
-      harmonization method choice (there is $\sim$20\% relative difference
-      between trajectories in the reference scenario in 2100). Override methods
-      have been applied to correct the distortion.  }
+      SSP2 and SSP2-45 scenarios. Panel \textbf{a} shows harmonized and
+      overridden-harmonized (respectively) trajectories for the \gls{sas}
+      region, and panel \textbf{b} shows the global total trajectory. In this
+      case, the \gls{sas} trajectory again shows not only a magnitude distortion, but
+      also a shape distortion at the tail of the trajectory. Additionally,
+      global trajectories are greatly affected by the harmonization method
+      choice (there is $\sim$20\% relative difference between trajectories in
+      the reference scenario in 2100). Override methods have been applied to
+      correct the distortion.  }
   \end{center}

### 2018-02-04 paper.tex
fix gls entries

@@
 \ref{fig:nh3}, Panel \textbf{b}). For example, in the original model result,
-\gls{SAS} \nht agricultural emissions contribute $\sim$30\% of total global
-emissions, whereas in the unharmonized case with default methods, \gls{SAS}
+\gls{sas} \nht agricultural emissions contribute $\sim$30\% of total global
+emissions, whereas in the unharmonized case with default methods, \gls{sas}
 comprises $\sim$50\% of global emissions. By using a \code{constant_offset}
@@
 trajectory. The harmonized emissions using overrides result in a contribution of
-$\sim$32\% of total global emissions by the \gls{SAS} region which aligns
+$\sim$32\% of total global emissions by the \gls{sas} region which aligns
 closely with the unharmonized model results.

### 2018-02-05 paper.tex
final self revisions

@@
 work describes the novel methodology for determining such harmonization methods
-and an open-source Python software library implementing the methodology. Results
-are shown for two example scenarios (with and without climate policy cases)
-using the IAM MESSAGE-GLOBIOM that satisfactorily harmonize over 96\% of the
-total emissions trajectories while having a negligible effect on key long-term
-climate indicators. This new capability enhances the comparability across
-different models, increases transparency and robustness of results, and allows
-other teams to easily participate in intercomparison exercises by using the
-same, openly available harmonization mechanism.
+and an open-source Python software library implementing the methodology. A case
+study is presented for two example scenarios (with and without climate policy
+cases) using the IAM MESSAGE-GLOBIOM that satisfactorily harmonize over 96\% of
+the total emissions trajectories while having a negligible effect on key
+long-term climate indicators. This new capability enhances the comparability
+across different models, increases transparency and robustness of results, and
+allows other teams to easily participate in intercomparison exercises by using
+the same, openly available harmonization mechanism.
 \end{abstract}
@@
 library (approximately 2000 lines of code). The \texttt{aneris} software was
-developed by the lead author, Matthew J. Gidden, Ph.D., whose contact
-information is shown on the title page of this manuscript. Documentation for the
-\texttt{aneris} Python package, including software requirements, is available
+developed by the lead author whose contact information is shown on the title
+page of this manuscript. Documentation for the \texttt{aneris} Python package,
+including software requirements, is available
 \href{http://mattgidden.com/aneris/}{online}.
@@
 harmonization software \code{aneris} \cite{matthew_gidden_2017_802832}, written
-in the Python programming language (detailed documentation available
+in the Python programming language (detailed documentation is available
 \href{http://mattgidden.com/aneris/}{online}). Section \ref{sec:meths} provides
 a detailed description of the underlying mathematical components of
-\code{aneris} as well as the procedural workflow. The results of applying the
+\code{aneris} as well as the procedural workflow. A case study of applying the
 automated harmonization mechanism on two example \gls{iam} scenarios, one with
@@
 
-\subsection{Harmonization Methods}
+\subsection{The Conceptual Basis for Choosing Harmonization Methods}\label{sec:concept}
+
+The goals of any harmonization exercise is threefold: aligning model results in
+the harmonization year to a common historical data source, faithfully
+representing the original \gls{iam}’s internal consistency between the driver of
+emissions (e.g. energy use) and emissions, and maintaining critical parameters
+from the original scenario design. Any harmonization method achieves the first
+goal by design. If the difference between the model base year and historical
+values are small, considering the second and third goals leads to a method
+choice that matches modeled drivers (e.g., a \code{ratio} method discussed in
+Section \ref{sec:harm-methods}) and converges prior to the final model year. It
+preserves the relationship between \gls{iam} output and emissions inventory in
+the base year while also matching the original model output at some point in the
+modeled time period. It furthermore maintains the consistency of the model's
+usage of energy technology, volume of agricultural activities, and abatement
+options with harmonized emissions trajectory.
+
+However, other concerns may lead to a better-informed choice than using a
+blanket method for all emissions trajectories. For example, emissions from
+\gls{luluc} are known to have high year-to-year variation, and therefore
+historical data may change drastically depending on the base year considered. In
+such a situation, a method that converges at a year past the modeled time period
+is a better choice in order to smooth out discrepancies between the historical
+data used to develop model and the new data source being used for harmonization.
+
+Separately, if there are large discrepancies between the model results in the
+base year and the historic data used for harmonization, convergence methods can
+result in harmonized trajectories that do not faithfully represent the
+underlying drivers of emissions. Furthermore, if models report negative
+emissions, as is possible in scenarios designed to depict the deployment of
+climate mitigation policies with large \cotwo reductions and eventual storage,
+then end-of-century emissions characteristics can be of chief import in order to
+faithfully match the design parameters of the original scenarios, such as global
+mean temperature and other climate metrics.
+
+Accordingly, we have developed a \textit{decision tree} approach to
+harmonization method choice, discussed in Section \ref{sec:tree}, in order to
+balance each of these concerns and use cases in a robust and automated manner.
+
+\subsection{Harmonization Methods}\label{sec:harm-methods}
 
@@
 \end{table}
-
-\subsection{The Conceptual Basis for Choosing Harmonization Methods}\label{sec:concept}
-
-The goals of any harmonization exercise is threefold: aligning model results in
-the harmonization year to a common historical data source, faithfully
-representing the original \gls{iam}’s internal consistency between the driver
-of emissions (e.g. energy use) and emissions, and maintaining critical
-parameters from the original scenario design. All harmonization methods achieve
-the first goal by design. If the difference between the model base year and
-historical values are small, considering the second and third goals leads to a
-\code{reduce\_ratio} method that converges prior to the final model year as the
-natural best choice. It preserves the ratio of \gls{iam} output and emissions
-inventory in the base year while also matching the original model output at some
-point in the modeled time period. It furthermore maintains the consistency of
-the model's usage of energy technology, volume of agricultural activities, and
-abatement options with harmonized emissions trajectory.
-
-However, other concerns may lead to a better-informed choice than using a
-blanket \code{reduce\_ratio} method for all emissions trajectories. For example,
-emissions from \gls{luluc} are known to have high year-to-year variation, and
-therefore historical data may change drastically depending on the base year
-considered. In such a situation, a \code{reduce\_ratio} that converges at a year
-past the modeled time period is a better choice in order to smooth out
-discrepancies between the historical data used to develop model and the new data
-source being used for harmonization.
-
-Separately, if there are large discrepancies between the model results in the
-base year and the historic data used for harmonization, convergence methods can
-result in harmonized trajectories that do not faithfully represent the
-underlying drivers of emissions. Furthermore, if models report negative
-emissions, as is possible in scenarios designed to depict the deployment of
-climate mitigation policies with large \cotwo reductions and eventual storage,
-then end-of-century emissions characteristics can be of chief import in order to
-faithfully match the design parameters of the original scenarios, such as global
-mean temperature and other climate metrics.
-
-Accordingly, we have developed a \textit{decision tree} approach to
-harmonization method choice, discussed in Section \ref{sec:tree}, in order to
-balance each of these concerns and use cases in a robust and automated manner.
 
@@
 a realization of $c_v$ for a single combination of native model region, sector,
-and emissions species \footnote{A full listing of all sectors and species is
+and emissions species\footnote{A full listing of all sectors and species is
   presented in the case-study discussion in Section \ref{sec:results}, Table
@@
 
-We herein present \code{aneris}' Python implementation, depicted in Figure
-\ref{fig:software}, and conceptual, depicted in Figure \ref{fig:workflow}. The
-library is composed of a number of utilities as well as three primary
-components: the \texttt{HarmonizationDriver}, \texttt{Harmonizer}, and data
-processing routines (shown in red, green, and blue, respectively in Figures
-\ref{fig:software} and \ref{fig:workflow}).
+We herein present \code{aneris}' Python implementation and conceptual
+design.. The library is composed of a number of utilities as well as three
+primary components: the \texttt{HarmonizationDriver}, \texttt{Harmonizer}, and
+data processing routines shown in red, green, and blue, respectively in Figures
+\ref{fig:software} and \ref{fig:workflow}.
 
@@
 
-The \texttt{Harmonizer} is a Python class whose concern is to harmonize model
-value trajectories given historical data and possible user method
-\textit{overrides}, i.e., non-default methods (described further below). It is
-used by the \texttt{HarmonizationDriver}; however it is also available to the
+The \texttt{Harmonizer} is a class whose concern is to harmonize model value
+trajectories given historical data and possible user method \textit{overrides},
+i.e., non-default methods (described further in Section \ref{sec:overrides}). It
+is used by the \texttt{HarmonizationDriver}; however it is also available to the
 user as a first-class object. The \texttt{Harmonizer} requires that input data
@@
 The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins
-by cleaning input data, which adds (0-valued) model trajectories that exist in
-the historical data set but are not provided by the model input and detects any
-issues that would cause the harmonization process to fail. The methods used to
-harmonize the data are then determined and the harmonization process is
-executed. Upon completion of the harmonization process, aggregation of common
-analysis regions is performed. A common regional aggregation used in the
-\gls{iam} community was defined in the \glspl{rcp}
-\cite{vuuren_representative_2011}, shown in Figure \ref{fig:regions}. Finally,
-any exogenous trajectories the user provides are added. Exogenous trajectories
-are normally provided for unmodeled gases with well-accepted scenario
-trajectories, e.g., chlorofluorocarbons provided by the \gls{wmo}
-\cite{wmo2014}. Upon completion, the harmonized trajectories and meta data
-regarding the harmonization process are returned. A description of all returned
-meta data is provided in Table \ref{tab:metadata}.
+by cleaning input data. Cleaning operations include adding (0-valued) model
+trajectories that exist in the historical data set but are not provided by the
+model input and detecting any issues that would cause the harmonization process
+to fail. The methods used to harmonize the data are then determined and the
+harmonization process is executed. Upon completion of the harmonization process,
+spatial aggregation to common analysis regions is performed. For example, the
+5-region aggregation developed in the \glspl{rcp}
+\cite{vuuren_representative_2011} process is commonly used in the \gls{iam}
+community and is shown in Figure \ref{fig:regions}. Finally, any exogenous
+trajectories the user provides are added. Exogenous trajectories are normally
+provided for unmodeled gases with well-accepted scenario trajectories, e.g.,
+chlorofluorocarbons provided by the \gls{wmo} \cite{wmo2014}. Upon completion,
+the harmonized trajectories and meta data regarding the harmonization process
+are returned. A description of all returned meta data is provided in Table
+\ref{tab:metadata}.
 
@@
       \label{fig:regions}
-      The 5 regions used in the \glspl{rcp} with their 11-region constituents: Asia
-      (\gls{cpa}, \gls{sas}, \gls{pas}) [yellows], \gls{lam}
-      [magenta], the \gls{mea} and \gls{afr} [greens], the OECD (\gls{nam},
-      \gls{weu}, and \gls{pao}) [blues], and the Reforming Economies
-      (\gls{eeu} and \gls{fsu}) [reds].  }
+      The 5 regions used in the \glspl{rcp} with their MESSAGE-GLOBIOM 11-region
+      constituents: Asia (\gls{cpa}, \gls{sas}, \gls{pas}) [yellows], \gls{lam}
+      [magenta], Africa and the Middle East (\gls{mea} and \gls{afr}) [greens],
+      the OECD (\gls{nam}, \gls{weu}, and \gls{pao}) [blues], and the Reforming
+      Economies (\gls{eeu} and \gls{fsu}) [reds].  }
   \end{center}
@@
 \end{table}
+
+\subsection{User-Defined Override Methods}\label{sec:overrides}
 
@@
 and $200$\%, respectively, are provided based on experiences of the authors'
-using \code{aneris} to date.
+use of \code{aneris} to date.
 
@@
 sensitivity of key indicators, such as end-of-century radiative forcing (which
-is used to estimate mean global temperature increase), to the timing and
+is used to estimate mean global temperature response), to the timing and
 magnitude of net-zero and total negative \cotwo emissions. Therefore, these two
-scenarios represent two relative extreme cases in the use of a harmonization
-approach and thus provide a case study as to its general applicability. Figure
-\ref{fig:kyoto} shows the different trends of Kyoto Gases in each scenarios.
+scenarios represent two bounding cases in the use of a harmonization approach
+and thus provide a case study as to its general applicability. Figure
+\ref{fig:kyoto} shows the different trends of Kyoto Gases, a measure of
+aggregate \gls{ghg} emissions, in each scenarios.
 
@@
 MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
-mapped directly to the 5-region definition used in the \glspl{rcp}. Historical data is
-taken from previously described \gls{luc} and anthropogenic sources, which comprise 10
-separate pollutant and \gls{ghg} species and 12 sectors shown in Table \ref{tab:sp}. A
-total of 970 distinct trajectories were harmonized for each scenario, and
-therefore 1940 trajectories were harmonized in total. \noxx generated from the
-Energy sector provides an example of an emissions species and sector in which
-all regions were satisfactorily harmonized with the default methods. Figure
-\ref{fig:nox} shows the results of harmonization in Asia, and Table
-\ref{tab:nox} describes the parameters that underlie the choice of method for
-each harmonized trajectory.
+mapped directly to the 5-region definition used in the \glspl{rcp}. Historical
+data is taken from previously described \gls{luc} and anthropogenic sources,
+which comprise 10 separate pollutant and \gls{ghg} species and 12 sectors shown
+in Table \ref{tab:sp}. A total of 970 distinct trajectories were harmonized for
+each scenario, and therefore 1940 trajectories were harmonized in total. \noxx
+generated from the Energy sector provides an example of an emissions species and
+sector in which all regions were satisfactorily harmonized with the default
+methods. Figure \ref{fig:nox} shows the results of harmonization in Asia, and
+Table \ref{tab:nox} describes the parameters that underlie the choice of method
+for each harmonized trajectory. Importantly, the default methods provide
+regional trajectories for both the reference and mitigation scenarios closely
+matching the unharmonized model behavior and also result in global aggregates
+that agree with the unharmonized trajectory over the modeled time period.
 
@@
 the harmonization year for carbon monoxide (CO) emissions in the industrial
-sector specifically for the \gls{sas} and \gls{afr} MESSAGE-GLOBIOM
-regions. Furthermore, emissions in both regions see relatively large expansion
-or contraction, depending on the scenario; therefore, both regions are
-candidates for choosing harmonization method overrides. The default method
-chosen in both cases (\code{constant_ratio}) maintains model trends for the
-region; however, overall model results are distorted. By applying a
-\code{constant_offset} overrides, the regional trends and magnitudes are
-maintained. Through the use of harmonization overrides also ameliorates issues
-seen in the harmonized global trajectory of industrial CO (Figure \ref{fig:co},
-Panel \textbf{c}). Not only does the global trajectory with overrides more
-accurately match the original unharmonized model behavior and magnitude of
-emissions, but the relative importance of the underlying regional trajectories
-is also maintained.
+sector specifically for the \gls{sas} and \gls{afr} regions. Furthermore,
+emissions in both regions see relatively large expansion or contraction,
+depending on the scenario; therefore, both regions are candidates for choosing
+harmonization method overrides. The default method chosen in both cases
+(\code{constant_ratio}) maintains model trends for the region; however, overall
+model results are distorted as can be seen in both the regional and global
+panels in Figure \ref{fig:co}. By applying \code{constant_offset} overrides, the
+regional trends and magnitudes are maintained. The use of harmonization
+overrides also ameliorates issues seen in the harmonized global trajectory of
+industrial CO (Figure \ref{fig:co}, Panel \textbf{c}). Not only does the global
+trajectory with overrides more accurately match the original unharmonized model
+behavior and magnitude of emissions, but the relative importance of the
+underlying regional trajectories is also maintained.
 
@@
 it largely affects the post-2080 shape of the global trajectory (Figure
-\ref{fig:nh3}, Panel \textbf{b}). For example, in the original model result,
-\gls{sas} \nht agricultural emissions contribute $\sim$30\% of total global
+\ref{fig:nh3}, Panel \textbf{b}) as well as the relative regional contributions
+to the global aggregate trajectory. For example, in the original model result,
+\gls{sas} \nht~ agricultural emissions contribute $\sim$30\% of total global
 emissions, whereas in the unharmonized case with default methods, \gls{sas}
-comprises $\sim$50\% of global emissions. By using a \code{constant_offset}
-method as an override, this distortion is addressed and more accurately reflects
-unharmonized results in the \gls{sas} region, the relative importance between
-regions, and global results for agricultural ammonia emissions, each of which
-contributes to a better harmonization quality for the harmonized \gls{sas}
-trajectory. The harmonized emissions using overrides result in a contribution of
-$\sim$32\% of total global emissions by the \gls{sas} region which aligns
-closely with the unharmonized model results.
+comprises $\sim$50\% of global emissions by 2100. By using a
+\code{constant_offset} method as an override, this distortion is addressed and
+more accurately reflects unharmonized results in the \gls{sas} region, the
+relative importance between regions, and global results for agricultural ammonia
+emissions, each of which contributes to a better harmonization quality for the
+harmonized \gls{sas} trajectory. The harmonized emissions using overrides result
+in a contribution of $\sim$32\% of total global emissions by the \gls{sas}
+region which aligns closely with the unharmonized model results.
 
@@
 This work provides a new direction and framework which the \gls{iam} and climate
-communities can build upon in order to reduce the necessity of consistent expert
-opinion and increase transparency and reproducibility of harmonization
-exercises. Furthermore, it provides an open-source, tested, and documented
-software library which can be used and improved upon by these communities. Both
-of these are clear steps in a positive direction for future climate and
-integrated assessment modeling exercises.
+science communities can build upon in order to reduce the necessity of
+consistent expert opinion and increase transparency and reproducibility of
+harmonization exercises. Furthermore, it provides an open-source, tested, and
+documented software library which can be used and improved upon by these
+communities. Both of these are clear steps in a positive direction for future
+climate and integrated assessment modeling exercises.
 

### 2018-02-05 paper.tex
luc to luluc

@@
 \newacronym{esm}{ESM}{Earth System Model}
-\newacronym{luc}{LUC}{Land-use Change}
-\newacronym{luluc}{LULUC}{Land-use and Land-use Change}
+\newacronym{luc}{LULUC}{Land-use and Land-use Change}
 \newacronym{weu}{WEU}{Western Europe}
@@
 historical emissions data set for both anthropogenic emissions (i.e., the
-\gls{ceds} \cite{hoesly_historical_2017} and open-burning \gls{luluc} emissions
+\gls{ceds} \cite{hoesly_historical_2017} and open-burning \gls{luc} emissions
 \cite{van_marle_historic_2017}) which, in conjunction with the \gls{ssp}
@@
 blanket method for all emissions trajectories. For example, emissions from
-\gls{luluc} are known to have high year-to-year variation, and therefore
+\gls{luc} are known to have high year-to-year variation, and therefore
 historical data may change drastically depending on the base year considered. In

### 2018-02-06 paper.tex
rm kyoto gas title and add it to the caption

@@
       \label{fig:kyoto}
-      Unharmonized Kyoto gas emissions for SSP2-Ref, a scenario with generally
-      increasing global emissions trends, and SSP2-45, a scenario with generally
-      decreasing global emissions trends.  }
+      Unharmonized global Kyoto gas emissions for SSP2-Ref, a scenario with
+      generally increasing global emissions trends, and SSP2-45, a scenario with
+      generally decreasing global emissions trends.  }
   \end{center}

### 2018-02-06 correspondance/revision1_letter/letter.tex
adding revision letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%       PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}
 \begin{document}

%----------------------------------------------------------------------------------------
%       ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{
    Professor S. Reis\\
    Editor, Environmental Modelling \& Software\\
    \hfill\break
    Professor D.P. Ames\\
    Editor-in-Chief, Environmental Modelling \& Software\\
  }

%----------------------------------------------------------------------------------------
%       YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

\address{Matthew Gidden\\
Laimgrubengasse 17/7\\
Vienna, Austria 1060}

%----------------------------------------------------------------------------------------
%       LETTER CONTENT SECTION
%----------------------------------------------------------------------------------------

\opening{Professors Reis and Ames,}

Please find enclosed our revised manuscript and response to reviewers regarding
the submitted manuscript entitled: ``A Methodology and Implementation of
Automated Emissions Harmonization for Use in Integrated Assessment Models''
(Ref: \texttt{ENVSOFT\_2017\_668}).

We very much appreciate the time and thoroughness taken by the three reviewers
of our manuscript, and believe that the resulting revised manuscript is
non-trivially improved over the submitted version. While we have provided direct
responses to each point made by the reviewers in the attached response document,
we summarize their primary suggestions below:

\begin{itemize}
\item We should better explain the intuition behind the harmonization method choices
\item We should greatly expand on the implementation details of the software
\item We should clarify and add a listing of all used acronyms
\item We should provide more context and sources of further reading for various terms and models discussed in the text.
\item We should consider framing the results section as a case study.
\end{itemize}

We have fully addressed and taken on board these core suggestions in the revised
manuscript. In order to better prime the discussion in the \textit{Methodology \&
  Implementation} section, we have written an introductory subsection entitled
\textit{The Conceptual Basis for Choosing Harmonization Methods} which outlines
the intuition for various methods to be used in the harmonization process. We
have greatly expanded the \textit{Python Implementation and Workflow} subsection
to include specific details of the software implementation, code snippets and
CLI listings, and a new figure (Figure 5 in the text) which provides a graphical
description of the implementation.

We have additionally made use of a global acronyms table which identifies and
lists all acronyms used in the text and have placed this table at the end of the
manuscript for readers' reference. In all instances raised by the reviewers for
further clarification on terms and models, we have added footnotes with an
additional description and links or references for further reading on the
topic. Finally, we have reframed slightly the last section, renaming it to
\textit{Case Study: Harmonizing Results from a Global IAM}. We agree with
Reviewer 3 that even in its initial form it is in fact a study of a specific use
case with \texttt{aneris} that details both the positive aspects (that the vast
majority of trajectories were harmonized satisfactorily) as well as what we did
to solve specific cases where default methods led to non-preferable harmonized
results.

After revision, we believe that the manuscript has been improved over the
initially submitted draft. The authors would like to thank the reviewers for
their constructive comments and suggestions as well as for their time taken to
review our article. We would also again like to thank you again for your
consideration of this work.

\closing{
Warm Regards,

Matthew Gidden, Ph.D.
}

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}

### 2018-02-06 correspondance/revision1_letter/letter.tex
adjust slight wording and spacing of letter

@@
     Professor D.P. Ames\\
-    Editor-in-Chief, Environmental Modelling \& Software\\
+    Editor-in-Chief, Environmental Modelling \& Software
+    \hfill\break
   }
@@
 
-After revision, we believe that the manuscript has been improved over the
-initially submitted draft. The authors would like to thank the reviewers for
-their constructive comments and suggestions as well as for their time taken to
-review our article. We would also again like to thank you again for your
-consideration of this work.
+The authors believe that the manuscript has been improved over the initially
+submitted draft. We would like to thank the reviewers for their constructive
+comments and suggestions as well as for their time taken to review our
+article. We would also again like to thank you again for your consideration of
+this work.
 
 \closing{
-Warm Regards,
-
+  \hfill\break
+  \hfill\break
+  Warm Regards,\\
+  \hfill\break
   Matthew Gidden, Ph.D.

### 2018-02-07 paper.tex
Update on Overleaf.

@@
 a number of international studies, which notably include projections of climate
-and economic futures. Recently, the \gls{iam} community has developed scenarios
+and energy futures. Recently, the \gls{iam} community has developed scenarios
 based on the \glspl{ssp} \cite{fujimori_ssp3:_2017, van_vuuren_energy_2017,
@@
 emissions, as is possible in scenarios designed to depict the deployment of
-climate mitigation policies with large \cotwo reductions and eventual storage,
+climate mitigation policies with large \cotwo reductions and storage,
 then end-of-century emissions characteristics can be of chief import in order to
@@
 
-A number of the classic methods are implemented in \code{aneris} including
+A number  methods are implemented in \code{aneris} including
 ratio-convergence shown in Equation \ref{eqs:ratio}, offset-convergence shown in
-Equation \ref{eqs:offset}, and interpolation shown in Equation
+Equation \ref{eqs:offset}, and linear interpolation shown in Equation
 \ref{eqs:interpolate}. In all equations the region, species, and sector indices
@@
       $dH$ is defined as $\left|\frac{h(t_i) - m(t_i)}{h(t_i)}\right|$, and
-      decision-making thresholds for $c_v$ and $dH$ are described below.  Where
+      decision-making thresholds for $c_v$ and $dH$ are described in the main text.  Where
       present, convergence years of default methods are provided below the
@@
 select based on the effect of the characteristic on the potential harmonized
-trajectory. For example, it is rare but possible for models to report zero
-values in the harmonization year in situations in which technologies are
-introduced in future time periods in regions or for sectors which produce an
-emissions species that is absent in the initial modeling period. In such cases,
+trajectory. For example, it is rare but possible for a sector to have emissions 
+reported in the historical data, but the model to report zero for the harmonization year, 
+with non-zero future values. In such cases,
 an offset method is required as a ratio method would mask future emissions and
-erroneously harmonize the trajectory.
+erroneously harmonize the trajectory to zero.
 
@@
 mid-century but also have negative emissions. Models can report negative
-emissions in contexts in which a gas species which can be extracted from the
-environment and stored, as is the case for \cotwo in future scenarios with
+emissions for \cotwo in future scenarios with
 climate mitigation enacted via the deployment of carbon capture and
-sequestration technologies or direct carbon capture technologies. Again, the
+storage technologies. Again, the
 relative difference with historical values differ between the panels to explore
@@
 
-Finally, consideration is taken with respect to the relative difference between
+Finally, the default harmonization decisions depend on the relative difference between
 the historic and model values in the harmonization time period. In order to
@@
 
-The \texttt{Harmonizer} is a class whose concern is to harmonize model value
+The \texttt{Harmonizer} is a class whose task is to harmonize model value
 trajectories given historical data and possible user method \textit{overrides},
@@
       The various objects and their relation to one another in the \code{aneris}
-      code base as well as a short description of their scope of concern.  
+      code base as well as a short description of their scope.  
     }
@@
 %%     \caption{ Panel \ref{fig:software} shows the various objects in the
-%%       \code{aneris} code base as well as a short description of their scope of
-%%       concern. Panel \ref{fig:workflow} displays the full harmonization process
+%%       \code{aneris} code base as well as a short description of their scope.
+%%       Panel \ref{fig:workflow} displays the full harmonization process
 %%       as executed by \code{aneris} for a single instance of a model and
@@
 The full harmonization workflow, outlined in Figure \ref{fig:workflow}, begins
-by cleaning input data. Cleaning operations include adding (0-valued) model
-trajectories that exist in the historical data set but are not provided by the
+by cleaning input data. Cleaning operations include adding model
+trajectories with 0 values where a sector/emission combination exists in the historical data set but are not provided by the
 model input and detecting any issues that would cause the harmonization process
@@
 
-In order to help identify cases where overrides may be of interest,
-harmonization \textit{diagnostics} are provided which analyze the relative
+In order to help identify cases where overrides may be needed,
+harmonization \textit{diagnostics} are provided which report the relative
 difference between harmonized and unharmonized trajectories at their mid and
-end-points. These values are configurable by the user, but defaults of $400$\%
+end-points when these values exceed specified threshold. The diagnostic reporting thresholds  are configurable by the user, but defaults of $400$\%
 and $200$\%, respectively, are provided based on experiences of the authors'
@@
 \code{aneris} harmonization procedure, we focus on the harmonization of results
-of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Two scenarios from
+of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Harmonization results for two scenarios from
 the \gls{ssp} scenario library\footnote{We refer the reader to the broad
@@
   \cite{vuuren_new_2013,oneill_scenario_2016, oneill_new_2013} for a more in
-  depth discussion of the scenario architecture and design.} are presented. The
-SSP2 \cite{Riahi2017153,Rao2017346}, or ``middle of the road'', scenario
-(referred to as SSP2-Ref) is chosen to be analyzed because MESSAGE-GLOBIOM is
-the marker scenario\footnote{The concept of ``marker'' scenarios are used to
-  illustrate distinctive scenario types within the SSP scenario framework. See
+  depth discussion of the scenario architecture and design.} are presented here. We use the
+SSP2-reference\cite{Riahi2017153,Rao2017346}, or ``middle of the road'', scenario
+(referred to as SSP2-Ref) as an example because MESSAGE-GLOBIOM is
+the marker scenario\footnote{The ``marker'' scenario concept is used to
+  designate the archetype scenario used as a reference within each SSP scenario family. See
   \cite{vuuren_new_2013} and \cite{Riahi2017153} for a more lengthy
   description.} for this \gls{ssp}. This SSP2 scenario lies between two
-\glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing is a
-  measurement of the capacity for different emissions species to affect
-  tropopausal irradiance and has a relationship to global mean equilibrium
-  temperature change. We refer the reader to \cite{IPCCWG1PhysicalStocker2013}
+\glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing in this concept is the energy inbalance at the top of the atmsophere caused by anthropogenic influences relative to a pre-industrial reference point.
+Higher radiative forcing leads to larger global changes, such as surface temperture.
+We refer the reader to \cite{IPCCWG1PhysicalStocker2013}
   for a more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
@@
 magnitude of net-zero and total negative \cotwo emissions. Therefore, these two
-scenarios represent two bounding cases in the use of a harmonization approach
+scenarios represent two contrasting cases in the use of a harmonization approach
 and thus provide a case study as to its general applicability. Figure
@@
 
-MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
-mapped directly to the 5-region definition used in the \glspl{rcp}. Historical
+MESSAGE-GLOBIOM includes a representation of 11 distinct regions, and harmonization is performed at the native regional resolution of each model. Historical
 data is taken from previously described \gls{luc} and anthropogenic sources,
@@
 in Table \ref{tab:sp}. A total of 970 distinct trajectories were harmonized for
-each scenario, and therefore 1940 trajectories were harmonized in total. \noxx
+each scenario, and therefore 1940 trajectories were harmonized in total for these two illustrative scenarios. 
+\noxx
 generated from the Energy sector provides an example of an emissions species and
@@
 for each harmonized trajectory. Importantly, the default methods provide
-regional trajectories for both the reference and mitigation scenarios closely
-matching the unharmonized model behavior and also result in global aggregates
-that agree with the unharmonized trajectory over the modeled time period.
+regional trajectories for both the reference and mitigation scenarios that
+match the general unharmonized model behavior and also result in global aggregated emissions
+that are similar to the unharmonized trajectory.
 
@@
 industrial CO (Figure \ref{fig:co}, Panel \textbf{c}). Not only does the global
-trajectory with overrides more accurately match the original unharmonized model
+trajectory with overrides more closely match the original unharmonized model
 behavior and magnitude of emissions, but the relative importance of the
@@
       markers) emissions are presented for SSP2 (purple lines) and SSP2-45 (red
-      lines) scenarios. Panel \textbf{a} shows harmonized and
-      overridden-harmonized (respectively) trajectories for the \gls{sas} region
-      while panels \textbf{b} and \textbf{c} show similar trajectories for the
+      lines) scenarios. The left column shows the deault harmonization and the right column shows harmonization with overrides. Panel \textbf{a} shows  trajectories for the \gls{sas} region
+      while panels \textbf{b} and \textbf{c} show  trajectories for the
       \gls{afr} region and (total) global emissions, respectively. Notably, the
-      \gls{sas} and \gls{afr} regional trajectory displays a distorted
-      trajectory due to the harmonization-year difference between history and
+      \gls{sas} and \gls{afr} regional trajectories are distorted
+      when the default methods are used due to the harmonization-year difference between history and
       model results in both scenarios. The distortion is large enough to affect
-      global results, as shown in Panel \textbf{c}.  }
+      global results, as shown in Panel \textbf{c}.  The use of overrides (right column) results in better consistantcy with the unharmonized model scenario at the regional and global levels. }
   \end{center}
@@
 \gls{sas} \nht~ agricultural emissions contribute $\sim$30\% of total global
-emissions, whereas in the unharmonized case with default methods, \gls{sas}
+emissions, whereas in the harmonized case with default methods, \gls{sas}
 comprises $\sim$50\% of global emissions by 2100. By using a
@@
 were analyzed in Section \ref{sec:results}. Of the remaining trajectories,
-harmonization method overrides were applied, and a discussion was provided
+harmonization method overrides were applied, and  discussion was provided
 detailing why overrides were deemed necessary.
@@
 historical and future emissions trajectories. Furthermore, the automated
-approach continues to scale well as models become more detailed in both the
+approach should continue to scale well as models become more detailed in both the
 regional and sectoral dimensions. Finally, expert opinion is still allowed to
@@
 software and underlying methodology. As with any software project, additional
-users will provide use cases for more robust handling of input/output concerns
+users will provide use cases for more robust handling of input/output issues
 and corner cases. Further configuration parameters can also be added in the

### 2018-02-07 paper.tex
spell checked and updated schematic

@@
       The various objects and their relation to one another in the \code{aneris}
-      code base as well as a short description of their scope.  
+      code base as well as a short description of their scope of concern.  
     }
@@
 
-In order to help identify cases where overrides may be needed,
-harmonization \textit{diagnostics} are provided which report the relative
-difference between harmonized and unharmonized trajectories at their mid and
-end-points when these values exceed specified threshold. The diagnostic reporting thresholds  are configurable by the user, but defaults of $400$\%
-and $200$\%, respectively, are provided based on experiences of the authors'
-use of \code{aneris} to date.
+In order to help identify cases where overrides may be needed, harmonization
+\textit{diagnostics} are provided which report the relative difference between
+harmonized and unharmonized trajectories at their mid and end-points when these
+values exceed specified thresholds. The diagnostic reporting thresholds are
+configurable by the user, but defaults of $400$\% and $200$\%, respectively, are
+provided based on experiences of the authors' use of \code{aneris} to date.
 
@@
 \code{aneris} harmonization procedure, we focus on the harmonization of results
-of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Harmonization results for two scenarios from
-the \gls{ssp} scenario library\footnote{We refer the reader to the broad
-  literature discussing the SSPs, e.g.,
+of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Harmonization
+results for two scenarios from the \gls{ssp} scenario library\footnote{We refer
+  the reader to the broad literature discussing the SSPs, e.g.,
   \cite{vuuren_new_2013,oneill_scenario_2016, oneill_new_2013} for a more in
-  depth discussion of the scenario architecture and design.} are presented here. We use the
-SSP2-reference\cite{Riahi2017153,Rao2017346}, or ``middle of the road'', scenario
-(referred to as SSP2-Ref) as an example because MESSAGE-GLOBIOM is
-the marker scenario\footnote{The ``marker'' scenario concept is used to
-  designate the archetype scenario used as a reference within each SSP scenario family. See
-  \cite{vuuren_new_2013} and \cite{Riahi2017153} for a more lengthy
-  description.} for this \gls{ssp}. This SSP2 scenario lies between two
-\glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing in this concept is the energy inbalance at the top of the atmsophere caused by anthropogenic influences relative to a pre-industrial reference point.
-Higher radiative forcing leads to larger global changes, such as surface temperture.
-We refer the reader to \cite{IPCCWG1PhysicalStocker2013}
-  for a more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
+  depth discussion of the scenario architecture and design.} are presented
+here. We use the SSP2-reference\cite{Riahi2017153,Rao2017346}, or ``middle of
+the road'', scenario (referred to as SSP2-Ref) as an example because
+MESSAGE-GLOBIOM is the marker scenario\footnote{The ``marker'' scenario concept
+  is used to designate the archetype scenario used as a reference within each
+  SSP scenario family. See \cite{vuuren_new_2013} and \cite{Riahi2017153} for a
+  more lengthy description.} for this \gls{ssp}. This SSP2 scenario lies between
+two \glspl{rcp}, 6 and 8.5, with a radiative forcing\footnote{Radiative forcing
+  in this context is the energy imbalance at the top of the atmosphere caused by
+  anthropogenic influences relative to a pre-industrial reference point. Higher
+  radiative forcing leads to larger global changes, such as surface
+  temperature. We refer the reader to \cite{IPCCWG1PhysicalStocker2013} for a
+  more detailed discussion.} level of approximately 6.5 Wm$^{-2}$. We
 additionally present the results for the SSP2-based mitigation scenario leading
@@
 
-MESSAGE-GLOBIOM includes a representation of 11 distinct regions, and harmonization is performed at the native regional resolution of each model. Historical
-data is taken from previously described \gls{luc} and anthropogenic sources,
-which comprise 10 separate pollutant and \gls{ghg} species and 12 sectors shown
-in Table \ref{tab:sp}. A total of 970 distinct trajectories were harmonized for
-each scenario, and therefore 1940 trajectories were harmonized in total for these two illustrative scenarios. 
-\noxx
-generated from the Energy sector provides an example of an emissions species and
-sector in which all regions were satisfactorily harmonized with the default
+MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
+mapped directly to the 5-region definition used in the RCPs (Figure
+\ref{fig:regions}); harmonization is performed directly on the native regional
+spatial resolution. Historical data is taken from previously described \gls{luc}
+and anthropogenic sources, which comprise 10 separate pollutant and \gls{ghg}
+species and 12 sectors shown in Table \ref{tab:sp}. A total of 970 distinct
+trajectories were harmonized for each scenario, and therefore 1940 trajectories
+were harmonized in total for these two illustrative scenarios.  \noxx generated
+from the Energy sector provides an example of an emissions species and sector in
+which all regions were satisfactorily harmonized with the default
 methods. Figure \ref{fig:nox} shows the results of harmonization in Asia, and
@@
 for each harmonized trajectory. Importantly, the default methods provide
-regional trajectories for both the reference and mitigation scenarios that
-match the general unharmonized model behavior and also result in global aggregated emissions
-that are similar to the unharmonized trajectory.
+regional trajectories for both the reference and mitigation scenarios that match
+the general unharmonized model behavior and also result in global aggregated
+emissions that are similar to the unharmonized trajectory.
 
@@
       markers) emissions are presented for SSP2 (purple lines) and SSP2-45 (red
-      lines) scenarios. The left column shows the deault harmonization and the right column shows harmonization with overrides. Panel \textbf{a} shows  trajectories for the \gls{sas} region
+      lines) scenarios. The left column shows the default harmonization and the right column shows harmonization with overrides. Panel \textbf{a} shows  trajectories for the \gls{sas} region
       while panels \textbf{b} and \textbf{c} show  trajectories for the
@@
       model results in both scenarios. The distortion is large enough to affect
-      global results, as shown in Panel \textbf{c}.  The use of overrides (right column) results in better consistantcy with the unharmonized model scenario at the regional and global levels. }
+      global results, as shown in Panel \textbf{c}.  The use of overrides (right column) results in better consistency with the unharmonized model scenario at the regional and global levels. }
   \end{center}

### 2018-02-07 paper.tex
clarify calculation and update intro text for steves comment

@@
 incorporate data on energy systems, land use, economics, demographics and
-emissions sources and concentrations, among other data, in order to provide
-consistent existing trajectories of modeled variables. The models then provide
+emissions sources and concentrations, among other data, in order to provide a
+consistent starting point for future projections. The models then provide
 estimates of future trajectories of these variables under various socio-economic
@@
 species and 12 sectors shown in Table \ref{tab:sp}. A total of 970 distinct
-trajectories were harmonized for each scenario, and therefore 1940 trajectories
-were harmonized in total for these two illustrative scenarios.  \noxx generated
-from the Energy sector provides an example of an emissions species and sector in
-which all regions were satisfactorily harmonized with the default
-methods. Figure \ref{fig:nox} shows the results of harmonization in Asia, and
-Table \ref{tab:nox} describes the parameters that underlie the choice of method
-for each harmonized trajectory. Importantly, the default methods provide
-regional trajectories for both the reference and mitigation scenarios that match
-the general unharmonized model behavior and also result in global aggregated
-emissions that are similar to the unharmonized trajectory.
+trajectories\footnote{Table \ref{tab:sp} compiles 24 global trajectories and 86
+  regional trajectories. Therefore, with 11 model regions, 970 total
+  trajectories are harmonized.} were harmonized for each scenario, and therefore
+1940 trajectories were harmonized in total for these two illustrative scenarios.
+\noxx generated from the Energy sector provides an example of an emissions
+species and sector in which all regions were satisfactorily harmonized with the
+default methods. Figure \ref{fig:nox} shows the results of harmonization in
+Asia, and Table \ref{tab:nox} describes the parameters that underlie the choice
+of method for each harmonized trajectory. Importantly, the default methods
+provide regional trajectories for both the reference and mitigation scenarios
+that match the general unharmonized model behavior and also result in global
+aggregated emissions that are similar to the unharmonized trajectory.
 

### 2018-02-07 paper.tex
update refs, add airchemmip, add make bib target

@@
 \newacronym{smip}{ScenarioMIP}{Scenario Model Intercomparison Project}
+\newacronym{amip}{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
 \newacronym{ceds}{CEDS}{Community Emissions Data System}
@@
 ongoing, worldwide scientific community efforts in the \gls{cmip6}
-\cite{eyring_overview_2016}, which is utilizing a number of marker \gls{ssp} scenarios
-developed by the \gls{iam} community (\gls{smip})
-\cite{oneill_scenario_2016}. These trajectories are endogenously calculated by
-modeling the individual technologies and sectors that contribute towards the
-emissions of different air pollutants and \glspl{ghg} as well as various
-mitigation technologies. However, the historical emissions starting points of
-models can differ by large amounts depending on the region, sector, and
-emissions species.
+\cite{eyring_overview_2016}, which is utilizing a number of marker \gls{ssp}
+scenarios developed by the \gls{iam} community
+(\gls{smip}\cite{oneill_scenario_2016},
+\gls{amip}\cite{collins_aerchemmip:_2017}). These trajectories are endogenously
+calculated by modeling the individual technologies and sectors that contribute
+towards the emissions of different air pollutants and \glspl{ghg} as well as
+various mitigation technologies. However, the historical emissions starting
+points of models can differ by large amounts depending on the region, sector,
+and emissions species.
 
@@
 historical emissions data set for both anthropogenic emissions (i.e., the
-\gls{ceds} \cite{hoesly_historical_2017} and open-burning \gls{luc} emissions
+\gls{ceds} \cite{hoesly_historical_2018} and open-burning \gls{luc} emissions
 \cite{van_marle_historic_2017}) which, in conjunction with the \gls{ssp}
@@
       The distribution of $c_v$ values for \gls{luc} and non-\gls{luc} historical
-      trajectories is shown. \gls{ceds} historical data \cite{hoesly_historical_2017}
+      trajectories is shown. \gls{ceds} historical data \cite{hoesly_historical_2018}
       is used for non-\gls{luc} data and \cite{van_marle_historic_2017} is used for

### 2018-02-07 paper.tex
add among others
add c footnotes for all ag sectors

@@
 (\gls{smip}\cite{oneill_scenario_2016},
-\gls{amip}\cite{collins_aerchemmip:_2017}). These trajectories are endogenously
-calculated by modeling the individual technologies and sectors that contribute
-towards the emissions of different air pollutants and \glspl{ghg} as well as
-various mitigation technologies. However, the historical emissions starting
-points of models can differ by large amounts depending on the region, sector,
-and emissions species.
+\gls{amip}\cite{collins_aerchemmip:_2017}, among others). These trajectories are
+endogenously calculated by modeling the individual technologies and sectors that
+contribute towards the emissions of different air pollutants and \glspl{ghg} as
+well as various mitigation technologies. However, the historical emissions
+starting points of models can differ by large amounts depending on the region,
+sector, and emissions species.
 
@@
     \hline
-    Black Carbon (BC)                 & Agricultural Waste Burning          \\
+    Black Carbon (BC)                 & Agricultural Waste Burning\footnoteref{x_co2}          \\
     Hexafluoroethane (C\textsubscript{2}F\textsubscript{6})
@@
       historical data.}
-                                      & Agriculture                         \\
+                                      & Agriculture\footnoteref{x_co2}                         \\
     Tetrafluoromethane (CF\textsubscript{4})
@@
       are harmonized for each model region.}
-                                      & Forest Burning                      \\
-    Carbon Monoxide (CO)              & Grassland Burning                   \\
+                                      & Forest Burning\footnoteref{x_co2}                      \\
+    Carbon Monoxide (CO)              & Grassland Burning\footnoteref{x_co2}                   \\
     Hydrofluorocarbons (HFCs)

### 2018-02-07 paper.tex
one chief import to should be considered

@@
 
-The goals of any harmonization exercise is threefold: aligning model results in
+The goals of any harmonization exercise are threefold: aligning model results in
 the harmonization year to a common historical data source, faithfully
@@
 emissions, as is possible in scenarios designed to depict the deployment of
-climate mitigation policies with large \cotwo reductions and storage,
-then end-of-century emissions characteristics can be of chief import in order to
+climate mitigation policies with large \cotwo reductions and storage, then
+end-of-century emissions characteristics should be considered in order to
 faithfully match the design parameters of the original scenarios, such as global
@@
 was provided in Section \ref{sec:meths}. \code{aneris} was able to
-satisfactorily harmonized over 96\% of the 1940 individual trajectories that
+satisfactorily harmonize over 96\% of the 1940 individual trajectories that
 were analyzed in Section \ref{sec:results}. Of the remaining trajectories,

### 2018-02-08 paper.tex
adding suggested edits from Keywan

@@
 \texttt{aneris}, first made available in 2017, is available online at
-\url{https://github.com/gidden/aneris} as a free and open-source Python software
+\url{https://github.com/iiasa/aneris} as a free and open-source Python software
 library (approximately 2000 lines of code). The \texttt{aneris} software was
@@
 page of this manuscript. Documentation for the \texttt{aneris} Python package,
-including software requirements, is available
-\href{http://mattgidden.com/aneris/}{online}.
+including software requirements, is available online at
+\url{http://data.ene.iiasa.ac.at/gidden/aneris/}.
 
@@
 and energy futures. Recently, the \gls{iam} community has developed scenarios
-based on the \glspl{ssp} \cite{fujimori_ssp3:_2017, van_vuuren_energy_2017,
-  calvin_ssp4:_2017, fricko_marker_2017, kriegler_fossil-fueled_2017} which
-quantify a variety of potential global futures. The \glspl{ssp} are designed to
-be used in research that include \gls{esm} simulations, climate impact,
-adaptation and climate mitigation studies \cite{vuuren_new_2013}.
+based on the \glspl{ssp} \cite{Riahi2017153} which quantify a variety of
+potential global futures. The \glspl{ssp} are designed to be used in research
+that include \gls{esm} simulations, climate impact, adaptation and climate
+mitigation studies \cite{oneill_new_2013, vuuren_new_2013}.
 
@@
   numerous and have a long history in the scientific literature. Various IAMs
-  have collaborated to produce
-  \href{http://themasites.pbl.nl/models/advance/index.php/ADVANCE_wiki}{community
-    IAM documentation} which readers can access for a full treatment of model
-  implementation and features.}, including simulation and optimization, the core
-inputs and outputs are similar across different models. Modeling teams
-incorporate data on energy systems, land use, economics, demographics and
-emissions sources and concentrations, among other data, in order to provide a
-consistent starting point for future projections. The models then provide
-estimates of future trajectories of these variables under various socio-economic
-and technological assumptions as well as proposed policy constraints, e.g.,
-targets for future \gls{ghg} emissions.
+  have collaborated to produce community IAM documentation (available online:
+  \url{http://themasites.pbl.nl/models/advance/index.php/ADVANCE_wiki}) which
+  readers can access for a full treatment of model implementation and
+  features.}, including simulation and optimization, the core inputs and outputs
+are similar across different models. Modeling teams incorporate data on energy
+systems, land use, economics, demographics and emissions sources and
+concentrations, among other data, in order to provide a consistent starting
+point for future projections. The models then provide estimates of future
+trajectories of these variables under various socio-economic and technological
+assumptions as well as proposed policy constraints, e.g., targets for future
+\gls{ghg} emissions.
 
@@
 When participating in intercomparison exercises in which a consistent historical
-starting point is required (e.g., in \gls{cmip6}), model teams incorporate a single,
-common historical data set through \textit{harmonization}. Harmonization
-refers to the process of adjusting model results to match a selected
-historical time series such that the resulting future trajectories are also
-consistent with the original modeled results. In the emissions context, this
-means that each individual combination of model region, model sector, and
-emissions species must be harmonized. Depending on the total number of model
-regions, sectors, and emissions species, this can require the selection of
-thousands to tens-of-thousands of harmonization methods.
+starting point is required (e.g., in \gls{cmip6}), model teams incorporate a
+single, common historical data set through \textit{harmonization}. Harmonization
+refers to the process of adjusting model results to match a selected historical
+time series such that the resulting future trajectories are consistent with the
+original modeled results and provide a smooth transition from the common
+historical data. In the emissions context, this means that each individual
+combination of model region, model sector, and emissions species must be
+harmonized. Depending on the total number of model regions, sectors, and
+emissions species, this can require the selection of thousands to
+tens-of-thousands of harmonization methods.
 
@@
 harmonization software \code{aneris} \cite{matthew_gidden_2017_802832}, written
-in the Python programming language (detailed documentation is available
-\href{http://mattgidden.com/aneris/}{online}). Section \ref{sec:meths} provides
-a detailed description of the underlying mathematical components of
+in the Python programming language (detailed documentation is available online
+at \url{http://data.ene.iiasa.ac.at/gidden/aneris/}). Section \ref{sec:meths}
+provides a detailed description of the underlying mathematical components of
 \code{aneris} as well as the procedural workflow. A case study of applying the
@@
 
-The goals of any harmonization exercise are threefold: aligning model results in
-the harmonization year to a common historical data source, faithfully
+The goals of any scenario harmonization exercise are threefold: aligning model
+results in the harmonization year to a common historical data source, faithfully
 representing the original \gls{iam}’s internal consistency between the driver of
@@
 harmonization method choice, discussed in Section \ref{sec:tree}, in order to
-balance each of these concerns and use cases in a robust and automated manner.
+balance each of these concerns and use cases in a robust, systematic,
+reproducible, and transparent manner.
 
@@
 \gls{iam} emission results are provided along temporal (normally half decade or
-decade), spatial (i.e., model regions), gas species, and sectoral
+decade), spatial (i.e., model regions), emissions species, and sectoral
 dimensions. Each individual temporal trajectory, i.e., unique combinations of
@@
 i.e., how well a given harmonization algorithm performs, depends on a number of
-factors. Of chief import is the faithful representation of the original,
-unharmonized, trajectory as well as the representation of negative trajectories
+factors. Of chief import is the faithful representation of the original
+unharmonized trajectory as well as the representation of negative trajectories
 (i.e., if a trajectory becomes negative, both the timing and total magnitude
@@
 IMAGE\cite{van_vuuren_energy_2017}, GCAM4\cite{calvin_ssp4:_2017},
-MESSAGE-GLOBIOM\cite{fricko_marker_2017}, and
+MESSAGE-GLOBIOM\cite{fricko_marker_2017, message_globiom_2016}, and
 REMIND-MAGPIE\cite{kriegler_fossil-fueled_2017}, each of which have varying
@@
 \end{minted}
-\caption{High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://mattgidden.com/aneris/config.html}{tutorial}}
+\caption{High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://data.ene.iiasa.ac.at/gidden/aneris/tutorial.html}{tutorial}}
 \label{lst:tutorial}
@@
   --rc RC               Runcontrol YAML file (see
-                        http://mattgidden.com/aneris/config.html for
+                        http://data.ene.iiasa.ac.at/gidden/aneris/config.html for
                         examples).
@@
 (with examples provided
-\href{http://mattgidden.com/aneris/config.html}{online}). The primary mechanism
+\href{http://data.ene.iiasa.ac.at/gidden/aneris/config.html}{online}). The primary mechanism
 by which users control the process is by providing \textit{override} methods for
@@
 \code{aneris} harmonization procedure, we focus on the harmonization of results
-of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017}. Harmonization
-results for two scenarios from the \gls{ssp} scenario library\footnote{We refer
-  the reader to the broad literature discussing the SSPs, e.g.,
-  \cite{vuuren_new_2013,oneill_scenario_2016, oneill_new_2013} for a more in
-  depth discussion of the scenario architecture and design.} are presented
-here. We use the SSP2-reference\cite{Riahi2017153,Rao2017346}, or ``middle of
-the road'', scenario (referred to as SSP2-Ref) as an example because
+of the \gls{iam} MESSAGE-GLOBIOM \cite{fricko_marker_2017,
+  message_globiom_2016}. Harmonization results for two scenarios from the
+\gls{ssp} scenario library\footnote{We refer the reader to the broad literature
+  discussing the SSPs, e.g., \cite{Riahi2017153, oneill_new_2013,
+    vuuren_new_2013,oneill_scenario_2016} for a more in depth discussion of the
+  scenario architecture and design. Model results for various SSP scenarios are
+  available online at \url{https://tntcat.iiasa.ac.at/SspDb}} are presented
+here. We use the SSP2-reference\cite{fricko_marker_2017,Rao2017346}, or ``middle
+of the road'', scenario (referred to as SSP2-Ref) as an example because
 MESSAGE-GLOBIOM is the marker scenario\footnote{The ``marker'' scenario concept
   is used to designate the archetype scenario used as a reference within each
-  SSP scenario family. See \cite{vuuren_new_2013} and \cite{Riahi2017153} for a
+  SSP scenario family. See \cite{Riahi2017153} and \cite{vuuren_new_2013} for a
   more lengthy description.} for this \gls{ssp}. This SSP2 scenario lies between
@@
   emissions) to estimate future radiative forcing and mean global temperature
-  response. We refer the reader to the
-  \href{http://wiki.magicc.org/index.php?title=Model_Description}{MAGICC wiki}
-  for a more in-depth description.} \cite{meinshausen2011emulating,
-  meinshausen2011rcp}, for each harmonized and unharmonized scenario
-respectively as shown in Figure \ref{fig:forcing}. We find that the change due
-to harmonization is small, ranging between 1 and 2.5\% over the modeled time
-period. Relative near-term differences persist in the mitigation case (SSP-4.5)
-because differences in near-term emissions define to a larger degree the
-longer-term forcing outcome due to the cumulative nature of long-lived climate
-forcers like \cotwo.  The resulting difference in forcing in 2100 is 0.04
-$\frac{\text{W}}{\text{m}^2}$ for SSP2-4.5 and 0.01
+  response. We refer the reader to the MAGICC wiki (available online:
+  \url{http://wiki.magicc.org}) for a more in-depth description.}
+\cite{meinshausen2011emulating, meinshausen2011rcp}, for each harmonized and
+unharmonized scenario respectively as shown in Figure \ref{fig:forcing}. We find
+that the change due to harmonization is small, ranging between 1 and 2.5\% over
+the modeled time period. Relative near-term differences persist in the
+mitigation case (SSP-4.5) because differences in near-term emissions define to a
+larger degree the longer-term forcing outcome due to the cumulative nature of
+long-lived climate forcers like \cotwo.  The resulting difference in forcing in
+2100 is 0.04 $\frac{\text{W}}{\text{m}^2}$ for SSP2-4.5 and 0.01
 $\frac{\text{W}}{\text{m}^2}$ for SSP2-Ref, both of which are well within

### 2018-02-08 paper.tex
update data.ene to software.ene

@@
 including software requirements, is available online at
-\url{http://data.ene.iiasa.ac.at/gidden/aneris/}.
+\url{http://software.ene.iiasa.ac.at/aneris/}.
 
@@
 in the Python programming language (detailed documentation is available online
-at \url{http://data.ene.iiasa.ac.at/gidden/aneris/}). Section \ref{sec:meths}
+at \url{http://software.ene.iiasa.ac.at/aneris/}). Section \ref{sec:meths}
 provides a detailed description of the underlying mathematical components of
@@
 \end{minted}
-\caption{High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://data.ene.iiasa.ac.at/gidden/aneris/tutorial.html}{tutorial}}
+\caption{High-level user interaction with the \texttt{HarmonizationDriver} taken from the online \href{http://software.ene.iiasa.ac.at/aneris/tutorial.html}{tutorial}}
 \label{lst:tutorial}
@@
   --rc RC               Runcontrol YAML file (see
-                        http://data.ene.iiasa.ac.at/gidden/aneris/config.html for
+                        http://software.ene.iiasa.ac.at/aneris/config.html for
                         examples).
@@
 (with examples provided
-\href{http://data.ene.iiasa.ac.at/gidden/aneris/config.html}{online}). The primary mechanism
+\href{http://software.ene.iiasa.ac.at/aneris/config.html}{online}). The primary mechanism
 by which users control the process is by providing \textit{override} methods for
